---
title: "DANL 200: Classwork 6 - filter, arrange, select, and mutate"
date: 2023-10-13
execute: 
  warning: false
  message: false
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(nycflights13)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```

# Question 1. Filtering with `filter()`

## Q1a.

- Find all flights that
  - Had an arrival delay of two or more hours
  - Flew to Houston (`IAH` or `HOU`)
  - Were operated by United, American, or Delta
  - Departed in summer (July, August, and September)
  - Arrived more than two hours late, but didn’t leave late
  - Were delayed by at least an hour, but made up over 30 minutes in flight
  - Departed between midnight and 6am (inclusive)

<br><br>

## Q1b.

- How many flights have a missing `dep_time`? What other variables are missing? What might these rows represent?

<br><br>

## Q1c.

- Why is `NA^0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)



<br><br><br>


# Question 2. Arrange rows with `arrange()`

## Q2a.

- How could you use `arrange()` to sort all missing values to the start? (Hint: use `is.na()`).
```{r}
#| echo: true
df <- data.frame(num = c(1, 4, NA, 2, NA, 3, 7),
                 chr = c("A", "A", "A", "B", "X", "Z", "D"))
```

<br>

## Q2b.

- Sort flights to find the most delayed flights. Find the flights that left earliest.

<br>

## Q2c.

- Sort flights to find the fastest (highest speed) flights.

<br>

## Q2d.

- Which flights traveled the farthest? Which traveled the shortest?

<br><br><br>


# Question 3. Select columns with `select()`

## Q3a.

- Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from flights.

<br>

## Q3b.

- Does the result of running the following code surprise us? How do the select helpers deal with case by default? How can we change that default?

```{r, echo = T, eval = F}
select(flights, contains("TIME"))
```

<br>

## Q3c.
- Rename `air_time` to `air_time_min` to indicate units of measurement and move it to the beginning of the data frame.


<br><br><br>

# Question 4. Add new variables with `mutate()`

## Q4a.
 
- Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because their values are represented with format HHMM or HMM (H for hour and M for minute).
  - Convert them to a more convenient representation of number of **minutes** since midnight.

```{r}
#| echo: false
q4a <- flights %>% 
  select(dep_time, sched_dep_time, dep_delay)  %>% 
  mutate(dep_time = ifelse(dep_time == 2400, 0, dep_time),
         sched_dep_time = ifelse(sched_dep_time == 2400, 0, sched_dep_time)
         ) %>% 
  mutate(dep_time_hour = dep_time %/% 100,
         dep_time_minute = dep_time %% 100,
         dep_time_new = dep_time_hour * 60 + dep_time_minute,
         .after = dep_time) %>% 
  mutate(sched_dep_time_hour = sched_dep_time %/% 100,
         sched_dep_time_minute = sched_dep_time %% 100,
         sched_dep_time_new = sched_dep_time_hour * 60 + sched_dep_time_minute,
         .after = sched_dep_time)
```

<br>

## Q4b. 

- Compare `air_time` with `arr_time - dep_time`. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
#| echo: false
q4b <- flights %>% 
  select(air_time, arr_time, dep_time) %>% 
  mutate(dep_time = ifelse(dep_time == 2400, 0, dep_time),
         arr_time = ifelse(dep_time == 2400, 0, arr_time),
         ) %>% 
  mutate(dep_time_hour = dep_time %/% 100,
         dep_time_minute = dep_time %% 100,
         dep_time = dep_time_hour * 60 + dep_time_minute,
         .after = dep_time) %>% 
  mutate(arr_time_hour = arr_time %/% 100,
         arr_time_minute = arr_time %% 100,
         arr_time = arr_time_hour * 60 + arr_time_minute,
         .after = arr_time) %>% 
  mutate(air_time2 = arr_time - dep_time,
         diff_air_time = air_time - air_time2) %>% 
  relocate(diff_air_time, arr_time, air_time2)

# The flight passes midnight, so arr_time < dep_time. In these cases, the difference in airtime should be by 24 hours (1,440 minutes).

# The flight crosses time zones, and the total air time will be off by hours (multiples of 60). All flights in flights departed from New York City and are domestic flights in the US. This means that flights will all be to the same or more westerly time zones. Given the time-zones in the US, the differences due to time-zone should be 60 minutes (Central) 120 minutes (Mountain), 180 minutes (Pacific), 240 minutes (Alaska), or 300 minutes (Hawaii).
```

```{r}
#| echo: false
#| eval: false
ggplot(q4b) +
  geom_histogram(aes(x = diff_air_time), binwidth = 1)

# there are many flights in which the difference between air time and local arrival and departure times is not divisible by 60.

# To fix these time-zone issues, I would want to convert all the times to a date-time to handle overnight flights, and from local time to a common time zone, most likely UTC, to handle flights crossing time-zones. 


# What else might be going on? There seem to be too many problems for this to be data entry problems.

# It appears that the air_time variable refers to flight time, which is defined as the time between wheels-off (take-off) and wheels-in (landing).
# But the flight time (dep_time, arr_time) does not include time spent on the runway taxiing to and from gates.
  # air_time <= arr_time - dep_time

```

<br>

## Q4c.

- Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you expect those three numbers to be related?

```{r}
#| echo: false
q4c <- q4a %>% 
  mutate(dep_delay_new = dep_time_new - sched_dep_time_new,
         diff_dep_delay = dep_delay - dep_delay_new,
         .before = 1) %>% 
  relocate(dep_delay)

# All of these discrepancies are exactly equal to 1440 (24 hours), and the flights with these discrepancies were scheduled to depart later in the day.
```

<br>

## Q4d.

- Find the 20 most delayed flights using a ranking function. How do you want to handle ties? 

```{r}
#| echo: false
q4d <- flights %>% 
  mutate(rank_min = min_rank(-dep_delay),
         rank_dense = dense_rank(-dep_delay),
         rank_rownum = row_number(-dep_delay)
         ) %>% 
  filter( rank_min <= 20 | rank_dense <= 20 | rank_rownum <= 20 ) %>% 
  select(dep_delay, rank_min, rank_dense, rank_rownum) %>% 
  arrange(-dep_delay)
```

<br>

## Q4e.
- Read the following CSV file, `TSLA.csv`, as a data.frame with the name `tsla`:
```{r}
#| echo: true
#| eval: false
tsla <- read_csv("https://bcdanl.github.io/data/TSLA_2022_2023.csv")
```

- Describe the distribution of the percentage change in `Close` price of Tesla stock.

- Compare the daily trend of cumulative `Volume` of Tesla stock in 2022 with that in 2023.


