---
title: "DANL 200: Homework Assignment 3"
date: "`r format(Sys.time(), '%B %d, %Y')`"
execute: 
  echo: true
  eval: true
  warning: false
  message: false
  fig-width: 9

---
```{r}
#| eval: true
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
```

# Question 1

Load the data.frame for Question 1.

```{r}
#| echo: true
#| eval: true
restaurant <- read_csv('https://bcdanl.github.io/data/DOHMH_NYC_Restaurant_Inspection.csv')
```

```{r, result = 'asis', message = F, warning = F}
#| echo: false
rmarkdown::paged_table(restaurant)
```

## Variable Description

-   `CAMIS`:
    -   This is an unique identifier for the entity (restaurant);
    -   10-digit integer
-   `DBA`:
    -   This field represents the name (doing business as) of the entity (restaurant);
    -   Public business name, may change at discretion of restaurant owner
-   `BORO`:
    -   Borough in which the entity (restaurant) is located.;
    -   • `1` = MANHATTAN
    -   • `2` = BRONX
    -   • `3` = BROOKLYN
    -   • `4` = QUEENS
    -   • `5` = STATEN ISLAND
    -   • `0` = Missing;
-   `CUISINE DESCRIPTION`:
    -   This field describes the entity (restaurant) cuisine.
-   `ACTION`:
    -   This field represents the actions that is associated with each restaurant inspection. ;
    -   • Violations were cited in the following area(s).
    -   • No violations were recorded at the time of this inspection.
    -   • Establishment re-opened by DOHMH
    -   • Establishment re-closed by DOHMH
    -   • Establishment Closed by DOHMH.
    -   • Violations were cited in the following area(s) and those requiring immediate action were addressed.
-   `VIOLATION CODE`:
    -   Violation code associated with an establishment (restaurant) inspection
-   `VIOLATION DESCRIPTION`:
    -   Violation description associated with an establishment (restaurant) inspection
-   `CRITICAL FLAG`:
    -   Indicator of critical violation;
    -   • `Critical`
    -   • `Not Critical`
    -   • `Not Applicable`;
    -   Critical violations are those most likely to contribute to food-borne illness
-   `SCORE`:
    -   Total score for a particular inspection;
-   `GRADE`:
    -   Grade associated with the inspection;
    -   • `N` = Not Yet Graded
    -   • `A` = Grade A
    -   • `B` = Grade B
    -   • `C` = Grade C
    -   • `Z` = Grade Pending
    -   • `P` = Grade Pending issued on re-opening following an initial inspection that resulted in a closure

<br>

## Q1a.

What are the mean, standard deviation, first quartile, median, third quartile, and maximum of `SCORE` for each `GRADE` of restaurants?

```{r}
restaurant %>% group_by(GRADE) %>% skim(SCORE) %>% select(-n_missing)
```

<br>

## Q1b.

-   How many restaurants with a `GRADE` of `A` are there in NYC?

-   How much percentage of restaurants in NYC are a `GRADE` of `C`?

```{r}
freq <- as.data.frame( table(restaurant$GRADE) )
prop <- as.data.frame( 100 * prop.table(table(restaurant$GRADE)) )
```

```{r, result = 'asis', echo = F, message = F, warning = F}
#| echo: false
rmarkdown::paged_table(freq)
```

```{r, result = 'asis', echo = F, message = F, warning = F}
#| echo: false
rmarkdown::paged_table(prop)
```

<br>

## Q1c.

Provide both (1) ggplot code and (2) a simple comment to describe how the distribution of `SCORE` varies by `GRADE` and `CRITICAL FLAG`.


-   Boxplots

```{r}
ggplot(restaurant) +
  geom_boxplot(aes(x = SCORE, y = GRADE, fill = GRADE) ) +
  facet_grid( `CRITICAL FLAG` ~ . )
```

-   Histograms

```{r}
ggplot(restaurant) +
  geom_histogram(aes(x = SCORE), binwidth = 1 ) +
  facet_grid( `CRITICAL FLAG` ~ GRADE )
```


-   Histograms with varying x- scale across facets

```{r}
ggplot(restaurant) +
  geom_histogram(aes(x = SCORE), binwidth = 1 ) +
  facet_grid( `CRITICAL FLAG` ~ GRADE, scales = 'free_x' )
```


-   Histograms with varying x- and y- scales across facets

```{r}
ggplot(restaurant) +
  geom_histogram(aes(x = SCORE), binwidth = 1 ) +
  facet_wrap( `CRITICAL FLAG` ~ GRADE, scales = 'free' )
```

-   Mostly,
    -   The values of `SCORE` for `GRADE` `A` ranges from 0 to 13.
    -   The values of `SCORE` for `GRADE` `B` ranges 13 to 27.
    -   The values of `SCORE` for `GRADE` `C` ranges 24 to 75.
-   For `Not Critical` type, two `SCORE` values around 1 and 12 are most common, while 12 is the single most common `SCORE` value for `Critical` type.

<br>

## Q1d.

Provide both (1) ggplot code and (2) a simple comment to describe how the proportion of `CRITICAL FLAG` varies by `GRADE` and `BORO`.

```{r}
ggplot(restaurant) +
  geom_bar(aes(x = `CRITICAL FLAG`,
               y = after_stat(prop), group = 1)) +
  facet_grid( GRADE ~ BORO )
```

-   For `GRADE` `A`, the probability distribution of `CRITICAL FLAG` are similar across `BORO`s.

-   For `GRADE` `B`, the restaurants in `Staten Island` are more likely to be `Critical` than in other `BORO`s.

-   For `GRADE` `C`, the restaurants in `Bronx` are more likely to be `Critical` than in other `BORO`s.

<br>

## Q1e.

For the 10 most common `CUISINE DESCRIPTION` values, find the `CUISINE DESCRIPTION` value that has the highest proportion of `GRADE` `A`.

```{r}
q2e <- restaurant %>% 
  group_by(`CUISINE DESCRIPTION`) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(dense_rank(-n) <= 10) %>% 
  group_by(`CUISINE DESCRIPTION`, GRADE) %>% 
  count() %>% 
  group_by(`CUISINE DESCRIPTION`) %>% 
  mutate(prop_A = n / sum(n)) %>% 
  filter(GRADE == 'A') %>% 
  arrange(-prop_A)
```

```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(q2e)
```

<br>

## Q1f.

-   Find the 3 most common names of restaurants (`DBA`) in each `BORO`.
    -   If the third most common `DBA` values are multiple, please include all the `DBA` values.
-   Overall, which `DBA` value is most common in NYC?

```{r}
q2f <- restaurant %>% 
  select(DBA, BORO) %>% 
  group_by(BORO, DBA) %>% 
  summarize(n = n()) %>% 
  mutate(ranking = dense_rank(-n)) %>% 
  filter(ranking <= 3) %>% 
  arrange(BORO, ranking)

q2f_ <- restaurant %>% 
  group_by(DBA) %>% 
  count() %>% 
  arrange(-n)
```

```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(q2f)
rmarkdown::paged_table(q2f_)
```

-   Note that `chipotle mexican grill` and `subway` are both the third most popular franchise/chain in `Manhattan`.

-   Overall, `dunkin` is the most popular franchise/chain in NYC.

<br>

## Q1g.

For all the `DBA` values that appear in the result of Q1f, find the `DBA` value that is most likely to commit critical violation.

```{r}
q2g <- restaurant %>% 
  filter(DBA %in% q2f$DBA) %>% 
  group_by(DBA, `CRITICAL FLAG`) %>% 
  count() %>% 
  group_by(DBA) %>% 
  mutate(lag_n = lag(n),
         tot = sum(n),
         prop_crit = lag_n / tot) %>% 
  select(DBA, prop_crit) %>% 
  filter(!is.na(prop_crit)) %>% 
  arrange(-prop_crit)
```

```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(q2g)
```

-   Among popular franchises/chains, `subway` is most likely to commit `Critical` violation in NYC.
