---
title: "DANL 200: Homework Assignment 2 - Example Answers"
date: "`r format(Sys.time(), '%B %d, %Y')`"
execute: 
  echo: true
  eval: true
  warning: false
  message: false

---
```{r setup}
#| include: false
#| 
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggthemes)
library(ggthemr)
library(hrbrthemes)
library(hexbin)
library(ggforce)
library(skimr)

knitr::opts_chunk$set(fig.width=8, fig.height=5,
                      echo = T, eval = T, 
                      message=F, warning = F, fig.cap = " ")  

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5))
                ))
```

# Question 1. Working with Customer Data with Health Insurance

Read the data file, `custdata_rev.csv`, as the data.frame object with the name, `health_cust`:

```{r}
#| echo: true
library(tidyverse)
library(skimr)

health_cust <- read_csv(
  'https://bcdanl.github.io/data/custdata_rev.csv'
)
```



## Variable Description for `health_cust`

The following describes the variables  in the `health_cust` data.frame.

- `custid`: ID of customer
- `sex`: Sex
- `is_employed`: Employment status
  - `NA`: Unknown or not applicable
  - `TRUE`: Employed
  - `FALSE`: Unemployed
- `income`: Income (in $)
- `marital_status`: Marital status
- `housing_type`: Housing type
- `recent_move`: 
  - `TRUE`: Recently moved
  - `FALSE`: Not recently moved
- `age`: Age
- `state_of_res`: State of residence
- `gas_usage`: Gas usage
  - `NA`: Unknown or not applicable
  - `001`: Included in rent or condo fee
  - `002`: Included in electricity payment
  - `003`: No charge or gas not used
  - `004`-`999`: $4 to $999 (rounded and top-coded)
- `health_ins`: Health insuarance status
  - `TRUE`: customer with health insuarance 
  - `FALSE`: customer without health insuarance 

<br><br>

## Q1a 
Report summary statistics including a mean value 
for each logical variable in `health_cust`.

**Answer:**

```{r}
health_cust_sum <- skim(health_cust)

q1a <- health_cust_sum[health_cust_sum$skim_type == 'logical', 
                       c('skim_variable', 
                         'n_missing', 'complete_rate',
                         'logical.mean', 'logical.count')]
```

<br>





## Q1b 
Use `health_cust` data.frame to create a new data.frame, `health_cust2`, in which 
(1) `age` is greater than 0 and less than 100; and
(2) `income` is positive.

**Answer:**

```{r}


health_cust2 <- health_cust[health_cust$age > 0, ]
health_cust2 <- health_cust2[health_cust2$age < 100, ]
health_cust2 <- health_cust2[health_cust2$income > 0, ]
health_cust2_sum <- skim(health_cust2)

# Combining all conditions into one, which we will cover after the first midterm:
# health_cust2 <- health_cust[ ( 0 < health_cust$age < 100 ) &
#                                 ( health_cust$income > 0 ), ]

```

<br>





## Q1c 
Provide ggplot code and comments to describe 
the distribution of `income`.

**Answer:**

```{r}
ggplot(health_cust2, aes(x = income)) +
  geom_histogram()

ggplot(health_cust2, aes(x = income)) +
  geom_histogram(bins = 100)

ggplot(health_cust2, aes(x = log(income))) +
  geom_histogram(bins = 100)

ggplot(health_cust2, aes(x = log(income))) +
  geom_freqpoly(bins = 100)

```

- Provide some comments

<br>





## Q1d 
Provide ggplot code and comments to describe 
the distribution of `marital_status`.

**Answer:**

```{r}

ggplot(health_cust2,
       aes(y = marital_status,
           fill = marital_status)) +
  geom_bar(show.legend = F)
```

- Provide some comments

<br>





## Q1e 
Suppose you randomly pick one customer's observation 
from each state in data.frame `health_cust`.

Use ggplot code to find:
(1) in which state a customer is most likely to have a health insurance; and
(2) in which state a customer is least likely to have a health insurance.

**Answer:**

```{r, fig.width=12, fig.height=20}
ggplot(health_cust2) +
  geom_bar(aes(y = health_ins,
               group = 1)) +
  facet_wrap(state_of_res ~ . , ncol = 5)

ggplot(health_cust2) +
  geom_bar(aes(y = health_ins, x = stat(prop),
               group = 1)) +
  facet_wrap(state_of_res ~ . , ncol = 5)
```
- The proportional bar chart faceted by `state_of_res` is the one we need.
  - `Alaska` customer is most likely to have health insurance.
  - `Hawaii` customer is least likely to have health insurance.
  - Why? 


The following might mislead...
```{r, fig.width=12, fig.height=20}
ggplot(health_cust2) +
  geom_bar(aes(y = state_of_res, x = stat(prop),
               group = 1)) +
  facet_wrap(health_ins ~ . , ncol = 5)

```

<br>





## Q1f 
Provide ggplot code and comments to describe 
the relationship between `age` and `income`.

**Answer:**

```{r}
ggplot(health_cust2,
       aes(x = age, y = income)) +
  geom_point(alpha = .2, size = .75) +
  geom_smooth() +
  geom_smooth(method = lm, color = 'red')

ggplot(health_cust2,
       aes(x = age, y = income)) +
  geom_hex() +
  geom_smooth() +
  geom_smooth(method = lm, color = 'red')

# Considering income level below $250,000
health_cust3 <- health_cust2[health_cust2$income < 2.5*10^5, ]
ggplot(health_cust3,
       aes(x = age, y = income)) +
  geom_hex() +
  geom_smooth() +
  geom_smooth(method = lm, color = 'red')

```

- Considering the income group whose level is below $250,000, `income` is positively associated with `age` up to `age` around 55. `income` is then negatively associated from age around 55 to 100.

<br>





## Q1g 
Provide ggplot code and comments to describe 
the relationship between `age` and `health_ins`.

**Answer:**

```{r}
ggplot(health_cust2) + 
  geom_histogram(aes(x = age), bins = 75) +
  facet_wrap(health_ins ~.)

ggplot(health_cust2) + 
  geom_freqpoly(aes(x = age), bins = 75) +
  facet_wrap(health_ins ~.)

```

<br>





## Q1h 
Provide ggplot code and comments to describe 
the relationship between `marital_status` and `health_ins`.

**Answer:**

```{r}
ggplot(health_cust2) +
  geom_bar(aes(y = marital_status, 
               fill = health_ins))

ggplot(health_cust2) +
  geom_bar(aes(y = marital_status, 
               fill = health_ins),
           position = "fill")

ggplot(health_cust2) +
  geom_bar(aes(y = marital_status, 
               fill = health_ins),
           position = "dodge")

ggplot(health_cust2) +
  geom_bar(aes(y = marital_status)) +
  facet_wrap(health_ins ~.)

ggplot(health_cust2) +
  geom_bar(aes(y = marital_status, 
               x = stat(prop), group = 1)) +
  facet_wrap(health_ins ~.)

ggplot(health_cust2) +
  geom_bar(aes(y = health_ins, 
               x = stat(prop), group = 1)) +
  facet_grid(marital_status ~.)

```

<br>





## Q1i 
Provide ggplot code and comments to describe 
the relationship between `marital_status` and `housing_type`.

**Answer:**

```{r}
ggplot(health_cust2) +
  geom_bar(aes(y = marital_status, 
               fill = housing_type))

ggplot(health_cust2) +
  geom_bar(aes(y = marital_status, 
               fill = housing_type),
           position = "fill")

ggplot(health_cust2) +
  geom_bar(aes(y = marital_status, 
               fill = housing_type),
           position = "dodge")

ggplot(health_cust2) +
  geom_bar(aes(y = marital_status)) +
  facet_wrap(housing_type ~.)

ggplot(health_cust2) +
  geom_bar(aes(y = marital_status, 
               x = stat(prop), group = 1)) +
  facet_wrap(housing_type ~.)

ggplot(health_cust2) +
  geom_bar(aes(y = housing_type, 
               x = stat(prop), group = 1)) +
  facet_grid(marital_status ~.)

```

<br>



