---
title: "DANL 200: Classwork 3"
date: 2023-09-20
execute: 
  echo: false
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
```

# Question 1. Data Visualization - Facets

## Q1a.
What happens if you facet on a continuous variable?

<br>

## Q1b.
What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?
```{r, echo = T, eval = T}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

<br>

## Q1c.
What plots does the following code make? What does `.` do?
```{r, echo = T, eval = F}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

<br>

## Q1d.
Consider the following faceted plot:
```{r, echo = T, eval = F}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
  - What are the advantages to using faceting instead of the color aesthetic? 
  - What are the disadvantages? 
  - How might the balance change if you had a larger dataset?
  
<br>

## Q1e.
Read `?facet_wrap`. 
  - What does `nrow` do? 
  - What does `ncol` do? 
  - What other options control the layout of the individual panels? 
  - Why doesn’t `facet_grid()` have `nrow` and `ncol` arguments?

<br>

## Q1f.
When using `facet_grid`, you should usually put the variable with more unique levels in the columns. Why?

<br>

## Q1g.
Use the following data.frame.

```{r, eval = F, echo = T}
tvshows_web <- read_csv(
		'https://bcdanl.github.io/data/tvshows.csv')
```

- Describe the relationship between audience size (`GRP`) and audience engagement (`PE`) using `ggplot`. 

- Explain the relationship in words. 


<br>

