---
title: "DANL 200: Classwork 2"
date: 2023-09-08
execute: 
  echo: false
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
```

## Question 1. Data Visualization - First Steps

Load the following data.frame, `mpg`.
```{r}
#| echo: true
#| warning: false
#| message: false

library(tidyverse)
mpg <- ggplot2::mpg
```


```{r}
#| eval: true

paged_table(mpg, 
            options = list(rows.print = 10,
                           cols.print = 3) )
```

```{r}
#| echo: true
#| warning: false
#| message: false

library(skimr)
skim(mpg)
```



### Q1a. 
Run `ggplot(data = mpg)`. What do you see?
```{r}
ggplot(data = mpg)
```

<br>

### Q1b.
How many rows are in `mpg`? How many columns?
```{r}
nrow(mpg)
ncol(mpg)
```

<br>


### Q1c.
Make a scatterplot of `hwy` vs. `cyl`.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

<br>

### Q1d.
What happens if you make a scatterplot of `class` vs. `drv`? Why is the plot not useful?


```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

<br>

### Q1e.
What’s gone wrong with this code? Why are the points not blue?

```{r}
#| echo: true
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

<br>

### Q1f.
Which variables in `mpg` are categorical? Which variables are continuous?

<br>

### Q1g.
Map a continuous variable to `color`, `size`, and `shape`. How do these aesthetics behave differently for categorical vs. continuous variables?

<br>

### Q1h.
What happens if you map an aesthetic to something other than a variable name, like `aes(color = displ < 5)`? 


