---
title: "DANL 200: Homework Assignment 2"
date: "`r format(Sys.time(), '%B %d, %Y')`"
execute: 
  echo: false
  eval: false

---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
```

# Question 1. Working with Customer Data with Health Insurance

Read the data file, `custdata_rev.csv`, as the data.frame object with the name, `health_cust`:

```{r}
#| echo: true
library(tidyverse)
library(skimr)

health_cust <- read_csv(
  'https://bcdanl.github.io/data/custdata_rev.csv'
)
```



## Variable Description for `health_cust`

The following describes the variables  in the `health_cust` data.frame.

- `custid`: ID of customer
- `sex`: Sex
- `is_employed`: Employment status
  - `NA`: Unknown or not applicable
  - `TRUE`: Employed
  - `FALSE`: Unemployed
- `income`: Income (in $)
- `marital_status`: Marital status
- `housing_type`: Housing type
- `recent_move`: 
  - `TRUE`: Recently moved
  - `FALSE`: Not recently moved
- `age`: Age
- `state_of_res`: State of residence
- `gas_usage`: Gas usage
  - `NA`: Unknown or not applicable
  - `001`: Included in rent or condo fee
  - `002`: Included in electricity payment
  - `003`: No charge or gas not used
  - `004`-`999`: $4 to $999 (rounded and top-coded)
- `health_ins`: Health insuarance status
  - `TRUE`: customer with health insuarance 
  - `FALSE`: customer without health insuarance 

<br><br>

## Q1a 
Report summary statistics including a mean value 
for each logical variable in `health_cust`.

<br>





## Q1b 
Use `health_cust` data.frame to create a new data.frame, `health_cust2`, in which 
(1) `age` is greater than 0 and less than 100; and
(2) `income` is positive.

<br>





## Q1c 
Provide ggplot code and comments to describe 
the distribution of `income`.

<br>





## Q1d 
Provide ggplot code and comments to describe 
the distribution of `marital_status`.

<br>





## Q1e 
Suppose you randomly pick one customer's observation 
from each state in data.frame `health_cust`.

Use ggplot code to find:
(1) in which state a customer is most likely to have a health insurance; and
(2) in which state a customer is least likely to have a health insurance.

<br>





## Q1f 
Provide ggplot code and comments to describe 
the relationship between `age` and `income`.

<br>





## Q1g 
Provide ggplot code and comments to describe 
the relationship between `age` and `health_ins`.

<br>





## Q1h 
Provide ggplot code and comments to describe 
the relationship between `marital_status` and `health_ins`.

<br>





## Q1i 
Provide ggplot code and comments to describe 
the relationship between `marital_status` and `housing_type`.

<br>



