---
title: "DANL 200: Homework Assignment 1"
date: "`r format(Sys.time(), '%B %d, %Y')`"
execute: 
  echo: false
  eval: false

---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
```

# Question 1. Data Analytics Survey

- Complete the following survey:
  - [https://forms.gle/Ya2HbPgHynkAp5e28](https://forms.gle/Ya2HbPgHynkAp5e28)
  

<br><br><br>

# Question 2. Working with `data.frame`

## Q2a.
- Download the CSV file, `starwars.csv`, from the Brightspace. 

- Read the data file, `starwars.csv`, as the data.frame object with the name, `starwars`, using (1) the `read_csv()` function and (2) the absolute path name of the file `starwars.csv` from your local hard disk drive in your laptop.

```{r}
#| echo: true
#| warning: false
#| message: false

library(tidyverse)
path_starwars <- 'PATHNAME_OF_THE_CSV_FILE, starwars.csv'
starwars <- read_csv(path_starwars)
```

<br>

## Variable Description for `starwars` data.frame

The following describes the variables in the `starwars` data.frame.


- `films`
List of films the character appeared in

- `name`
Name of the character

- `species`
Name of species

- `height`
Height (cm)

- `mass`
Weight (kg)

- `hair_color`, `skin_color`, `eye_color`
Hair, skin, and eye colors

- `birth_year`
Year born (BBY = Before Battle of Yavin)

- `sex`
The biological sex of the character, namely male, female, hermaphroditic, or none (as in the case for Droids).

- `gender`
The gender role or gender identity of the character as determined by their personality or the way they were programmed (as in the case for Droids).

- `homeworld`
Name of homeworld


<br>

## Q2b.
- Provide the R code to find how many characters are `Droid` `species`.

```{r}
library(skimr)

# The following gives not only Droid but also NA
# q2a <- starwars[starwars$species == 'Droid', ]

droid <- which(starwars$species == 'Droid')
q2a <- starwars[droid,]

nrow(
  q2a
)
```

<br>

## Q2c.
- Provide the R code to find which `Droid` character did appear most in the Star Wars `films`.

```{r}
q2b <- q2a[ , c('name','films')]
```


<br>

## Q2d.
- Provide the R code to find the minimum, first quartile, median, third quartile, and the maximum of `height`.

- [Statistics] Explain why the median value is different from the mean value.
  - When is the median value greater than the mean value?
  - When is the median value less than the mean value?

```{r}
skim(starwars$height)
```

<br>

## Q2e.
- Provide the R code to find the number of unique/distinct `homeworld` values.


```{r}
skim(starwars$homeworld)
```


<br><br><br>

# Question 3. Data Visualization

## Q3a.
Provide the R code to install the R package, `gapminder`.

```{r}
install.packages("gapminder")
```

<br>

## Q3b.
Load the `gapminder` data.frame provided by the R package, `gapminder`.


```{r}
library(gapminder)
gapminder
?gapminder
```


<br>


## Variable Description for `gapminder` data.frame

- The following describe the variables in the `gapminder` data.frame.

  - `country`
  factor with 142 levels
  
  - `continent`
  factor with 5 levels
  
  - `year`
  ranges from 1952 to 2007 in increments of 5 years
  
  - `lifeExp`
  life expectancy at birth, in years
  
  - `pop`
  population
  
  - `gdpPercap`
  GDP per capita (US$, inflation-adjusted)

<br>

## Q3c.
Provide ggplot code and comments to describe the relationship between `gdpPercap` and `lifeExp`.

<br>

## Q3d.
Provide ggplot code and comments to describe how the relationship between `gdpPercap` and `lifeExp` varies by `continent`.

<br>
