---
title: "DANL 200: Classwork 1"
date: 2023-09-06
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    # highlight-style: espresso
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
```

## Question 1. Working with `data.frame`

Load the following data.frame, `tvshows`.
```{r}
#| echo: true
#| warning: false
#| message: false

library(tidyverse)
tvshows <- read_csv(
		'https://bcdanl.github.io/data/tvshows.csv')
```


```{r}
#| eval: true
#| echo: false
paged_table(tvshows, 
            options = list(rows.print = 10,
                           cols.print = 3) )
```

### Q1a. 
Provide the R code to return the shows whose `Duration` values are `30`.

```{r}
q1a <- tvshows[tvshows$Duration == 30, ]
```

```{r}
q1a_indices <- which(tvshows$Duration == 30)
q1a <- tvshows[ q1a_indices, ]
```

```{r}
#| eval: true
#| echo: false
paged_table(q1a, 
            options = list(rows.print = 10,
                           cols.print = 3) )
```

The number of shows whose `Duration` is 30 is `r nrow(q1a)`.

<br>

### Q1b.
Provide the R code to return the shows whose `GRP` values are greater than the mean value of `GRP` in all the shows in the `tvshows` data.frame.


Hint: We can calculate the mean value of numeric vector `x` using the `mean(x)` function.
```{r}
q1b <- tvshows[tvshows$GRP > mean(tvshows$GRP), ]
```

The mean value of `GRP` is `r mean(tvshows$GRP)`.

```{r}
#| eval: true
#| echo: false
paged_table(q1b, 
            options = list(rows.print = 10,
                           cols.print = 3) )
```


<br>

### Q1c.
Provide the R code to return the data.frame with only three variables---`Show`, `PE`, and `GRP`---for which `PE` values are greater than the mean value of `PE`.

```{r}
q1c <- tvshows[tvshows$PE > mean(tvshows$PE), ]
q1c <- q1c[ , c("Show", "PE", "GRP")]
```

```{r}
#| eval: true
#| echo: false
paged_table(q1c, 
            options = list(rows.print = 10,
                           cols.print = 3) )
```

