---
title: "Fall 2023 DANL 200 Midterm Exam 2"
# subtitle: "Version 1"
author: "Byeong-Hak Choe"
institute: "SUNY Geneseo"
date: "2023-11-17"
format: 
  html
    # toc: true
    # toc-depth: 2
    # toc-expand: true
    # toc-title: Contents

execute:
  echo: false
  eval: true
  message: false
  warning: false
---
```{r, include=F}
library(tidyverse)
library(skimr)
library(lubridate)
library(hrbrthemes)

knitr::opts_chunk$set(fig.width=9, fig.height=6)  

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5) ),
                axis.title.y = element_text(size = rel(1.5),
                                            angle = 0))
          )

```
# Honor Pledges

I solemnly swear that I will not cheat or engage in any form of academic dishonesty during this exam. 

I will not communicate with other students or use unauthorized materials. 

I will uphold the integrity of this exam and demonstrate my own knowledge and abilities.

By taking this pledge, I acknowledge that academic dishonesty undermines the academic process and is a violation of the trust placed in me as a student. 

I accept the consequences of any violation of this promise.

- Student's Name: 
- Student's ID: 
- Student's Signature: 



# Load R packages
- Here we are loading all the R packages we need for the Midterm Exam II, so that you do not need to load the R packages in your code.
```{r, echo = T}
library(tidyverse)
library(skimr)
```


<br><br>

# Question 1



- The following is the data frames for Question 1.

```{r, echo = T}
games <- read_csv('https://bcdanl.github.io/data/games.csv')
```

## Variable description

- **gamename**:	Name of games
- **date**:	Date of measure
- **year**:	Year of measure
- **month**:	Month of measure
- **avg**:	Average number of players at the same time
- **peak**:	Highest number of players at the same time



<br>


- The following is the data.frame, **games**.
```{r, results = 'asis'}
rmarkdown::paged_table(games,
                       options = list(rows.print = 20,
                                      cols.print = 6,
                                      pages.print = 0,
                                      paged.print = F
                                      )) 
```

<br>

```{r, echo = T}
nrow(games)
ncol(games)
```

<br>

```{r, echo = T}
unique(games$gamename)
```

<br>


## Q1a
- Provide R code to recreate the ggplot figure illustrating the monthly trend of the average for the top 5 **gamename**s. 
  - The top 5 **gamename**s should be determined by the highest mean values of **avg** for each **gamename**.

<br>

- **Answer**:

```{r}
#| eval: false
Q1a <- close_elections[ , c('score', 'democrat') ]

```

<br><br><br><br><br><br><br><br><br><br><br><br>

