---
title: "ECON 340: Homework Assignment 3"
date: 2023-11-08
execute: 
  echo: false
  eval: true
  warning: false
  message: false
  fig-width: 9
  fig-height: 6
---
```{r}
#| include: false

library(tidyverse)
library(hrbrthemes)
```


# Question 1. Provision of the Sustailibilities
The town of Geneseo has two residents: Ben and Tiana. The town currently funds its sustainability department solely from the individual contributions of Ben and Tiana. 

Ben and Tiana have their own utility functions, $U_{B}$ and $U_{T}$ respectively, over private goods ( $X$ ) and total sustainability workers ( $M$ ) of the forms:

$$U_{B} = 3 × \log(X_{B}) + 3 × \log(M).$$ 
$$U_{T} = 4 × \log(X_{T}) + 2 × \log(M).$$


The total provision of sustainability workers hired, $M$, is the sum of the number hired by each of the two persons: 

$$M = M_{B} + M_{T}.$$ 

Ben and Tiana both have income of $\$100$, and the price of both the private good and the sustainability worker is $\$1$. Therefore, they are limited to providing between 0 and 100 sustainability workers.

<br>

## Q1a.
- How many sustainability workers are hired if the Geneseo government does not intervene? 
  - How many are paid for by Ben? By Tiana?
  - Describe the Ben's and the Tiana's decisions on provision of sustainability workers using a graph in the $(M_{B}, M_{T})$ space.

<br>

Taking the Tiana's action and the prices of the private good and the sustainability worker as given, Ben maximizes his utility:

$$
\begin{align}
\max_{M_{B}}\; U_{B} &= 3 × \log(100 -  M_{B}) + 3 × \log(M_{B} + M_{T}),\tag{UMax-B}
\end{align}
$$
where $X_{B} = 100 - M_{B}$.

The first-order condition for problem (Umax-B) is:

$$
\frac{d U_{B}}{dM_{B}} = -\frac{3}{100-M_{B}} + \frac{3}{M_{B} + M_{T}} = 0.\tag{FOC-B}
$$
(FOC-B) simplifies to

$$
M_{B} = 50 - \frac{1}{2}M_{T}.\tag{Action-B}
$$

Taking the Ben's action and the prices of the private good and the sustainability worker as given, Tiana maximizes his utility:

$$
\begin{align}
\max_{M_{T}}\; U_{T} &= 4 × \log(100 -  M_{T}) + 2 × \log(M_{B} + M_{T}),\tag{UMax-T}
\end{align}
$$
where $X_{T} = 100 - M_{T}$.

The first-order condition for problem (Umax-T) is:

$$
\frac{d U_{T}}{dM_{T}} = -\frac{4}{100-M_{T}} + \frac{2}{M_{T} + M_{T}} = 0.\tag{FOC-T}
$$
(FOC-T) simplifies to

$$
M_{T} = \frac{1}{3}(100 - 2M_{B}).\tag{Action-T}
$$

Both Ben's and Tiana's optimal decisions on how many sustainability workers to hire are characterized by (Action-B) and (Action-T), respectively. Solving the two unknowns---$M_{B}$ and $M_{T}$---in the two equations---(Action-B) and (Action-T)---gives us:

$$
(M_{B}^{*}, M_{T}^{*}) = (50, 0).
$$
Therefore, the total provision of sustainability workers under no government intervention is $M^{*} = M_{B}^{*} + M_{T}^{*} = 50$.
<br>

The following describes Ben's and Tiana's reactions to the other individual's choice on the number of sustainability workers to hire:

```{r}

df <- data.frame(m_i = seq(0, 150, 5),
                 Ben = 50 - seq(0, 150, 5),
                 Tiana = 50 - 1.5*seq(0, 150, 5)
)

df <- df %>%
  rename(`Ben's Reaction to Tiana` = Ben,
         `Tiana's Reaction to Ben` = Tiana) %>% 
  pivot_longer(`Ben's Reaction to Tiana`:`Tiana's Reaction to Ben`,
               names_to = "Reaction",
               values_to = "fn")


ggplot(df, aes(x = m_i, y = fn )) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  geom_line(aes(color = Reaction), size = 1.5) +
  geom_point(data = data.frame(x = 0, y = 50),
             aes(x = x, y = y), size = 2.5) +
  labs(x = "Number of Tiana's Sustainability Workers",
       y = "Number of Ben's Sustainability Workers",
  ) +
  scale_x_continuous(breaks = seq(0,150,5)) +
  scale_y_continuous(breaks = seq(0,150,5)) +
  coord_cartesian(xlim = c(0,60), 
                  ylim = c(0,60)) +
  theme_ipsum() +
  theme(axis.title.x = element_text(size = rel(1.25)),
        axis.title.y = element_text(size = rel(1.25)))
```



## Q1b.
- What is the socially optimal number of sustainability workers if the Social planner appreciates each individual's budget constraint? 

<br>

Taking both Ben's and Tiana's budget constraints as given, the Social Planner chooses $M_{B}$ and $M_{T}$ to maximize the social welfare $SW = U_{B} + U_{T}$:
$$
\begin{align}
\max_{M_{B}, M_{T}}\; SW &= 3 × \log(100 -  M_{B}) + 4 × \log(100 -  M_{T}) + 5 × \log(M_{B} + M_{T}),\tag{SWMax}
\end{align}
$$

First-order conditions for problem (SWMax) are:

$$
\begin{align}
\frac{\partial SW}{\partial M_{B}} = -\frac{3}{100-M_{B}} + \frac{5}{M_{B}+M_{T}} = 0; \tag{FOC-SW-B}\\
\frac{\partial SW}{\partial M_{B}} = -\frac{4}{100-M_{T}} + \frac{5}{M_{B}+M_{T}} = 0. \tag{FOC-SW-T}
\end{align}
$$
(FOC-SW-B) and (FOC-SW-T) simplify to

$$
\begin{align}
M_{B} = \frac{1}{8}(500 - 3M_{T});\tag{Action-SW-B}\\
M_{T} = \frac{1}{9}(100 - 4M_{B}).\tag{Action-SW-T}
\end{align}
$$

The Social Planner's allocation of the number of Ben's and Tiana's sustainability workers is characterized by (Action-SW-B) and (Action-SW-T). Solving the two unknowns---$M_{B}$ and $M_{T}$---in the two equations---(Action-SW-B) and (Action-SW-T)---gives us:

$$
(M_{B}^{o}, M_{T}^{o}) = \left(50, \frac{100}{3}\right).
$$
Therefore, the socially optimal level of sustainability workers is $M^{o} = M_{B}^{o} + M_{T}^{o} = \frac{250}{3}$, which is greater than the one in Q1a. The sustainability workers are underprovided in the private market.


<br>

The following plots (Action-SW-B) and (Action-SW-T), describing how the Social Planner allocates Ben's and Tiana's sustainability workers:

```{r}

df <- data.frame(m_i = seq(0, 180, 5),
                 Ben = (1/8)*( 500 - 3*seq(0, 180, 5) ),
                 Tiana = (1/4)*( 500 - 9*seq(0, 180, 5) )
)

df <- df %>%
  rename(`Ben's Reaction to Tiana` = Ben,
         `Tiana's Reaction to Ben` = Tiana) %>% 
  pivot_longer(`Ben's Reaction to Tiana`:`Tiana's Reaction to Ben`,
               names_to = "Reaction",
               values_to = "fn")


ggplot(df, aes(x = m_i, y = fn )) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  geom_line(aes(color = Reaction), size = 1.5) +
  geom_point(data = data.frame(x = 100/3, y = 50),
             aes(x = x, y = y), size = 2.5) +
  geom_segment(data = data.frame(x1 = 100/3, y1 = 50,
                                 x2 = 0, y2 = 50),
               aes(x = x1, y = y1, xend = x2, yend = y2),
               linetype = 3) +
  geom_segment(data = data.frame(x1 = 100/3, y1 = 50,
                                 x2 = 100/3, y2 = 0),
               aes(x = x1, y = y1, xend = x2, yend = y2),
               linetype = 3) +
  labs(x = "Number of Tiana's Sustainability Workers",
       y = "Number of Ben's Sustainability Workers",
  ) +
  scale_x_continuous(breaks = seq(0,180,10)) +
  scale_y_continuous(breaks = seq(0,180,10)) +
  coord_cartesian(xlim = c(0,180), 
                  ylim = c(0,130)) +
  theme_ipsum() +
  theme(axis.title.x = element_text(size = rel(1.25)),
        axis.title.y = element_text(size = rel(1.25)))
```

- The allocation proposed by the Social Planner serves as a benchmark for environmental protection that the Geneseo government aims to implement.




## Q1c.
- If your answer in Q1b differs from Q1a, explain economically and intuitively why they are different.

<br>

- The socially optimal level of provision of sustainability workers is larger than the one without the Geneseo government's intervention. 

- Notice that Ben's contribution to sustainability goes down as Tiana's contribution goes up, and vice versa. This implies that both individuals have incentive to free ride on each other's.

- Intuitively, in the computation in Q1a, we set the marginal utility of the last sustainability worker to each resident equal to the marginal utility of consumption for that resident. In Q1b, we set the sum of the marginal utilities of the last sustainability worker---the social marginal utility of the sustainability worker---equal to the marginal utility of consumption for either resident. Since the social marginal utility of sustainability workers exceeds the individual marginal utilities of that sustainability workers, Geneseo optimally hires more than individuals would if they were acting alone.

<br><br><br>

# Question 2. Calculating the Value of Statistical Life

As an economist, you have been doing a project trying to estimate the willingness-to-accept risk on the part of workers. You have gathered data on a variety of blue-collar occupations, ranging from clerical work to truck driving to coal mining. Controlling for age, gender, education, work experience and race, you have uncovered the following relationship:

```{r}
df <- data.frame(
  risk = 0:5,
  extra_wage = seq(0,2500, 500)
)

```

```{r}
#| echo: false
#| eval: true

ggplot(df, aes(x = risk, y = extra_wage)) +
  geom_line() +
  geom_point() +
  labs(x = "Mortality Risk (out of 10,000)",
       y = "Extra wage per year") + 
  hrbrthemes::scale_y_comma(labels = scales::dollar) +
  hrbrthemes::theme_ipsum() +
  theme(axis.title.x = element_text(angle = 0,
                                    size = rel(1.33)),
        axis.title.y = element_text(angle = 0,
                                    size = rel(1.33)),
        )
  

```

## Q2a.	
- What is your estimate of the “value of a statistical life” (VSL)? 
  - What does this VSL mean?

<br>

- **Answer**: If we draw a line through the points, we notice that for each $\frac{1}{10,000}$ increase in mortality risk, the wage premium rises by about $500. Thus, 10,000 workers trade $500 each for a total of $5 million per life lost: the value of a “statistical life”.

## Q2b.	
- Why might this VSL value be a poor estimate of the value placed on lives saved from pollution reduction? (Three reasons)

<br>

-	**Three reasons**:
    1.	This is WTA for voluntary risk; environmental risk can be involuntary.
    2.	Sample selection bias: are the individuals surveyed (e.g., police officers, fire fighters) more “risk lovers” than the average individual?
    3.	Rationality: do the individuals really know the risk they are shouldering in exchange for the wage premium? 



<br><br><br>
