---
title: "ECON 340: Homework Assignment 3"
date: 2023-11-04
execute: 
  echo: false
---

# Question 1. Provision of the Sustailibilities
The town of Geneseo has two residents: Ben and Tiana. The town currently funds its sustainability department solely from the individual contributions of Ben and Tiana. 

Ben and Tiana have their own utility functions, $U_{B}$ and $U_{T}$ respectively, over private goods ( $X$ ) and total sustainability workers ( $M$ ) of the forms:

$$U_{B} = 3 × \log(X_{B}) + 3 × \log(M).$$ 
$$U_{T} = 4 × \log(X_{T}) + 2 × \log(M).$$ 

The total provision of sustainability workers hired, $M$, is the sum of the number hired by each of the two persons: 

$$M = M_{B} + M_{T}.$$ 

Ben and Tiana both have income of $\$100$, and the price of both the private good and the sustainability worker is $\$1$. Therefore, they are limited to providing between 0 and 100 sustainability workers.

<br>

## Q1a.
- How many sustainability workers are hired if the Geneseo government does not intervene? 
  - How many are paid for by Ben? By Tiana?
  - Describe the Ben's and the Tiana's decisions on provision of sustainability workers using a graph in the $(M_{B}, M_{T})$ space.

<br>

## Q1b.
- What is the socially optimal number of sustainability workers if the Social planner appreciates each individual's budget constraint? 

<br>

## Q1c.
- If your answer in Q1b differs from Q1a, explain economically and intuitively why they are different.

<br><br><br>

# Question 2. Calculating the Value of Statistical Life

As an economist, you have been doing a project trying to estimate the willingness-to-accept risk on the part of workers. You have gathered data on a variety of blue-collar occupations, ranging from clerical work to truck driving to coal mining. Controlling for age, gender, education, work experience and race, you have uncovered the following relationship:

```{r}
#| include: false
#| echo: false
library(tidyverse)
df <- data.frame(
  risk = 0:5,
  extra_wage = seq(0,2500, 500)
)

```

```{r}
#| echo: false
#| eval: true

ggplot(df, aes(x = risk, y = extra_wage)) +
  geom_line() +
  geom_point() +
  labs(x = "Mortality Risk (out of 10,000)",
       y = "Extra wage per year") + 
  hrbrthemes::scale_y_comma(labels = scales::dollar) +
  hrbrthemes::theme_ipsum() +
  theme(axis.title.x = element_text(angle = 0,
                                    size = rel(1.33)),
        axis.title.y = element_text(angle = 0,
                                    size = rel(1.33)),
        )
  

```

## Q2a.	
- What is your estimate of the “value of a statistical life” (VSL)? 
  - What does this VSL mean?

<br>

## Q2b.	
- Why might this VSL value be a poor estimate of the value placed on lives saved from pollution reduction? (Three reasons)

<br><br><br>
