---
title: "DANL 200, Homework Assignment 1"
subtitle: "Data Visualization with `ggplot`<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "September 17, 2022"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---


```{r, include = F}
# install.packages("ggridges")
# install.packages("colorspace")
# install.packages("ggthemes")
library(tidyverse)
library(lubridate)
library(ggridges)
library(ggthemes)
theme_set(theme_bw())
library(colorspace)
```

# Instruction
The starting R script for Homework Assignment 1 is provided in the Assignment section in our Canvas.
 
Submit your R script with the following file name with no spaces---`DANL210_hw1_LASTNAME_FIRSTNAME.R`---to Homework Assignment 1 in the Assignment section in our Canvas.

The due for the Homework Assignment 1 is September 26, 2022, 11:59 PM Eastern Time.


# Question 1

## (a)

Download the data file, `bikeshare_dc.csv`, from the Files section in our Canvas web-page. Read the data file, `bikeshare_dc.csv`, as the *data.frame* object with the name, `bikeshare`, using (1) the `read_csv()` function and (2) the *absolute path* name of the file `titanic_all.csv` from your local hard disk drive in your laptop.

```{r, echo = T, warning=F, message=F}
library(tidyverse)
bikeshare <- read_csv('/Users/byeong-hakchoe/Google Drive/suny-geneseo/teaching-materials/lecture-data/bikeshare_dc.csv')

```

