---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Lecture 15 Class Exercises<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```

### Loading R packages for Class Exercises

```{r, eval = F, echo = T}
library(tidyverse)
```

<br><br>


# Tweets with #climatechange
The following data is for Class Exercises:

```{r, echo = T, eval = F}
NY_CC_tweets <- read_csv(
  'https://bcdanl.github.io/data/NY_tweets_cc_no_contents.csv')
```

## Variable Description

- `id_user`: a unique identification number for a Twitter user whom retweeted to a tweet with *#climatechange* or *#globalwarming*.
- `id_city`: a unique identification number for a city.
- `FIPS`: a unique identification number for a county.

Each row represents an observation of a retweet to a tweet with *#climatechange* or *#globalwarming*.

Each row includes a Twitter user's geographic information at city or county levels (variables `FIPS`, `county`, `city`) as well as information about timing when a Twitter user retweeted (variables `year`, `month`, `day`, `hour`, `minute`, `second`). 


## Q1a
How many Twitter users retweeted on the date, January 1, 2017?

<br><br>

## Q1b
Which city is with the third highest number of retweets on the date, December 1, 2017?

<br><br>

## Q1c
For each year, find the top 5 Twitter users in NY state in terms of the number of retweets they made in NY state. In which city and county do these users live in?

<br><br>

## Q1d
Summarize the data set into the data frame with county and month levels of retweets with the following variables:

- `FIPS`, `county`, `year`, `month`;
- `n_retweets`: the number of retweets in `year` *YYYY* and `month` *MM* from `county` `C`.

The `unique()` or `distinct()` functions can be used to keep only unique/distinct rows from a data frame.

<br><br>


## Q1e
Describe the relationship between the number of retweets and `county` using ggplot. Make a simple comment on your plot.
