---
title: "Spring 2023, DANL 200: Introduction to Data Analytics"
subtitle: "Midterm Exam"
author: "Byeong-Hak Choe"
---

```{r setup, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggthemes)
library(ggtech)
library(ggthemr)
library(hrbrthemes)
library(hexbin)
library(ggforce)
library(skimr)

knitr::opts_chunk$set(fig.width=8, fig.height=5,
                      echo = T, eval = T, message = F, warning = F)  

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray")))
```

### Loading R packages

```{r, eval = F, echo = T}
library(tidyverse)
```

<br><br>

# Question 1

Load the data.frame for Question 1.

```{r}
fao_stat <- read_csv('https://bcdanl.github.io/data/fao_stat.csv')
```

```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(fao_stat)
```

<br>

## Variable Description

-   `Area`:
    -   Country
-   `Year`:
    -   Year
-   `SSA`:
    -   `TRUE` if `Area` belongs to Sub-Saharan Africa;
    -   `FALSE` otherwise
-   `gdp_per_capita`:
    -   Gross domestic product per capita, PPP, (constant 2017 international \$)
-   `drinking_water`:
    -   Percentage of population using at least basic drinking water services (percent)
-   `sanitation_service`:
    -   Percentage of population using at least basic sanitation services (percent)
-   `children_stunted`:
    -   Percentage of children under 5 years of age who are stunted (percent)
    -   What does it mean stunted?
        -   Stunting is the impaired growth and development that children experience from poor nutrition, repeated infection, and inadequate psychosocial stimulation.
-   `children_overweight`:
    -   Percentage of children under 5 years of age who are overweight (percent)
-   `investment_pct`:
    -   Investment share of gross domestic product (GDP), (percent)
    -   `investment_pct` = 100 * I/GDP, where GDP = (Consumption) + (Investment) + (Government Expenditure) + (Net Export)
    -   It measures how much of the new value added in the economy is invested.

<br>

## Q1a.

-   For each `SSA` value, calculate the mean value of each numeric variable.

-   For each numeric variable, how much is the difference between Sub-Saharan Africa and non-Sub-Saharan Africa in their mean values?


<br>

## Q1b

Provide both (1) ggplot code and (2) a simple comment to describe how the relationship between `investment_pct` and `children_stunted` varies by `SSA`.


<br>

## Q1c

Provide both (1) ggplot code and (2) a simple comment to describe how the relationship between `investment_pct` and `children_stunted` varies by `SSA` and `Year`.

<br>


## Q1d.

Provide both (1) ggplot code and (2) a simple comment to describe how overall the yearly trend of each `Area`'s `children_stunted` varies by `SSA`.

<br><br>

# Question 2

Load the data.frame for Question 2.

```{r}
restaurant <- read_csv('https://bcdanl.github.io/data/DOHMH_NYC_Restaurant_Inspection.csv')
```



```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(restaurant)
```


## Variable Description

-   `CAMIS`:
    -   This is an unique identifier for the entity (restaurant);
    -   10-digit integer
-   `DBA`:
    -   This field represents the name (doing business as) of the entity (restaurant);
    -   Public business name, may change at discretion of restaurant owner
-   `BORO`:
    -   Borough in which the entity (restaurant) is located.;
    -   • `1` = MANHATTAN
    -   • `2` = BRONX
    -   • `3` = BROOKLYN
    -   • `4` = QUEENS
    -   • `5` = STATEN ISLAND
    -   • `0` = Missing;
-   `CUISINE DESCRIPTION`:
    -   This field describes the entity (restaurant) cuisine.
-   `ACTION`:
    -   This field represents the actions that is associated with each restaurant inspection. ;
    -   • Violations were cited in the following area(s).
    -   • No violations were recorded at the time of this inspection.
    -   • Establishment re-opened by DOHMH
    -   • Establishment re-closed by DOHMH
    -   • Establishment Closed by DOHMH.
    -   • Violations were cited in the following area(s) and those requiring immediate action were addressed.
-   `VIOLATION CODE`:
    -   Violation code associated with an establishment (restaurant) inspection
-   `VIOLATION DESCRIPTION`:
    -   Violation description associated with an establishment (restaurant) inspection
-   `CRITICAL FLAG`:
    -   Indicator of critical violation;
    -   • `Critical`
    -   • `Not Critical`
    -   • `Not Applicable`;
    -   Critical violations are those most likely to contribute to food-borne illness
-   `SCORE`:
    -   Total score for a particular inspection;
-   `GRADE`:
    -   Grade associated with the inspection;
    -   • `N` = Not Yet Graded
    -   • `A` = Grade A
    -   • `B` = Grade B
    -   • `C` = Grade C
    -   • `Z` = Grade Pending
    -   • `P` = Grade Pending issued on re-opening following an initial inspection that resulted in a closure

<br>


## Q2a.

What are the mean, standard deviation, first quartile, median, third quartile, and maximum of `SCORE` for each `GRADE` of restaurants?

<br>


## Q2b.

-   How many restaurants with a `GRADE` of `A` are there in NYC?

-   How much percentage of restaurants in NYC are a `GRADE` of `C`?




<br>

## Q2c.

Provide both (1) ggplot code and (2) a simple comment to describe how the distribution of `SCORE` varies by `GRADE` and `CRITICAL FLAG`.

<br>


## Q2d.

Provide both (1) ggplot code and (2) a simple comment to describe how the probability distribution of `CRITICAL FLAG` varies by `GRADE` and `BORO`.


<br>


## Q2e.

For the 10 most common `CUISINE DESCRIPTION` values, find the `CUISINE DESCRIPTION` value that has the highest proportion of `GRADE` `A`.

<br>


## Q2f.

-   Find the 3 most common names of restaurants (`DBA`) in each `BORO`.
    -   If the third most common `DBA` values are multiple, please include all the `DBA` values.
-   Overall, which `DBA` value is most common in NYC?


<br>


## Q2g.

For all the `DBA` values that appear in the result of Q2f, find the `DBA` value that is most likely to commit critical violation.

