---
title: "DANL 310: Lecture 05. Make a Plot"
author: "Byeong-Hak Choe"
date: February 10, 2022
output: 
  ioslides_presentation:
    transition: 0
    widescreen: TRUE
    smaller: FALSE
    smart: TRUE
---

## Loading the R packages and setting the theme for ggplot {.build .faster}

```{r setup}
library(tidyverse)
library(skimr)   # a better summary of data
theme_set(theme_minimal()) # setting the minimal theme for ggplot
```

- Use `theme()` if you want to tweak the display of an existing theme.  

- The following lists the ggplot themes:  

theme           | theme
----------------|----------------
theme_grey()    | theme_gray()
theme_bw()      | theme_linedraw()
theme_light()   | theme_dark()
theme_minimal() | theme_classic()
theme_void()    | theme_test()


## gapminder data {.build .faster}

- The `gapminder` package include the `gapminder` data frame.

```{r}
library(gapminder)
gapminder
```

```{r}
head(gapminder)
```

## Descriptive statistics {.build .faster}
```{r skim, include=FALSE}
skimmed <- skim(gapminder)
skimmed <- skimmed %>% 
  select(-factor.ordered, -complete_rate)
skimmed <- as_tibble(skimmed)
skimmed_chr <- skimmed[1:2,2:5]
skimmed_numeric <- skimmed[3:6,c(2:3,5:13)]
skimmed_numeric[,4:10] <- round(skimmed_numeric[,4:10], 2)
skimmed_numeric <- skimmed_numeric %>% 
  select(-factor.top_counts)
```

```{r skim_chr, eval=TRUE, echo=TRUE, fig.align='center', out.height="75%"}
knitr::kable(skimmed_chr, "simple",   # from skimr::skim() 
             col.names = c('Variable', 'n_missing', 'n_Categories', 'Counts'))  
```

## Descriptive statistics {.build .faster}

```{r skim_n, eval=TRUE, echo=TRUE, fig.align='center', out.height="75%"}
knitr::kable(skimmed_numeric, "simple",   # from skimr::skim()
             col.names = c('Variable', 'n_missing', 'Mean', 'SD',
                           'Min', 'Q1', 'Median','Q3', 'Max', 'Hist'))
```

## Mappings link data to things you see {.build .faster}

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, results="asis"}
p <- ggplot(data = gapminder)

p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p

p + geom_point() 

```

## Mappings link data to things you see {.build .faster}

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=4, results="asis"}
p <- ggplot(data = gapminder)

p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p
```

## Mappings link data to things you see {.build .faster}

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, results="asis"}
p <- ggplot(data = gapminder)

p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p
```


## Mappings link data to things you see {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, results="asis"}
p + geom_point() 

```

## Build your plots layer by layer {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, results="asis"}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y=lifeExp))
p + geom_smooth()
```
## Build your plots layer by layer {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, results="asis"}
p + geom_point() + geom_smooth() 
```


## Build your plots layer by layer {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, results="asis"}
p + geom_point() + geom_smooth(method = "lm") 
```


## Build your plots layer by layer {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, results="asis"}
p + geom_point() +
    geom_smooth(method = "gam") +
    scale_x_log10()
```


## Build your plots layer by layer {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, results="asis"}
p + geom_point() +
    geom_smooth(method = "gam") +
    scale_x_log10(labels = scales::dollar)
```


## Build your plots layer by layer {.build .faster}

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, results="asis"}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y=lifeExp))
p + geom_smooth()

p + geom_point() + geom_smooth() 

p + geom_point() + geom_smooth(method = "lm") 

p + geom_point() +
    geom_smooth(method = "gam") +
    scale_x_log10()

p + geom_point() +
    geom_smooth(method = "gam") +
    scale_x_log10(labels = scales::dollar)
```

## Mapping aesthetics vs setting them {.build .faster}

```{r}
p <-  ggplot(data = gapminder,
             mapping = aes(x = gdpPercap,
                            y = lifeExp,
                            color = continent))
```

## Mapping aesthetics vs setting them {.build .faster}

```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = "purple"))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()

```

## Mapping aesthetics vs setting them {.build .faster}

```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p + geom_point(color = "purple") +
    geom_smooth(method = "loess") +
    scale_x_log10()

```

## Mapping aesthetics vs setting them {.build .faster}

```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp)) 
p + geom_point(alpha = 0.3) +
    geom_smooth(color = "orange", se = FALSE, size = 8, method = "lm") +
    scale_x_log10()

```

## Mapping aesthetics vs setting them {.build .faster}

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y=lifeExp))
p + geom_point(alpha = 0.3) +
    geom_smooth(method = "gam") +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapminder.")

```

## Mapping aesthetics vs setting them {.build .faster}

```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = continent))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

## Mapping aesthetics vs setting them {.build .faster}

```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = continent,
                          fill = continent))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()

```

## Aesthetics can be mapped per geom {.build .faster}

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point(mapping = aes(color = continent)) +
    geom_smooth(method = "loess") +
    scale_x_log10()

```

## Aesthetics can be mapped per geom {.build .faster}

```{r}
library(scales)
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p + geom_point(mapping = aes(color = log(pop))) +
      scale_x_continuous(trans = scales::log_trans())   # natural log scale
     # scale_x_log10()  for base 10 log scale
```

## Save your work {.build .faster}

```{r}
knitr::opts_chunk$set(fig.width=8, fig.height=5) 
```

## Save your work {.build .faster}

```{r example}
p + geom_point()
```

## Save your work {.build .faster}

```{r, fig.width = 12, fig.height = 9}
ggsave(filename = "my_figure.png")
```

## Save your work {.build .faster}

```{r}
ggsave(filename = "my_figure.pdf")
```

## Save your work {.build .faster}

```{r}
p_out <- p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
ggsave("my_figure.pdf", plot = p_out)
```

## Save your work {.build .faster}

```{r}
# here() starts at the directory where the project is created.
# Create the "figures" folder in the directory where the project is created. 
ggsave("lifexp_vs_gdp_gradient.pdf", plot = p_out)
```

## Save your work {.build .faster}

```{r}
ggsave("lifexp_vs_gdp_gradient.pdf",
       plot = p_out, height = 8, width = 10, units = "in")
```

