---
title: "R Lab 4 - Joins"
author: "Byeong-Hak Choe"
---

```{r setup, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggthemes)
library(ggtech)
library(ggthemr)
library(hrbrthemes)
library(hexbin)
library(ggforce)
library(skimr)

knitr::opts_chunk$set(fig.width=8, fig.height=5)  

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray")))


```

## Loading R packages

```{r, eval = T, echo = T}
library(tidyverse)
library(nycflights13)
library(skimr)
```


# Question 1


Consider the following example DataFrames.
```{r}
student_data1 <- data.frame(
  student_id = c('S1', 'S2', 'S3', 'S4', 'S5'),
  name = c('Danniella Fenton', 'Ryder Storey', 'Bryce Jensen', 'Ed Bernal', 'Kwame Morin'),
  marks = c(200, 210, 190, 222, 199)
)

student_data2 <- data.frame(
  student_id = c('S4', 'S5', 'S6', 'S7', 'S8'),
  name = c('Scarlette Fisher', 'Carla Williamson', 'Dante Morse', 'Kaiser William', 'Madeeha Preston'),
  marks = c(201, 200, 198, 219, 201)
)

```




```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(student_data1)
```


```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(student_data2)
```


<br>

## Q1a
Write a R code to join the two given data.frames along **rows** and assign all data.


<br>


## Q1b
Write a R code to join the two given data.frames along **columns** and assign all data.



<br><br><br>

# Question 2

Consider the three data.frames `flights`, `airports`, and `airlines` from the `nycflights13` package.


- `flights` data.frame
```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(flights)
```

```{r}
colnames(flights)  # returns column names in the flights data.frame
```

- `airports` data.frame
```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(airports)
```

```{r}
colnames(airports)  # returns column names in the flights data.frame
```


- `airlines` data.frame

```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(airlines)
```

```{r}
colnames(airlines)
```

<br>

## Q2a

Join `flights` with `airports`, so that we can see information about each flight's destination airport in the `flights` data.frame.



<br>

## Q2b

Find the full name of the airline that has the longest positive `dep_delay` on average.



<br>

## Q2c

Find the full name of the airline that has the largest proportion of flights with longer than 30-minute `dep_delay`.



