---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Homework Assignment 3 - Example Answers<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```

### Loading R packages for Homework Assignment 3

```{r, eval = T, echo = T, warning = F, message = F}
library(tidyverse)
library(lubridate)
library(ggthemes)
```

<br><br>


# Question 1
Read the data file, `NY_school_enrollment_socioecon.csv`, as the data.frame object with the name, `NY_pincp`, using (1) the `read_csv()` function and (2) its URL, `https://bcdanl.github.io/data/NY_school_enrollment_socioecon.csv`.

```{r, echo = T, eval = F}
url <- 'https://bcdanl.github.io/data/NY_school_enrollment_socioecon.csv'
NY_school_enrollment_socioecon <- read_csv(url)
```



<br><br>

For description of variables in `NY_school_enrollment_socioecon`, refer to the file, `ny_school_enrollment_socioecon_description.zip`, which is in the Files section in our Canvas web-page. (I recommend you to extract the zip file, and then read the file, ny_school_enrollment_socioecon_description.csv, using Excel or Numbers.)


<br><br>

## Q1a
Look up the meaning of variable `pincp` from the data.frame, `NY_pincp`. Create the data frame, `NY_pincp`, which has only the following four variables from `NY_school_enrollment_socioecon`:

- `fips` 
- `year` 
- `county_name`
- `pincp`

```{r, eval = T, echo = F, warning = F, message = F}
library(tidyverse)

url <- 'https://bcdanl.github.io/data/NY_school_enrollment_socioecon.csv'
NY_school_enrollment_socioecon <- read_csv(url)
```

```{r, eval = T, echo = T, warning = F, message = F}
NY_pincp <- NY_school_enrollment_socioecon %>% 
  select(fips:pincp)

NY_pincp
```



<br><br>

## Q1b 
Create the data frame, `NY_pincp_wide`, which has one observation for each county (62 observations in total) and the following eight variables:

- `fips`: ID number for county
- `county_name`: County Name
- `pincp2015`: Annual personal income in year 2015
- `pincp2016`: Annual personal income in year 2016
- `pincp2017`: Annual personal income in year 2017
- `pincp2018`: Annual personal income in year 2018
- `pincp2019`: Annual personal income in year 2019
- `pincp2020`: Annual personal income in year 2020

```{r, eval = T, echo = T, warning = F, message = F}
NY_pincp_wide <- NY_pincp %>% 
  pivot_wider(names_from = year,
              values_from = pincp,
              names_prefix = 'pincp')
```





<br><br>

## Q1c
Use the data.frame `NY_pincp_wide` from Q1b to create the data frame, NY_pincp_long, which has six observations for each county (372 observations in total) and the following four variables:

- `fips`: ID number for county 
- `year`: integer variable of year (e.g., The possible values of `year` are 2015, 2016, 2017, 2018, 2019, and 2020.)
- `county_name`: County Name
- `pincp`: Annual personal income


```{r, eval = T, echo = T, warning = F, message = F}
# (1) using separate()
NY_pincp_long <- NY_pincp_wide %>% 
  pivot_longer( cols = pincp2015:pincp2020,
                names_to = "year",
                values_to = "pincp") %>% 
  separate(year, into = c("dum", "year"), 
           sep = "pincp", convert = T) %>% 
  select(-dum)

# () using str_replace()
NY_pincp_long <- NY_pincp_wide %>% 
  pivot_longer( cols = pincp2015:pincp2020,
                names_to = "year",
                values_to = "pincp") %>% 
  mutate( year = str_replace(year, "pincp", ""),
          year = as.integer(year) )
```




<br><br>

## Q1d
Create the data.frame, `NY_pincp_geo`, by join the two data.frames, `NY_county_geo` and `NY_pincp`.

```{r, eval = T, echo = T, warning = F, message = F}
NY_county_geo <- read_csv('https://bcdanl.github.io/data/NY_county_geo.csv')
```

The data.frame, `NY_pincp_geo`, must include all the observations and variables in the data.frame, `NY_county_geo`.

```{r, eval = T, echo = T, warning = F, message = F}
NY_pincp_geo <- NY_county_geo %>% 
  left_join(NY_pincp, by = c("FIPS" = "fips"))
```




<br><br>

## Q1e
In the following ggplot code with `geom_polygon()`, replace the blanks ([?]) with the appropriate object to draw a yearly county map of `pincp`.

```{r, eval = F, echo = T, warning = F, message = F}
# ggplot(data = [?]) +
#   geom_polygon(mapping = aes(x = [?], y = [?], group = group, 
#                              fill = [?] ) ,
#                color = "grey", size = 0.1) +
#   facet_wrap( [?] ) +
#   labs( title = "Income in NY State" ) +
#   coord_map("bonne", parameters = 41.6) +  # for better aspect ratio
#   scale_fill_gradient2( #  mapping pincp rate to color level
#   low = 'red', 
#   high = 'blue', 
#   na.value = "grey50",
#   midpoint = quantile(
#     NY_school_enrollment_socioecon$pincp, 50)) +
#   theme_map() +    # a better theme for map drawing
#   theme(legend.position='top') 
```


```{r, eval = T, echo = T, warning = F, message = F}
ggplot(data = NY_pincp_geo) +
  geom_polygon(mapping = aes(x = long, y = lat, group = group, 
                             fill = pincp ) ,
               color = "grey", size = 0.1) +
  facet_wrap( ~ year ) +
  labs( title = "Average Income in NY Counties, 2015-2020",
        caption = "Source: Breau of Labor Statistics") +
  coord_map("bonne", parameters = 41.6) +  # for better aspect ratio
  scale_fill_gradient2( #  mapping pincp values to color levels
    low = 'red', 
    high = 'blue', 
    na.value = "grey50",
    midpoint = quantile(
      NY_school_enrollment_socioecon$pincp, .50),
    name = "Average Income") +
  theme_map() +    # a better theme for map drawing
  theme(legend.position= "bottom",
        plot.title = element_text(hjust = .5, size = 16,
                                  color = "purple"),
        strip.background = element_rect(color = 'white', 
                                        fill = 'white')) 
```



<br><br><br>


# Question 2
Read the data file, `tech_stock.csv`, as the data.frame object with the name, `tech_stock`, using (1) the `read_csv()` function and (2) its URL, `https://bcdanl.github.io/data/tech_stock`.

```{r, eval = T, echo = T, warning = F, message = F}
tech_stock <- read_csv("https://bcdanl.github.io/data/tech_stock.csv")
tech_stock
```

<br><br>


## Q2a
Describe the daily trend of `Close` for each company since 2010 in one ggplot.


```{r, eval = T, echo = T, warning = F, message = F}
ggplot(filter(tech_stock, Date > ymd("2010-01-01")),
       aes(x = Date, y= Close,
           color = company)) +
  geom_line() +
  hrbrthemes::theme_ipsum()

ggplot(filter(tech_stock, Date > ymd("2010-01-01")),
       aes(x = Date, y= Close,
           color = fct_reorder2(company, Date, Close))) +
  geom_line() + 
  hrbrthemes::theme_ipsum()
```


<br><br>



## Q2b
- Describe the relationship between `log(Close)` and `log(Volume)` for the companies in `tech_stock`.

```{r, eval = T, echo = T, warning = F, message = F}
p0 <- ggplot(tech_stock , 
             aes(x = log(Close), y = log(Volume)) )
p0 + geom_point(alpha = .05,
                shape = 1,
                aes(color = company)) + geom_smooth(method = lm) +
  guides(colour = guide_legend(override.aes = list(alpha=1)))
```



<br><br>

## Q2c
Does the relationship between `log(Close)` and `log(Volume)` vary by companies and by the time periods?

```{r, eval = T, echo = T, warning = F, message = F}
# all years
p0 <- ggplot(tech_stock , 
             aes(x = log(Close), y = log(Volume),
                 color = company) )
p0 + geom_point(alpha = .05,
                shape = 1) + geom_smooth(method = lm) +
  facet_grid(.~company) +
  theme_minimal()+
  labs(title = "All years") +
  theme_bw() +
  guides(color = "none")



# since 2017
tech_stock2 <- tech_stock %>% 
  filter(Date >= ymd("2017-01-01")) %>% 
  mutate(year = year(Date))

p0 <- ggplot(tech_stock2 , 
             aes(x = log(Close), y = log(Volume),
                 color = company) )
p0 + geom_point(alpha = .075,
                shape = 1) + 
  geom_smooth(method = lm, se = F, size = .75) + 
  facet_grid(year~company, scales = "free_x") +
  labs(title = "Since 2017") +
  theme_bw() +
  guides(color = "none")

```

