---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Midterm Exam, Spring 2022<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```

### Loading R packages for the Midterm Exam

```{r, eval = F, echo = T}
library(tidyverse)
```

<br><br>


# Question 1
## Q1a. 

Download `dominick_oj_q1a.csv` from the Midterm Exam in the Assignments or the Files sections in our Canvas.

Then import the `dominick_oj_q1a.csv` using the following lines:

```{r, echo = T, eval = F}
oj_q1a <- read_csv('ABSOLUTE_PATH_NAME_FOR_THE_FILE_dominick_oj_q1a.csv')
table(oj_q1a$brand)
```

<br>

You need to provide the absolute path name for the file, `dominick_oj_q1a.csv` to the above `read_csv()` function to read the file.

### Variable Description

- `sales`: the number of orange juice (OJ) cartons sold in a week
- `price`: price of OJ carton
- `brand`: OJ brand
- `feat`: Advertisement status--- 1 if advertised; 0 if not advertised.


- Report (1) minimum, (2) median, (3) maximum, (4) mean, and (5) standard deviation of variable `price` for the `brand`, Dominick's OJ.


<br><br>

## Q1b
For Question 1b, run the following function to read the `dominick_oj.csv` file:
```{r, echo = T, eval = F}
oj_q1b <- read_csv(
  'https://bcdanl.github.io/data/dominick_oj.csv'
)
```

<br>

The description of variables in `oj_q1b` is the same as `oj_q1a`.

- Describe the relationship between the log of `price` and the log of `sales` by `brand` using ggplot. Make a simple comment on your ggplot figure.


<br><br><br>

# Question 2

For Question 2, run the following R command to read the `nyc_dogs.csv` file.

```{r, echo = T, eval = F}
nyc_dogs <- read_csv('https://bcdanl.github.io/data/nyc_dogs.csv')
```

<br>

## Q2a 
Describe the distribution of animal_gender using ggplot. Make a simple comment on your ggplot figure.

<br><br>

## Q2b
Find the five most popular breeds in NYC.

<br><br>

## Q2c
Describe the relationship between the five popular breeds and borough using ggplot. Make a simple comment on your ggplot figure.

<br><br>

## Q2d
Find the five most popular breeds for each borough in NYC.

<br><br>

## Q2e 
Find the five most popular dog names for each gender in NYC.

<br><br>

## Q2f
Find the five most popular dog names for each gender for each borough in NYC.

<br><br>

## Q2g
Assume that all dogs in the `nyc_dogs` data frame are alive as of today.

Describe the distribution of age for each borough using ggplot. Make a simple comment on your ggplot.

<br><br><br>

# Question 3

For Question 3, run the following function to read the NYC's Citywide Payroll Data.

```{r, echo = T, eval = F}
nyc_payroll <- read_csv(
  'https://bcdanl.github.io/data/nyc_payroll.csv'
)
```

<br>

Description of variables in the nyc_payroll dataset is provided at the end of the R script.

<br>

## Q3a. 
Create a variable, `payroll`, which is defined as:

$$ \texttt{payroll} = \texttt{regular_gross_paid} + \texttt{total_ot_paid}$$
where `regular_gross_paid` and `total_ot_paid` are variables in the `nyc_payroll` data frame.


<br><br>

## Q3b. 
Calculate the mean of `payroll` by `title_description`.


<br><br>


## Q3c. 
Calculate the mean of `payroll` by `work_location_borough`.



<br><br>

## Variable Description

1. Fiscal Year: Fiscal Year

2. Payroll Number: Payroll Number

3. Agency Name: The Payroll agency that the employee works for

4. Last Name: Last name of employee

5. First Name: First name of employee

6. Mid Init: Middle initial of employee

7. Agency Start Date: Date which employee began working for their current agency Date & Time

8. Work Location Borough: Borough of employee's primary work location

9. Title Description: Civil service title description of the employee

10. Leave Status as of June 30: Status of employee as of the close of the relevant fiscal year: Active, Ceased, or On Leave

11. Base Salary: Base Salary assigned to the employee

12. Pay Basis: Lists whether the employee is paid on an hourly, per diem or annual basis

13. Regular Hours: Number of regular hours employee worked in the fiscal year

14. Regular Gross Paid: The amount paid to the employee for base salary during the fiscal year

15. OT Hours: Overtime Hours worked by employee in the fiscal year

16. Total OT Paid: Total overtime pay paid to the employee in the fiscal year

17. Total Other Pay: Includes any compensation in addition to gross salary and overtime pay, ie Differentials, lump sums, uniform allowance, meal allowance, retroactive pay increases, settlement amounts, and bonus pay, if applicable.



