---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Midterm Exam, Spring 2022<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```

### Loading R packages for the Midterm Exam

```{r, eval = F, echo = T}
library(tidyverse)
```

<br><br>


# Question 1
## Q1a. 

Download `dominick_oj_q1a.csv` from the Midterm Exam in the Assignments or the Files sections in our Canvas.

Then import the `dominick_oj_q1a.csv` using the following lines:

```{r, echo = T, eval = F}
oj_q1a <- read_csv('ABSOLUTE_PATH_NAME_FOR_THE_FILE_dominick_oj_q1a.csv')
table(oj_q1a$brand)
```

<br>

You need to provide the absolute path name for the file, `dominick_oj_q1a.csv` to the above `read_csv()` function to read the file.

### Variable Description

- `sales`: the number of orange juice (OJ) cartons sold in a week
- `price`: price of OJ carton
- `brand`: OJ brand
- `feat`: Advertisement status--- 1 if advertised; 0 if not advertised.


- Report (1) minimum, (2) median, (3) maximum, (4) mean, and (5) standard deviation of variable `price` for the `brand`, Dominick's OJ.


```{r, echo = T, eval = T, message = F, warning= F}
library(tidyverse)
oj_q1a <- read_csv('/Users/byeong-hakchoe/Google Drive/suny-geneseo/teaching-materials/lecture-data/dominick_oj_q1a.csv')
table(oj_q1a$brand)

library(skimr)
oj_q1a_d <- filter(oj_q1a, brand == 'dominicks')
skim(oj_q1a_d$price)
```

 (1) minimum, (2) median, (3) maximum, (4) mean, and (5) standard deviation of variable `price` for the `brand`, `dominicks` are ......


<br><br>

## Q1b
For Question 1b, run the following function to read the `dominick_oj.csv` file:
```{r, echo = T, eval = T, message = F, warning= F}
oj_q1b <- read_csv(
  'https://bcdanl.github.io/data/dominick_oj.csv'
)
```

<br>

The description of variables in `oj_q1b` is the same as `oj_q1a`.

- Describe the relationship between the log of `price` and the log of `sales` by `brand` using ggplot. Make a simple comment on your ggplot figure.


```{r, echo = T, eval = T, warning= F, message= F}
summary(oj_q1b)

ggplot(data = oj_q1b, 
       mapping = aes(x = log(price), y = log(sales),
                     color = brand)) +
  geom_point(alpha = .05) +
  geom_smooth(method = lm, se = F)
```

Here we observe the downward sloping demand curve for each OJ brand.

For the same level of price, the demand for `tropicana` is the highest, followed by `minute.maid` and `dominicks`.

- `tropicana` is the luxurious OJ; 
- `dominicks` OJ seems to be a budget option.
- `minute.maid` is somewhat between the two.


<br><br><br>

# Question 2

For Question 2, run the following R command to read the `nyc_dogs.csv` file.

```{r, echo = T, eval = T, warning= F, message= F}
nyc_dogs <- read_csv('https://bcdanl.github.io/data/nyc_dogs.csv')
```

<br>

## Q2a 
Describe the distribution of `animal_gender` using ggplot. Make a simple comment on your ggplot figure.

```{r, echo = T, eval = T, warning= F, message= F}
ggplot(nyc_dogs) +
  geom_bar(aes(x = animal_gender))

ggplot(nyc_dogs) +
  geom_density(aes(x = animal_gender))

ggplot(nyc_dogs) +
  geom_boxplot(aes(x = animal_gender))

ggplot(nyc_dogs) +
  geom_bar(aes(x = animal_gender)) +
  facet_wrap(~borough)

```

- There are more male dogs than female dogs.

- There are more male dogs than female dogs in each borough.

<br><br>

## Q2b
Find the five most popular breeds in NYC.

```{r, echo = T, eval = T, warning= F, message= F}
# 0.
table(nyc_dogs$breed_rc)


# 1.
q2b <- group_by(nyc_dogs, 
                breed_rc)

q2b <- summarise(q2b,
                 n = n())

q2b <- arrange(q2b,
               -n)  # or desc(n)
q2b

# 2.
q2b <- group_by(nyc_dogs, 
                breed_rc)

q2b <- mutate(q2b,
              n = n())

q2b <- arrange(q2b,
               desc(n))

q2b <- select(q2b, 
              breed_rc, n)

q2b <- distinct(q2b)
q2b
```

1 Unknown                  54586
2 Yorkshire Terrier        30379
3 Labrador (or Crossbreed) 28399
4 Shih Tzu                 27407
5 Pit Bull (or Mix)        24393
6 Chihuahua                21211


<br><br>

## Q2c
Describe the relationship between the five popular breeds and borough using ggplot. Make a simple comment on your ggplot figure.

```{r, echo = T, eval = T, warning= F, message= F}
q2c <- filter(nyc_dogs,
              breed_rc %in% c("Yorkshire Terrier", 
                              "Labrador (or Crossbreed)",
                              "Shih Tzu", 
                              "Pit Bull (or Mix)",
                              "Chihuahua"),
              !is.na(borough)
                         )

q2c <- filter(nyc_dogs,
              breed_rc == "Yorkshire Terrier" | 
              breed_rc == "Labrador (or Crossbreed)" | 
              breed_rc == "Shih Tzu" | 
              breed_rc == "Pit Bull (or Mix)" | 
              breed_rc == "Chihuahua",
              !is.na(borough)
)

# 1. stacked bar charts
# distribution of `breed_rc` by `borough` 
ggplot(q2c) +
  geom_bar(aes(x = breed_rc, fill = borough)) +
  coord_flip()

# distribution of `borough` by `breed_rc` 
ggplot(q2c) +
  geom_bar(aes(fill = breed_rc, x = borough)) +
  coord_flip()


# 2. stacked bar charts of relative frequencies
# distribution of `breed_rc` by `borough` 
ggplot(q2c) +
  geom_bar(aes(x = breed_rc, fill = borough), 
           position = "fill") +
  coord_flip()

# distribution of `borough` by `breed_rc` 
ggplot(q2c) +
  geom_bar(aes(fill = breed_rc, x = borough), 
           position = "fill") +
  coord_flip()


# 3. side-by-side bar charts 
# distribution of `breed_rc` by `borough` 
ggplot(q2c) +
  geom_bar(aes(x = breed_rc, fill = borough), 
           position = "dodge2") +
  coord_flip()

# distribution of `borough` by `breed_rc` 
ggplot(q2c) +
  geom_bar(aes(fill = breed_rc, x = borough), 
           position = "dodge2") +
  coord_flip()


# 4. side-by-side bar charts of relative frequencies
# distribution of `breed_rc` by `borough` 
ggplot(q2c) +
  geom_bar(aes(x = breed_rc, fill = borough, 
               y = ..prop.., group = borough), 
           position = "dodge2") +
  coord_flip()

# distribution of `borough` by `breed_rc` 
ggplot(q2c) +
  geom_bar(aes(x = borough, fill = breed_rc, 
               y = ..prop.., group = breed_rc), 
           position = "dodge2") +
  coord_flip()

```


```{r, echo = T, eval = T, warning= F, message= F, fig.width= 8, fig.height= 60}
# What if the filtered data could not be obtained ... 
ggplot(data = nyc_dogs)+
  geom_bar(mapping = aes(x=borough))+
  facet_wrap(~breed_rc, ncol = 4) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  # optional
```

<br><br>

## Q2d
Find the five most popular breeds for each borough in NYC.

```{r, echo = T, eval = T, warning= F, message= F}
# 1. 
q2d <- group_by(nyc_dogs,
                breed_rc, borough)
q2d <- summarize(q2d,
                 n = n())
q2d <- arrange(q2d,
               borough, -n)
  
# 2.
q2d <- group_by(nyc_dogs,
                breed_rc, borough)
q2d <- mutate(q2d,
              n = n())
q2d <- arrange(q2d,
               -n)
q2d <- select(q2d,
              breed_rc, n, borough)
q2d <- distinct(q2d)


# filter by borough
q2d_bronx <- filter(q2d, borough == "Bronx")
q2d_bronx

q2d_brooklyn <- filter(q2d, borough == "Brooklyn")
q2d_brooklyn

q2d_manhattan <- filter(q2d, borough == "Manhattan")
q2d_manhattan

q2d_queens <- filter(q2d, borough == "Queens")
q2d_queens

q2d_staten_island <- filter(q2d, borough == "Staten Island")
q2d_staten_island



# 3.
q2d <- group_by(nyc_dogs,
                breed_rc, borough)

q2d <- mutate(q2d,
              n = n())

q2d <- group_by(q2d,
                borough)

q2d <- mutate(q2d,
              rank = dense_rank(desc(n)) )

q2d <- filter(q2d, rank <= 10 )

q2d <- select(q2d,
              breed_rc, n, rank, borough)

q2d <- distinct(q2d)

q2d <- arrange(q2d,
               borough,-n)

print(q2d, n = Inf)

```

<br><br>

## Q2e 
Find the five most popular dog names for each gender in NYC.

```{r, echo = T, eval = T, warning= F, message= F}
# 1.
q2e <- group_by(nyc_dogs,
                animal_name, animal_gender)
q2e <- summarise(q2e,
                 n = n())
q2e <- arrange(q2e,
               animal_gender, -n)


# 2. 
q2e <- group_by(nyc_dogs,
                animal_name, animal_gender)
q2e <- mutate(q2e,
              n = n())
q2e <- arrange(q2e,
               -n)
q2e <- select(q2e,
              animal_name, animal_gender, n)
q2e <- distinct(q2e)



# filter by animal_gender
q2e_F <- filter(q2e, animal_gender == "F")
q2e_F

q2e_M <- filter(q2e, animal_gender == "M")
q2e_M

```

<br><br>

## Q2f
Find the five most popular dog names for each gender for each borough in NYC.

```{r, echo = T, eval = T, warning= F, message= F}
popular_name <- nyc_dogs %>% 
  group_by(animal_name, animal_gender, borough) %>% 
  mutate(n = n()) %>% 
  arrange(borough, animal_gender, -n) %>% 
  select(animal_gender, animal_name, n) %>% 
  distinct()


# 1.
q2f <- group_by(nyc_dogs,
                animal_name, animal_gender, borough)
q2f <- summarise(q2f,
                 n = n())
q2f <- arrange(q2f,
               animal_gender, -n)


# 2. 
q2f <- group_by(nyc_dogs,
                animal_name, animal_gender, borough)
q2f <- mutate(q2f,
              n = n())
q2f <- arrange(q2f,
               -n)
q2f <- select(q2f,
              animal_name, animal_gender, borough, n)
q2f <- distinct(q2f)



# filter by animal_gender
q2f_F <- filter(q2f, animal_gender == "F")
q2f_M <- filter(q2f, animal_gender == "M")


# filter `q2f_F` by borough
q2f_F_bronx <- filter(q2f_F, borough == "Bronx")
q2f_F_bronx

q2f_F_brooklyn <- filter(q2f_F, borough == "Brooklyn")
q2f_F_brooklyn

q2f_F_manhattan <- filter(q2f_F, borough == "Manhattan")
q2f_F_manhattan

q2f_F_queens <- filter(q2f_F, borough == "Queens")
q2f_F_queens

q2f_F_staten_island <- filter(q2f_F, borough == "Staten Island")
q2f_F_staten_island


# filter `q2f_M` by borough
q2f_M_bronx <- filter(q2f_M, borough == "Bronx")
q2f_M_bronx

q2f_M_brooklyn <- filter(q2f_M, borough == "Brooklyn")
q2f_M_brooklyn

q2f_M_manhattan <- filter(q2f_M, borough == "Manhattan")
q2f_M_manhattan

q2f_M_queens <- filter(q2f_M, borough == "Queens")
q2f_M_queens

q2f_M_staten_island <- filter(q2f_M, borough == "Staten Island")
q2f_M_staten_island



# 3.
q2f <- group_by(nyc_dogs,
                animal_name, animal_gender, borough)

q2f <- mutate(q2f,
              n = n())

q2f <- group_by(q2f,
                animal_gender, borough)

q2f <- mutate(q2f,
              rank = dense_rank(desc(n)) )

q2f <- filter(q2f, 
              rank <= 10 )

q2f <- select(q2f,
              animal_name, animal_gender, n, rank, borough)

q2f <- distinct(q2f)

q2f <- arrange(q2f,
               borough, animal_gender, -n)

print(q2f, n = Inf)

```

<br><br>

## Q2g
Assume that all dogs in the `nyc_dogs` data frame are alive as of today.

Describe the distribution of age for each borough using ggplot. Make a simple comment on your ggplot.

```{r, echo = T, eval = T, warning= F, message= F}
nyc_dogs_age <- nyc_dogs %>% 
  mutate( age = 2022 - animal_birth_year )

nyc_dogs_age <- mutate(nyc_dogs, 
                       age = 2022 - animal_birth_year )

ggplot( nyc_dogs_age ) +
  geom_density(aes(x = age)) + 
  facet_wrap(~borough)

ggplot( filter(nyc_dogs_age, 
               age <= 25, !is.na(borough)) ) +
  geom_density(aes(x = age)) + 
  facet_wrap(~borough)


ggplot( filter(nyc_dogs_age, age <= 25, !is.na(borough)) ) +
  geom_histogram(aes(x = age), 
                 binwidth = 1) + 
  facet_wrap(~borough)

```

Add some comments here ....

<br><br><br>

# Question 3

For Question 3, run the following function to read the NYC's Citywide Payroll Data.

```{r, echo = T, eval = T, warning= F, message= F}
nyc_payroll <- read_csv(
  'https://bcdanl.github.io/data/nyc_payroll.csv'
)
```

<br>

Description of variables in the nyc_payroll dataset is provided at the end of the R script.

<br>

## Q3a. 
Create a variable, `payroll`, which is defined as:

$$ \texttt{payroll} = \texttt{regular_gross_paid} + \texttt{total_ot_paid}$$
where `regular_gross_paid` and `total_ot_paid` are variables in the `nyc_payroll` data frame.

```{r, echo = T, eval = T, warning= F, message= F}
q3a <-  mutate(nyc_payroll,
               payroll = regular_gross_paid + total_ot_paid)
```


<br><br>

## Q3b. 
Calculate the mean of `payroll` by `title_description`.


```{r, echo = T, eval = T, warning= F, message= F}
q3b <- group_by(q3a, 
                title_description)
q3b <-  summarise(q3b, 
                  avg_payroll = mean(payroll, na.rm = T))
q3b <- arrange(q3b, 
               -avg_payroll)
```

<br><br>


## Q3c. 
Calculate the mean of `payroll` by `work_location_borough`.

```{r, echo = T, eval = T, warning= F, message= F}
q3c <- group_by(q3a, 
                work_location_borough)
q3c <-  summarise(q3c, 
                  avg_payroll = mean(payroll, na.rm = T))
q3c <- arrange(q3c, 
               -avg_payroll)
```

<br><br>

## Variable Description

1. Fiscal Year: Fiscal Year

2. Payroll Number: Payroll Number

3. Agency Name: The Payroll agency that the employee works for

4. Last Name: Last name of employee

5. First Name: First name of employee

6. Mid Init: Middle initial of employee

7. Agency Start Date: Date which employee began working for their current agency Date & Time

8. Work Location Borough: Borough of employee's primary work location

9. Title Description: Civil service title description of the employee

10. Leave Status as of June 30: Status of employee as of the close of the relevant fiscal year: Active, Ceased, or On Leave

11. Base Salary: Base Salary assigned to the employee

12. Pay Basis: Lists whether the employee is paid on an hourly, per diem or annual basis

13. Regular Hours: Number of regular hours employee worked in the fiscal year

14. Regular Gross Paid: The amount paid to the employee for base salary during the fiscal year

15. OT Hours: Overtime Hours worked by employee in the fiscal year

16. Total OT Paid: Total overtime pay paid to the employee in the fiscal year

17. Total Other Pay: Includes any compensation in addition to gross salary and overtime pay, ie Differentials, lump sums, uniform allowance, meal allowance, retroactive pay increases, settlement amounts, and bonus pay, if applicable.



