---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Homework Assignment 1<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---


### Loading R packages for Homework Assignment 1

```{r, eval = F, echo = T}
library(tidyverse)
```

<br><br>

# Question 1

## Q1a
Download the compressed file, `bikeshare-2011-01-01.zip`, 
from the Files section in our Canvas web-site. Extract the file, `bikeshare-2011-01-01.zip`, so that you can use the file, `bikeshare-2011-01-01.csv`. Read the data file, `bikeshare-2011-01-01.csv`, as the data.frame object with the name, `bikeshare2011_01_01`, using (1) the `read_csv()` function and (2) the *absolute path* name of the file `bikeshare_2011_01_01.csv` from your local hard disk drive in your laptop.

```{r}

```

<br><br>

## Q1b

Report the mean, median, minimum, maximum, and standard deviation for each numeric variable in the data.frame `bikeshare2011_01_01`.

```{r}

```


<br><br><br>

# Question 2

## Q2a

Read the data file, `bikeshare_cleaned.csv`, as the data.frame object with the name, `bikeshare`, using (1) the `read_csv()` function and (2) its URL, `https://bcdanl.github.io/data/bikeshare_cleaned.csv`.


```{r}

```


<br><br>

Use the data.frame `bikeshare` for the rest of questions in Question 2.

### Description of variables in the data file, `bikeshare_cleaned.csv`

The data set, `bikeshare_cleaned.csv`, includes 17376 observations of hourly counts from 2011 to 2012 for bike rides (rentals) in Washington D.C.

<br>

- `cnt`: count of total bikes rented out
- `year`: year
- `month`: month
- `date`: date
- `hr`: hours 
- `wkday`: week day
- `holiday`: holiday if holiday == 1; non-holiday otherwise
- `seasons`: season
- `weather_cond`: weather condition
- `temp`: temperature, measured in standard deviations from average.
- `hum`: humidity, measured in standard deviations from average.
- `windspeed`: wind speed, measured in standard deviations from average.



<br><br>

## Q2b
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the distribution of `cnt`.

```{r}

```

<br><br>

## Q2c
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the distribution of `cnt` by `year` and `month`.

```{r}

```

<br><br>

## Q2d
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the distribution of `temp` by `year` and `month`.

```{r}

```



<br><br>

## Q2e
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the distribution of `hum` by `year` and `month`.

```{r}

```






<br><br>

## Q2f
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the distribution of `windspeed` by `year` and `month`.

```{r}

```


<br><br>

## Q2g
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the relationship between `temp` and `cnt`.

```{r}

```

<br><br>

## Q2h
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the relationship between `temp` and `cnt` by `year` and `month`.

```{r}

```

<br><br>

## Q2i
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the relationship between `weather_cond` and `cnt`.

```{r}

```



<br><br>

## Q2j
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the relationship between `weather_cond` and `cnt` by `hr`.

```{r}

```




<br><br>

## Q2k
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the relationship between `wkday` and `cnt`.

```{r}

```



<br><br>

## Q2l
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the relationship between `wkday` and `cnt` by `hr`.

```{r}

```




<br><br><br>

# Question 3
## Q3a
Read the data file, `NY_school_enrollment_socioecon.csv`, as the data.frame object with the name, `NY_school_enrollment_socioecon`, using (1) the `read_csv()` function and (2) its URL, `https://bcdanl.github.io/data/NY_school_enrollment_socioecon.csv`.

```{r}

```



<br><br>

For description of variables in `NY_school_enrollment_socioecon`, refer to the file, `ny_school_enrollment_socioecon_description.zip`, which is in the Files section in our Canvas web-page. (I recommend you to extract the zip file, and then read the file, ny_school_enrollment_socioecon_description.csv, using Excel or Numbers.)


<br><br>

## Q3b
Provide both (1) `ggplot` codes and (2) a couple of sentences to describe the relationship between college enrollment and educational attainment of population 45 to 64 years, and how such relationship varies by the type (public or private) of colleges.

```{r}

```






<br><br>

## Q3c 
Provide both (1) `ggplot` codes and  (2) a couple of sentences to describe how the relationships described in Q3b vary by gender of population 45 to 64 years.

```{r}

```





<br><br>

## Q3d
Provide both (1) `ggplot` codes and  (2) a couple of sentences to describe how the relationships described in Q3b vary by gender of college enrollment.

```{r}

```




