---
title: "DANL 100: Programming for Data Analytics"
subtitle: "DANL 100 - Homework Assignment 5 - Example Answers<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```


- The following is the Python libraries we need for this homework.
```{python, echo = T, eval = F}
import pandas as pd
import seaborn as sns
```



# Question 1

Load the CSV file, `beer_markets_cleaned.csv`, from the class website:
```{python, echo = T, eval = F}
beer_markets = pd.read_csv('https://bcdanl.github.io/data/beer_markets_cleaned.csv')
```


- Description of DataFrame `beer_markets` 
  - One single observation (row) corresponds to information regarding one household and its beer purchase.
  - DataFrame `beer_markets` does not have any missing values.

<br>

- Variable (column) description
  - `hh`: an identifier of the household;
  - `purchase_desc`: details on the purchased item;
  - `quantity`: the number of items purchased;
  - `brand`: Bud Light, Busch Light, Coors Light, Miller Lite, or Natural Light;
  - `spent`: total dollar value of purchase;
  - `beer_floz`: total volume of beer, in fluid ounces;
  - `price_per_floz`: price per fl.oz. (i.e., beer spent/beer floz);
  - `container`: the type of container;
  - `promo`: Whether the item was promoted (coupon or otherwise);
  - `market`: Scan-track market (or state if rural);
  - demographic data, including gender, marital status, household income, class of work, race, education, age, the size of household, and whether or not the household has a microwave or a dishwasher.

<br><br>

## Q1a
- Sort the DataFrame `beer_markets` by `hh` in ascending order.


<br>

*Answer*

```{python, echo = T, eval = F}
beer_markets = beer_markets.sort_values('hh')
```

<br><br>

## Q1b
Count the number of households for each market.
<br>

*Answer*

```{python, echo = T, eval = F}
# The below line counts the number of non-missing values
# for each variable for each group of "market" and "hh".
q1b = beer_markets.groupby(["market", "hh"]).count() 

# In the resulting DataFrame q1b above, 
# there is only one single observation for each group of
# "market" and "hh".
# So, the following line counts the number of households
# for each market.
q1b = q1b.groupby(["market"]).count()

q1b = q1b.loc[ :, ["brand"] ]
q1b.columns = ['n']   # rename `brand` with `n`

# For renaming, the following also works:
# q1b = q1b.rename({'brand':"n"}, axis = 'columns')
```

<br><br>

## Q1c

Find the top 5 beer markets in terms of the number of households that purchased beer.

<br>

*Answer*

```{python, echo = T, eval = F}
q1c = q1b.sort_values('n', ascending = False)
```

```{r, echo = F, eval = T, warning=F, message=F}
library(tidyverse)
library(kableExtra)
text_tbl <- data.frame(
  market = c("TAMPA", "DETROIT", "COLUMBUS",
             "MIAMI", "PHOENIX"),
  n = c(280, 214, 184, 178, 168)
)

kbl(text_tbl) %>%
  kable_paper("hover", full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, width = "5em")
```

<br><br>

## Q1d
Sum of `beer_floz` for each `market`.
<br>

*Answer*

```{python, echo = T, eval = F}
q1d = beer_markets.groupby(["market"]).sum()
q1d = q1d.loc[ :, ["beer_floz"] ]

# the following gives the Series, instead of the DataFrame
# q1d = q1d['beer_floz'] 
```

<br><br>

## Q1e
Find the top 5 beer markets in terms of the amount of total beer consumption.
<br>

*Answer*

```{python, echo = T, eval = F}
q1e = q1d.sort_values('beer_floz', ascending = False)
```


```{r, echo = F, eval = T, warning=F, message=F}
text_tbl <- data.frame(
  market = c("TAMPA", "PHOENIX", "DETROIT",
             "MIAMI", "COLUMBUS"),
  beer_floz = c(462904, 388824, 309588, 271870, 266896)
)

kbl(text_tbl) %>%
  kable_paper("hover", full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, width = "5em")
```

<br><br>

## Q1f
- Variable `price_per_floz` is continuous.
- Variable `brand` is categorical.

Describe the distribution of `price_per_floz` for each `brand` using `seaborn`.

Make a simple comment on comparison for the distribution of `price_per_floz` across brands.
<br>

*Answer*

```{python, echo = T, eval = F}
# distribution of price_per_floz for BUD LIGHT
sns.displot( x = 'price_per_floz', bins = 200,
             data = beer_markets[ beer_markets['brand'] == 'BUD LIGHT' ] )

# distribution of price_per_floz for COORS LIGHT
sns.displot( x = 'price_per_floz', bins = 200,
             data = beer_markets[ beer_markets['brand'] == 'COORS LIGHT' ] )

# distribution of price_per_floz for MILLER LITE
sns.displot( x = 'price_per_floz', bins = 200,
             data = beer_markets[ beer_markets['brand'] == 'MILLER LITE' ] )

# distribution of price_per_floz for BUSCH LIGHT
sns.displot( x = 'price_per_floz', bins = 200,
             data = beer_markets[ beer_markets['brand'] == 'BUSCH LIGHT' ] )

# distribution of price_per_floz for NATURAL LIGHT
sns.displot( x = 'price_per_floz', bins = 200,
             data = beer_markets[ beer_markets['brand'] == 'NATURAL LIGHT' ] )
```

```{r, echo = F, eval = T, warning=F, message=F}
library(tidyverse)
beer_markets <- read_csv('https://bcdanl.github.io/data/beer_markets_cleaned.csv')
beer <- c('BUD LIGHT', 'COORS LIGHT', 'MILLER LITE',
  'BUSCH LIGHT', 'NATURAL LIGHT')

my_function <- function(bname) { 
  ggplot(filter(beer_markets,
              brand == bname)) +
  geom_histogram(aes(x = price_per_floz), 
                 bins = 200,
                 fill = 'deepskyblue3',
                 color = 'grey70') +
  theme_minimal() +
  ggtitle(bname)
}

ggplot(filter(beer_markets,
              brand == 'BUD LIGHT')) +
  geom_histogram(aes(x = price_per_floz), 
                 bins = 200,
                 fill = 'deepskyblue3',
                 color = 'grey70') +
  theme_minimal() +
  ggtitle('BUD LIGHT')

ggplot(filter(beer_markets,
              brand == 'COORS LIGHT')) +
  geom_histogram(aes(x = price_per_floz), 
                 bins = 200,
                 fill = 'deepskyblue3',
                 color = 'grey70') +
  theme_minimal() +
  ggtitle('COORS LIGHT')

ggplot(filter(beer_markets,
              brand == 'MILLER LITE')) +
  geom_histogram(aes(x = price_per_floz), 
                 bins = 200,
                 fill = 'deepskyblue3',
                 color = 'grey70') +
  theme_minimal() +
  ggtitle('MILLER LITE')

ggplot(filter(beer_markets,
              brand == 'BUSCH LIGHT')) +
  geom_histogram(aes(x = price_per_floz), 
                 bins = 200,
                 fill = 'deepskyblue3',
                 color = 'grey70') +
  theme_minimal() +
  ggtitle('BUSCH LIGHT')

ggplot(filter(beer_markets,
              brand == 'NATURAL LIGHT')) +
  geom_histogram(aes(x = price_per_floz), 
                 bins = 200,
                 fill = 'deepskyblue3',
                 color = 'grey70') +
  theme_minimal() +
  ggtitle('NATURAL LIGHT')
```

<br>

- `BUD LIGHT`, `COORS LIGHT`, and `MILLER LITE` have the similar distribution of `price_per_floz` with each other.

- `BUSCH LIGHT` and `NATURAL LIGHT` have the similar distribution of `price_per_floz` with each other.

- Overall, `BUD LIGHT`, `COORS LIGHT`, and `MILLER LITE` are more expensive than `BUSCH LIGHT` and `NATURAL LIGHT`.



<br><br>

## Q1g

- Both variables `price_per_floz` and `beer_floz` are continuous.

- Describe the relationship between `price_per_floz` and `beer_floz` by brand using `seaborn`.

- Make a simple comment on the visualization result regarding how the relationship between `price_per_floz` and `beer_floz` varies by `brand`.

<br>

```{python, echo = T, eval = F}
sns.lmplot(x = "beer_floz",
           y = "price_per_floz",
           scatter_kws = {'alpha' : 0.2},
           data = beer_markets[ beer_markets['brand'] == 'BUD LIGHT' ] )


sns.lmplot(x = "beer_floz",
           y = "price_per_floz",
           scatter_kws = {'alpha' : 0.2},
           data = beer_markets[ beer_markets['brand'] == 'COORS LIGHT' ] )


sns.lmplot(x = "beer_floz",
           y = "price_per_floz",
           scatter_kws = {'alpha' : 0.2},
           data = beer_markets[ beer_markets['brand'] == 'MILLER LITE' ] )


sns.lmplot(x = "beer_floz",
           y = "price_per_floz",
           scatter_kws = {'alpha' : 0.2},
           data = beer_markets[ beer_markets['brand'] == 'BUSCH LIGHT' ] )


sns.lmplot(x = "beer_floz",
           y = "price_per_floz",
           scatter_kws = {'alpha' : 0.2},
           data = beer_markets[ beer_markets['brand'] == 'NATURAL LIGHT' ] )


# all the plots in one plot:
sns.lmplot(x = "beer_floz",
           y = "price_per_floz",
           hue = "brand",
           scatter_kws = {'alpha' : 0.2},   
           data = beer_markets)
```

```{r, echo = F, eval = T, warning = F, message= F}
ggplot(filter(beer_markets,
              brand == 'BUD LIGHT'),
       aes(x = beer_floz, y = price_per_floz)) +
  geom_point(color = 'deepskyblue4',
             alpha = .2)+
  geom_smooth(color = 'deepskyblue3',
              method = lm) +
  theme_minimal() +
  ggtitle('BUD LIGHT')
ggplot(filter(beer_markets,
              brand == 'COORS LIGHT'),
       aes(x = beer_floz, y = price_per_floz)) +
  geom_point(color = 'deepskyblue4',
             alpha = .2)+
  geom_smooth(color = 'deepskyblue3',
              method = lm) +
  theme_minimal() +
  ggtitle('COORS LIGHT')
ggplot(filter(beer_markets,
              brand == 'MILLER LITE'),
       aes(x = beer_floz, y = price_per_floz)) +
  geom_point(color = 'deepskyblue4',
             alpha = .2)+
  geom_smooth(color = 'deepskyblue3',
              method = lm) +
  theme_minimal() +
  ggtitle('MILLER LITE')
ggplot(filter(beer_markets,
              brand == 'BUSCH LIGHT'),
       aes(x = beer_floz, y = price_per_floz)) +
  geom_point(color = 'deepskyblue4',
             alpha = .2)+
  geom_smooth(color = 'deepskyblue3',
              method = lm) +
  theme_minimal() +
  ggtitle('BUSCH LIGHT')
ggplot(filter(beer_markets,
              brand == 'NATURAL LIGHT'),
       aes(x = beer_floz, y = price_per_floz)) +
  geom_point(color = 'deepskyblue4',
             alpha = .2)+
  geom_smooth(color = 'deepskyblue3',
              method = lm) +
  theme_minimal() +
  ggtitle('NATURAL LIGHT')

ggplot(beer_markets,
       aes(x = beer_floz , y = price_per_floz, 
           color = brand)) +
  geom_point(alpha = .1)+
  geom_smooth(method = lm) +
  theme_minimal()
```

<br>

- The law of demand holds in the beer market:
  - From the fitted straight line, we see the downward-sloping demand curve for each beer brand.
  
- The demand curves for `BUD LIGHT`, `COORS LIGHT`, and `MILLER LITE` are steeper than those for `BUSCH LIGHT` and `NATURAL LIGHT`.

- According to the demand curves, `BUD LIGHT`, `COORS LIGHT`, and `MILLER LITE` have larger demands than `BUSCH LIGHT` and `NATURAL LIGHT` given the same level of `price_per_floz` $\geq$ 0.0375.
  