---
title: "DANL 310 Lecture 05"
author: "Byeong-Hak Choe"
date: February 10, 2022
output: 
  ioslides_presentation:
    transition: 0
    widescreen: TRUE
    smaller: FALSE
    smart: TRUE
---
# Make a Plot

## Loading the R packages

```{r setup}
library(tidyverse)
library(skimr)   # a better summary of data frame
library(scales)  # scales for ggplot
```

## Setting the theme for ggplot {.build .faster}

```{r}
theme_set(theme_minimal()) # setting the minimal theme for ggplot
```

- Use `theme()` if you want to tweak the display of an existing theme.  

- The following lists the ggplot themes:  

theme            | theme
-----------------|-----------------
theme_grey()     | theme_gray()
theme_bw()       | theme_linedraw()
theme_light()    | theme_dark()
theme_minimal()  | theme_classic()
theme_void()     | theme_test()


## gapminder data {.build .faster}

- The `gapminder` package include the `gapminder` data frame.

```{r}
library(gapminder)
gapminder
```

## Descriptive statistics 
```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
skim(gapminder)
```

## Mappings link data to things you see 

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder)

p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p
```


## Mappings link data to things you see

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder)

p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p
```


## Mappings link data to things you see {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_point() 

```




## Build your plots layer by layer {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_smooth()
```

## Build your plots layer by layer {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_point() + geom_smooth() 
```


## Build your plots layer by layer {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_point() + geom_smooth(method = "lm") 
```


## Build your plots layer by layer {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_point() +
    geom_smooth(method = "gam") +
    scale_x_log10()
```


## Build your plots layer by layer {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_point() +
    geom_smooth(method = "gam") +
    scale_x_log10(labels = scales::dollar)
```


## Mapping aesthetics vs setting them {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p <-  ggplot(data = gapminder,
             mapping = aes(x = gdpPercap, y = lifeExp,
                           color = continent))
p
```

## Mapping aesthetics vs setting them

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, y = lifeExp,
                          color = "purple"))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

## Mapping aesthetics vs setting them 

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=5, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, y = lifeExp,
                          color = "purple"))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()

```

## Mapping aesthetics vs setting them

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p + geom_point(color = "purple") +
    geom_smooth(method = "loess") +
    scale_x_log10()

```

## Mapping aesthetics vs setting them

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=5, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p + geom_point(color = "purple") +
    geom_smooth(method = "loess") +
    scale_x_log10()

```

## Mapping aesthetics vs setting them {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE} 
p + geom_point(alpha = 0.3) +
    geom_smooth(color = "orange", se = FALSE, size = 8, method = "lm") +
    scale_x_log10()

```

## Mapping aesthetics vs setting them

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_point(alpha = 0.3) +
    geom_smooth(method = "gam") +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapminder.")
```
## Mapping aesthetics vs setting them

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=5, message=FALSE, warning=FALSE}
p + geom_point(alpha = 0.3) +
    geom_smooth(method = "gam") +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapminder.")
```

## Mapping aesthetics vs setting them

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = continent))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```


## Mapping aesthetics vs setting them

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=5, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = continent))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```


## Mapping aesthetics vs setting them 

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = continent,
                          fill = continent))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```


## Mapping aesthetics vs setting them 

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=5, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = continent,
                          fill = continent))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```


## Aesthetics can be mapped per geom

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(mapping = aes(color = continent)) +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

## Aesthetics can be mapped per geom

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=5, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(mapping = aes(color = continent)) +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

## Aesthetics can be mapped per geom {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_point(mapping = aes(color = log(pop))) +
      scale_x_continuous(trans = scales::log_trans())  # natural log scale
```

## Save your work

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5) 
```

## Save your work

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_point()
```

## Save your work

```{r, fig.width = 12, fig.height = 9, message=FALSE, warning=FALSE}
ggsave(filename = "my_figure.png")
```

## Save your work

```{r, message=FALSE, warning=FALSE}
ggsave(filename = "my_figure.pdf")
```

## Save your work

```{r, message=FALSE, warning=FALSE}
p_out <- p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
ggsave("my_figure.pdf", plot = p_out)
```

## Save your work {.build .faster}

```{r, message=FALSE, warning=FALSE}
library(here)
ggsave(here("figures", "lifexp_vs_gdp_gradient.pdf"), plot = p_out)
```

- `here()` starts at the directory where the project is created.

- Create the "figures" folder in the directory where the project is created. 

## Save your work

```{r, message=FALSE, warning=FALSE}
ggsave("lifexp_vs_gdp_gradient.pdf",
       plot = p_out, height = 8, width = 10, units = "in")
```


# Show the right number

## Grouped data and the “group” aesthetic {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = year,
                          y = gdpPercap))
p + geom_line() 
```


## Grouped data and the “group” aesthetic {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_line(aes(group=country)) 
```


## Facet to make small multiples {.build .faster}

```{r, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_line(aes(group = country)) + facet_wrap(~ continent)
```

## Facet to make small multiples

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_line(color="gray70", aes(group = country)) +
    geom_smooth(size = 1.1, method = "loess", se = FALSE) +
    scale_y_log10(labels=scales::dollar) +
    facet_wrap(~ continent, ncol = 5) +
    labs(x = "Year", y = "GDP per capita",
         title = "GDP per capita on Five Continents")
```


## Facet to make small multiples

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=5, message=FALSE, warning=FALSE}
p + geom_line(color="gray70", aes(group = country)) +
    geom_smooth(size = 1.1, method = "loess", se = FALSE) +
    scale_y_log10(labels=scales::dollar) +
    facet_wrap(~ continent, ncol = 5) +
    labs(x = "Year", y = "GDP per capita",
         title = "GDP per capita on Five Continents")
```


## Facet to make small multiples

```{r, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
p + geom_line(color="gray70", aes(group = country)) +
    geom_smooth(size = 1.1, method = "loess", se = FALSE) +
    scale_y_log10(labels=scales::dollar) +
    facet_wrap(~ continent, ncol = 5) +
    labs(x = "Year", y = "GDP per capita",
         title = "GDP per capita on Five Continents") + 
    theme(axis.text.x = element_text(angle = 45),
          axis.title.x = element_text(margin = margin(t = 25)))
```

## Facet to make small multiples

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=5, message=FALSE, warning=FALSE}
p + geom_line(color="gray70", aes(group = country)) +
    geom_smooth(size = 1.1, method = "loess", se = FALSE) +
    scale_y_log10(labels=scales::dollar) +
    facet_wrap(~ continent, ncol = 5) +
    labs(x = "Year", y = "GDP per capita",
         title = "GDP per capita on Five Continents") + 
    theme(axis.text.x = element_text(angle = 45),
          axis.title.x = element_text(margin = margin(t = 20)))
```


