---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Homework Assignment 4 - Example Answers<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```

### Loading R packages for Homework Assignment 4

```{r, eval = T, echo = T, warning = F, message = F}
library(tidyverse)
library(lubridate)
library(stargazer)
library(broom)
```

<br><br>


# Question 1
Read the data file, `bikeshare_cleaned.csv`, as the data.frame object with the name, `bikeshare`, using (1) the `read_csv()` function and (2) its URL, `https://bcdanl.github.io/data/bikeshare_cleaned.csv`.

```{r, echo = T, eval = F}
url <- 'https://bcdanl.github.io/data/bikeshare_cleaned.csv'
bikeshare <- read_csv(url)
```

Use the data.frame `bikeshare` for the rest of questions in Question 1.

<br>

- Description of variables in the data file, bikeshare_cleaned.csv
  - `cnt`: count of total rental bikes
  - `year`: year
  - `month`: month
  - `date`: date
  - `hr`: hours 
  - `wkday`: week day
  - `holiday`: holiday if `holiday` == 1, non-holiday otherwise
  - `seasons`: season
  - `weather_cond`: weather condition
  - `temp`: temperature, measured in standard deviations from average.
  - `hum`: humidity, measured in standard deviations from average.
  - `windspeed`: wind speed, measured in standard deviations from average.

<br><br>

## Q1a
Convert `year`, `month`, `wkday`, `hr`, `seasons`, and `weather_cond` into factor variables.

- Set `wkday` in order of 'sunday', 'monday', 'tuesday', etc.
- Set `seasons` in order of 'spring', 'summer', 'fall', 'winter'.

```{r, eval = T, echo = F, warning = F, message = F}
library(tidyverse)
url <- 'https://bcdanl.github.io/data/bikeshare_cleaned.csv'
bikeshare <- read_csv(url)
```

```{r, eval = T, echo = T, warning = F, message = F}
bikeshare <- bikeshare %>% 
  mutate( year = factor(year),
          month = factor(month),
          hr = factor(hr),
          weather_cond = factor(weather_cond), 
          wkday = factor(wkday, 
                         levels = c("sunday", "monday", "tuesday", "wednesday", 
                                    "thursday", "friday", "saturday") ),
          seasons = factor(seasons, 
                           levels = c("spring", "summer", "fall", "winter") ) 
          )


levels(bikeshare$year)
levels(bikeshare$month)
levels(bikeshare$hr)
levels(bikeshare$weather_cond)
levels(bikeshare$wkday)
levels(bikeshare$seasons)
```



<br><br>

## Q1b 
Divide the `bikeshare` data.frame into training and testing data.frames.

```{r, eval = T, echo = T, warning = F, message = F}
set.seed(20221121)
rn <- runif( nrow(bikeshare) )

# it is not necessary to do 50-50 split though
dtrain <- filter(bikeshare, rn >= .5)  # training data.frame
dtest <- filter(bikeshare, rn < .5)  # testing data.frame
```





<br><br>

## Q1c
Train the linear regression model using the following `formula` and the training data.frame.

Provide the summary of the regression result.


```{r, eval = F, echo = T, warning = F, message = F}
formula <- 
  cnt ~ temp + hum + windspeed + weather_cond + 
        hr + month + year
```


```{r, eval = F, echo = T, warning = F, message = F}
model <- lm(formula, data = dtrain)
summary(model)
```

```{r results = "asis", echo = TRUE, eval = TRUE}
formula <- 
  cnt ~ temp + hum + windspeed + weather_cond + 
        hr + month + year
model <- lm(formula, dtrain)
stargazer(model, type = "html")
```



<br><br>

## Q1d
Which `hr` is most strongly associated with changes in `cnt`? Interpret the beta estimate of that `hr`.

```{r, eval = T, echo = T, warning = F, message = F}
sum_model <- tidy(model) %>% 
  filter(str_detect(term, "hr")) %>% 
  arrange(-estimate) %>% 
  slice(1)
```

- All else being equal, an `hour17` relative to an `hour` 0 is associated with an increase in `cnt` by $373.85$.

- All else being equal, an `hour17` relative to an `hour` 0 is associated with an increase in `cnt` by $(373.85 - 2\times 7.62,\; 373.85 + 2\times 7.62)$.


<br><br>

## Q1e
Make a prediction on the outcome variable using the testing data.frame and the regression result from Q1c.

```{r, eval = T, echo = T, warning = F, message = F}
dtest <- dtest %>% 
  mutate( pred = predict(model, 
                         newdata = dtest) )
```


<br><br>

## Q1f
On average, are the predictions correct in the model in 3a? Are there systematic errors?

```{r, eval = T, echo = T, warning = F, message = F}

# actual vs. predicted outcome plot
ggplot( data = dtest, 
        aes(x = pred, y = cnt) ) +
  geom_point( alpha = 0.2, color = "darkgray" ) +
  geom_smooth( color = "darkblue" ) +  
  geom_abline( color = "red", linetype = 2 ) +  # y = x, perfect prediction line
  geom_vline(aes(xintercept = 0), lty = 'dotted') + 
  geom_hline(aes(yintercept = 0), lty = 'dotted')

# resitual plot
ggplot(data = dtest, 
       aes(x = pred, y = cnt - pred)) +
  geom_point(alpha = 0.2, color = "darkgray") +
  geom_smooth( color = "darkblue" ) +   
  geom_hline( aes( yintercept = 0 ),  # perfect prediction 
              color = "red", linetype = 2) + 
  geom_vline(aes(xintercept = 0), lty = 'dotted')

```

- There is a systematic error in the sense that the variance of residual increases with the value of `pred`.

- Additionally, some values of `pred` are less than zero, which may not make sense because `cnt` $\geq$ 0.
