---
title: "R Lab 3 - Pivoting"
author: "Byeong-Hak Choe"
---

```{r setup, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggthemes)
library(ggtech)
library(ggthemr)
library(hrbrthemes)
library(hexbin)
library(ggforce)
library(skimr)

knitr::opts_chunk$set(fig.width=8, fig.height=5)  

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray")))


```

## Loading R packages

```{r, eval = F, echo = T}
library(tidyverse)
library(skimr)
```

# Question 1 - Billboard Charts

The following data is for Question 1:

```{r, echo = T, eval = T, message=F}
billboard <- read_csv('https://bcdanl.github.io/data/billboard.csv')
```

```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(billboard) 
```

<br><br>


## Q1a

-   Describe how the distribution of rating varies across week 1, week 2, and week 3 using the **faceted** histogram.

```{r}

```

<br>

## Q1b

-   Which artist(s) have the most number of tracks in `billboard` data.frame?

```{r}

```

<br><br><br>

# Question 2 - Average Personal Income in NY Counties

The following data is for Question 2:

```{r, echo = T, eval = T, message=F}
ny_pincp <- read_csv('https://bcdanl.github.io/data/NY_pinc_wide.csv')
```

```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(ny_pincp) 
```

<br><br>


## Q2a
Make `ny_pincp` longer.

```{r}

```


<br><br><br>

# Question 3 - COVID-19 Cases

The following data is for Question 3:

```{r, echo = T, eval = T, message=F}
covid <- read_csv('https://bcdanl.github.io/data/covid19_cases.csv')
```

```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(covid) 
```

<br><br>


## Q3a
- Keep only the following three variables, `date`, `countriesAndTerritories`, and `cases`.

- Then make a wide-form data.frame of `covid` whose variable names are from `countriesAndTerritories` and values are from `cases`.

- Then drop the variable `date`.

<!-- Make a wide-form data.frame of `covid` whose variable names are from `countriesAndTerritories` and values are from `cases`. -->

```{r}

```

<br>

## Q3b

- Use the wide-form data.frame of `covid` to find the top 10 countries in terms of the correlation between their cases and the US case. 
  - First, keep only
  - Use `cor()`

```{r}

```
