---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Homework Assignment 2 - Example Answers<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```

### Loading R packages for Homework Assignment 2

```{r, eval = F, echo = T}
library(tidyverse)
```

<br><br>


# Question 1
Read the data file, `NY_school_enrollment_socioecon.csv`, as the data.frame object with the name, `NY_school_enrollment_socioecon`, using (1) the `read_csv()` function and (2) its URL, `https://bcdanl.github.io/data/NY_school_enrollment_socioecon.csv`.

```{r, echo = T, eval = F}
url <- 'https://bcdanl.github.io/data/NY_school_enrollment_socioecon.csv'
NY_school_enrollment_socioecon <- read_csv(url)
```



<br><br>

For description of variables in `NY_school_enrollment_socioecon`, refer to the file, `ny_school_enrollment_socioecon_description.zip`, which is in the Files section in our Canvas web-page. (I recommend you to extract the zip file, and then read the file, ny_school_enrollment_socioecon_description.csv, using Excel or Numbers.)


<br><br>

## Q1a
Look up the meaning of variables `d01_018`, `d01_021`, `d01_024`, and `d01_027`, from the file, *ny_school_enrollment_socioecon_description.csv*.

- Find the top 5 counties in terms of the value of variable `d01_018` for each year.

- Find the top 5 counties in terms of the value of variable `d01_021` for each year.

- Find the top 5 counties in terms of the value of variable `d01_024` for each year.

- Find the top 5 counties in terms of the value of variable `d01_027` for each year.

- Make a comment on the result.

```{r, eval = T, echo = F, warning = F, message = F}
library(tidyverse)

url <- 'https://bcdanl.github.io/data/NY_school_enrollment_socioecon.csv'
NY_school_enrollment_socioecon <- read_csv(url)
```

```{r, eval = T, echo = T, warning = F, message = F}
Q1a_25_34 <- NY_school_enrollment_socioecon %>% 
  select(year, county_name, d01_018) %>% 
  group_by(year) %>% 
  filter(min_rank(desc(d01_018)) <= 5) %>% 
  arrange(year, -d01_018)

Q1a_35_44 <- NY_school_enrollment_socioecon %>% 
  select(year, county_name, d01_021) %>% 
  group_by(year) %>% 
  filter(min_rank(desc(d01_021)) <= 5) %>% 
  arrange(year, -d01_021)

Q1a_45_64 <- NY_school_enrollment_socioecon %>% 
  select(year, county_name, d01_024) %>% 
  group_by(year) %>% 
  filter(min_rank(desc(d01_024)) <= 5) %>% 
  arrange(year, -d01_024)

Q1a_65_over <- NY_school_enrollment_socioecon %>% 
  select(year, county_name, d01_027) %>% 
  group_by(year) %>% 
  filter(min_rank(desc(d01_027)) <= 5) %>% 
  arrange(year, -d01_027)

# rank() and dense_rank() would also be okay.

# Here I do not provide any comments on the result.
```



<br><br>

## Q1b 
Look up the meaning of variables `c02_010`, `c04_010`, and `c06_010` from the file, *ny_school_enrollment_socioecon_description.csv*.

- Find the top 5 counties in terms of the value of variable `c02_010` for each year.

- Find the top 5 counties in terms of the value of variable `c04_010` for each year.

- Find the top 5 counties in terms of the value of variable `c06_010` for each year.

- Make a comment on the result.

```{r, eval = T, echo = T, warning = F, message = F}
Q1b_college <- NY_school_enrollment_socioecon %>% 
  select(year, county_name, c02_010) %>% 
  group_by(year) %>% 
  filter(min_rank(desc(c02_010)) <= 5) %>% 
  arrange(year, -c02_010)

Q1b_college_pub <- NY_school_enrollment_socioecon %>% 
  select(year, county_name, c04_010) %>% 
  group_by(year) %>% 
  filter(min_rank(desc(c04_010)) <= 5) %>% 
  arrange(year, -c04_010)

Q1b_college_prv <- NY_school_enrollment_socioecon %>% 
  select(year, county_name, c06_010) %>% 
  group_by(year) %>% 
  filter(min_rank(desc(c06_010)) <= 5) %>% 
  arrange(year, -c06_010)

# Here I do not provide any comments on the result.
```





<br><br>

## Q1c
Look up the meaning of variable `pincp` from the file, *ny_school_enrollment_socioecon_description.csv*.

Provide both (1) one ggplot code with `geom_line()` and (2) a couple of sentences to describe the yearly trend of `pincp` for Livingston county and top 4 counties in terms of the mean value of `pincp` over the years, `2015`-`2020`.

```{r, eval = T, echo = T, warning = F, message = F}
NY_school_enrollment_socioecon %>% 
  select(year, county_name, pincp) %>% 
  group_by(county_name) %>% 
  mutate( pincp_mean = mean(pincp) ) %>% 
  ungroup() %>% 
  filter( dense_rank(desc(pincp_mean)) <= 4 | 
            county_name == "Livingston" ) %>% 
  ggplot() +
  geom_line(aes(x = year, y = pincp, color = county_name)) +
  geom_point(aes(x = year, y = pincp, group = county_name))

# Here I do not provide any comments on the result.
```




<br><br>

## Q1d
Create the variable of the growth rate of `pincp`, where the growth rate of `pincp` is defined as:

$$ (\text{Growth rate of } \texttt{pincp} \text{ in year } \texttt{Y}) 
 = \frac{( \texttt{pincp} \text{ in year } \texttt{Y} ) - ( \texttt{pincp} \text{ in year } \texttt{Y-1} )}{(\texttt{pincp} \text{ in year } \texttt{Y-1} )} $$
Provide both (1) one ggplot code with `geom_line()` and (2) a couple of sentences to describe the yearly trend of the growth rate of `pincp` for Livingston county and top 5 counties in terms of the mean value of the growth rate of `pincp` over the years, `2016`-`2020`.

```{r, eval = T, echo = T, warning = F, message = F}
NY_school_enrollment_socioecon %>% 
  select(year, county_name, pincp) %>% 
  group_by(county_name) %>% 
  mutate(g = 100 * (pincp - lag(pincp)) / lag(pincp),
         g_mean = mean(g, na.rm = T) ) %>% 
  ungroup() %>% 
  filter( (dense_rank(desc(g_mean)) <= 4 | 
            county_name == "Livingston") & !is.na(g) ) %>% 
  ggplot() +
  geom_line(aes(x = year, y = g, color = county_name)) +
  geom_point(aes(x = year, y = g, group = county_name))

# Here I do not provide any comments on the result.

```




<br><br>

## Q1e
Look up the meaning of variable `poverty_hs` and `poverty_bachelor_or_higher` from the file, *ny_school_enrollment_socioecon_description.csv*.

- Summarize the mean value of `poverty_hs` and `poverty_bachelor_or_higher` from years `2015` to `2020` for each county.

- Provide both (1) a ggplot code with `geom_point()` and (2) a couple of sentences to describe the mean value of `poverty_hs`.

- Provide both (1) a ggplot code with `geom_point()` and (2) a couple of sentences to describe the mean value of `poverty_bachelor_or_higher`.

```{r, eval = T, echo = T, warning = F, message = F}
q1e <- NY_school_enrollment_socioecon %>% 
  group_by(county_name) %>% 
  summarise(poverty_hs = mean(poverty_hs),
            poverty_bachelor_or_higher = mean(poverty_bachelor_or_higher))

ggplot(q1e) +
  geom_point(aes(x = poverty_hs, 
                 y = county_name))
ggplot(q1e) +
  geom_point(aes(x = poverty_hs, 
                 y = reorder(county_name, poverty_hs)))

ggplot(q1e) +
  geom_point(aes(x = poverty_bachelor_or_higher, 
                 y = county_name))

ggplot(q1e) +
  geom_point(aes(x = poverty_bachelor_or_higher, 
                 y = reorder(county_name, poverty_bachelor_or_higher)))


# Here I do not provide any comments on the result.

```



<br><br><br>


# Question 2
Read the data file, `beer_markets.csv`, as the data.frame object with the name, `beer_markets`, using (1) the `read_csv()` function and (2) its URL, `https://bcdanl.github.io/data/beer_markets.csv`.

```{r, echo = T, eval = F}
beer_markets <- read_csv(
  'https://bcdanl.github.io/data/beer_markets.csv'
)
```

<br><br>

### Description of variables in the data file, `beer_markets.csv`

Each observation in `beer_markets.csv` is a household-level record for one transaction of beer.

<br>

- `hh`: an identifier of the household;
- `X_purchase_desc`: details on the purchased item;
- `quantity`: the number of items purchased;
- `brand`: Bud Light, Busch Light, Coors Light, Miller Lite, or Natural Light;
- `spent`: total dollar value of purchase;
- `beer_floz`: total volume of beer, in fluid ounces;
- `price_per_floz`: price per fl.oz. (i.e., beer spent/beer floz);
- `container`: the type of container;
- `promo`: Whether the item was promoted (coupon or otherwise);
- `market`: Scan-track market (or state if rural);
- demographic data, including gender, marital status, household income, class of work, race, education, age, the size of household, and whether or not the household has a microwave or a dishwasher.



<br><br>


## Q2a
- Find the top 5 markets in terms of the total volume of beer.
- Find the top 5 markets in terms of the total volume of *BUD LIGHT*.
- Find the top 5 markets in terms of the total volume of *BUSCH LIGHT*.
- Find the top 5 markets in terms of the total volume of *COORS LIGHT*.
- Find the top 5 markets in terms of the total volume of *MILLER LITE*.
- Find the top 5 markets in terms of the total volume of *NATURAL LIGHT*.

```{r, eval = T, echo = F, warning = F, message = F}
beer_markets <- read_csv(
  'https://bcdanl.github.io/data/beer_markets.csv'
)
```


```{r, eval = T, echo = T, warning = F, message = F}
Q2a1 <- beer_markets %>% 
   group_by(market) %>% 
   summarize(beer_floz_tot = sum(beer_floz, na.rm = T)) %>% 
   arrange(-beer_floz_tot) %>% 
   slice(1:5)

Q2a_bud <- beer_markets %>% 
  filter(brand == "BUD LIGHT") %>% 
  group_by(market) %>% 
  summarize(beer_floz_tot = sum(beer_floz, na.rm = T)) %>% 
  arrange(-beer_floz_tot) %>% 
  slice(1:5)

Q2a_busch <- beer_markets %>% 
  filter(brand == "BUSCH LIGHT") %>% 
  group_by(market) %>% 
  summarize(beer_floz_tot = sum(beer_floz, na.rm = T)) %>% 
  arrange(-beer_floz_tot) %>% 
  slice(1:5)

Q2a_coors <- beer_markets %>% 
  filter(brand == "COORS LIGHT") %>% 
  group_by(market) %>% 
  summarize(beer_floz_tot = sum(beer_floz, na.rm = T)) %>% 
  arrange(-beer_floz_tot) %>% 
  slice(1:5)

Q2a_miller <- beer_markets %>% 
  filter(brand == "MILLER LITE") %>% 
  group_by(market) %>% 
  summarize(beer_floz_tot = sum(beer_floz, na.rm = T)) %>% 
  arrange(-beer_floz_tot) %>% 
  slice(1:5)

Q2a_natural <- beer_markets %>% 
  filter(brand == "NATURAL LIGHT") %>% 
  group_by(market) %>% 
  summarize(beer_floz_tot = sum(beer_floz, na.rm = T)) %>% 
  arrange(-beer_floz_tot) %>% 
  slice(1:5)


```


<br><br>



## Q2b
- For households that purchased *BUD LIGHT*, what fraction of households did purchase only *BUD LIGHT*?

- For households that purchased *BUSCH LIGHT*, what fraction of households did purchase only *BUSCH LIGHT*?

- For households that purchased *COORS LIGHT*, what fraction of households did purchase only *COORS LIGHT*?

- For households that purchased *MILLER LITE*, what fraction of households did purchase only *MILLER LITE*?

- For households that purchased *NATURAL LIGHT*, what fraction of households did purchase only *NATURAL LIGHT*?

- Which beer brand does have the largest base of loyal consumers?



```{r, eval = T, echo = T, warning = F, message = F}
beer_markets <- beer_markets %>% 
  mutate(bud = ifelse(brand=="BUD LIGHT", 1, 0),
         busch = ifelse(brand=="BUSCH LIGHT", 1, 0),
         coors = ifelse(brand=="COORS LIGHT", 1, 0),
         miller = ifelse(brand=="MILLER LITE", 1, 0),
         natural = ifelse(brand=="NATURAL LIGHT", 1, 0) )

Q2b_bud <- beer_markets %>%
  select(hh, bud) %>%
  arrange(hh, -bud) %>%
  group_by(hh) %>%
  filter(sum(bud) > 0) %>% 
  mutate(frac_bud = sum(bud)/n(),
         loyal_bud = ifelse(frac_bud == 1, 1, 0)) %>%
  select(hh, frac_bud, loyal_bud) %>% 
  unique() %>% 
  ungroup() %>% 
  mutate(n_hh_bud = n()) %>%
  group_by(loyal_bud, n_hh_bud) %>% 
  summarise(n_obs = n()) %>% 
  ungroup() %>% 
  mutate(n_frac = n_obs/n_hh_bud )  # 0.6600816


Q2b_busch <- beer_markets %>%
  select(hh, busch) %>%
  arrange(hh, -busch) %>%
  group_by(hh) %>%
  filter(sum(busch) > 0) %>% 
  mutate(frac_busch = sum(busch)/n(),
         loyal_busch = ifelse(frac_busch == 1, 1, 0)) %>%
  select(hh, frac_busch, loyal_busch) %>% 
  unique() %>% 
  ungroup() %>% 
  mutate(n_hh_busch = n()) %>%
  group_by(loyal_busch, n_hh_busch) %>% 
  summarise(n_obs = n()) %>% 
  ungroup() %>% 
  mutate(n_frac = n_obs/n_hh_busch )  # 0.472973


Q2b_coors <- beer_markets %>%
  select(hh, coors) %>%
  arrange(hh, -coors) %>%
  group_by(hh) %>%
  filter(sum(coors) > 0) %>% 
  mutate(frac_coors = sum(coors)/n(),
         loyal_coors = ifelse(frac_coors == 1, 1, 0)) %>%
  select(hh, frac_coors, loyal_coors) %>% 
  unique() %>% 
  ungroup() %>% 
  mutate(n_hh_coors = n()) %>%
  group_by(loyal_coors, n_hh_coors) %>% 
  summarise(n_obs = n()) %>% 
  ungroup() %>% 
  mutate(n_frac = n_obs/n_hh_coors )  # 0.6390805


Q2b_miller <- beer_markets %>%
  select(hh, miller) %>%
  arrange(hh, -miller) %>%
  group_by(hh) %>%
  filter(sum(miller) > 0) %>% 
  mutate(frac_miller = sum(miller)/n(),
         loyal_miller = ifelse(frac_miller == 1, 1, 0)) %>%
  select(hh, frac_miller, loyal_miller) %>% 
  unique() %>% 
  ungroup() %>% 
  mutate(n_hh_miller = n()) %>%
  group_by(loyal_miller, n_hh_miller) %>% 
  summarise(n_obs = n()) %>% 
  ungroup() %>% 
  mutate(n_frac = n_obs/n_hh_miller )  # 0.6312989

Q2b_natural <- beer_markets %>%
  select(hh, natural) %>%
  arrange(hh, -natural) %>%
  group_by(hh) %>%
  filter(sum(natural) > 0) %>% 
  mutate(frac_natural = sum(natural)/n(),
         loyal_natural = ifelse(frac_natural == 1, 1, 0)) %>%
  select(hh, frac_natural, loyal_natural) %>% 
  unique() %>% 
  ungroup() %>% 
  mutate(n_hh_natural = n()) %>%
  group_by(loyal_natural, n_hh_natural) %>% 
  summarise(n_obs = n()) %>% 
  ungroup() %>% 
  mutate(n_frac = n_obs/n_hh_natural )  # 0.5096234

# Here I do not provide any comments on the result.

```



<br><br>

## Q2c
- Calculate the number of beer transactions for each household. 
- Calculate the fraction of each beer brand for each household.

```{r, eval = T, echo = T, warning = F, message = F}
Q2c <- beer_markets %>% 
  count(hh, brand) %>% 
  group_by(hh) %>% 
  mutate(n_tot = sum(n)) %>% 
  arrange(hh, brand) %>% 
  mutate( prop = n / n_tot )

```

