---
title: "DANL 310 Lecture 12"
author: "Byeong-Hak Choe"
date: March 24, 2022
output: 
  ioslides_presentation:
    transition: 0
    widescreen: TRUE
    smaller: FALSE
    smart: TRUE
---
# Interactive visualization


## R packages

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
# install.packages("gganimate") 
# install.packages("plotly") 
# install.packages("gifski")   # for PCs.
# install.packages("transformr")   # for manipulating polygons and paths
# install.packages("magick")  # for animation rendering
# install.packages("htmltools")  # for html rendering

library(tidyverse)
library(plotly)
library(gganimate)
library(gifski)
library(transformr)
library(magick)
library(htmltools)
theme_set(theme_bw())
```


## gganimate {.build .faster}
- Consider the relationship between *cyl* and *mpg* from **mtcars** data, which is provided by the ggplot2 package.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
data(mtcars)

p <- ggplot(data = mtcars,
       mapping = aes(x = factor(cyl), y = mpg)) +
  geom_boxplot()

p

```


## gganimate

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
data(mtcars)

p <- ggplot(data = mtcars,
       mapping = aes(x = factor(cyl), y = mpg)) +
  geom_boxplot()

p
```



## gganimate {.build .faster}
- Let us consider how the relationship between *cyl* and *mpg* vary by *gear*.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p + facet_wrap(~gear)

```


## gganimate

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p + facet_wrap(~gear)

```



## gganimate {.build .faster}

- **gganimate** turn our ggplot visualizations into moving images.

  - **gganimate** takes a ggplot figure and creates many different versions of it by changing a single parameter in the data.

```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}
my_anim <- p + 
  transition_states(gear)

my_anim
```


## gganimate

```{r, echo=F, eval=T, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}
my_anim <- p + 
  transition_states(gear)

my_anim
```


## gganimate {.build .faster}

- `transition_states()` is intended to use primarily with categorical variables. 

  - You can control the link for each transition and the amount of time spent on each of the states using the `transition_length` and `state_length` arguments.

```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}
my_anim2 <- p + 
  transition_states(gear,
    transition_length = 3,  # relative length
    state_length = 2)  # relative length

my_anim2
```



## gganimate {.build .faster}

```{r, echo=F, eval=T, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}
my_anim2 <- p + 
  transition_states(gear,
    transition_length = 3,  # relative length
    state_length = 2)  # relative length

my_anim2
```




## Cooperative Congressional Election Survey {.build .faster}

- Cooperative Congressional Election Survey (CCES) is a 50,000+ person national stratified sample survey.

- We can use `recode()` to create a new variable corresponding to other factor variable.

-  Here we summarize the number of Democrats and Republicans in a series of years.

  -  There are usually 435 seats in total in the House of Representatives.

```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}
cces <- read_csv(url("https://bcdanl.github.io/data/cces.csv"))
cces <- cces %>% 
  mutate(party = recode(dem,`1`="Democrat",`0`="Republican"))

cong_dat <- cces %>% 
  group_by(year, party) %>%
  summarise( seats =n())
```


```{r, include = F}
cces <- read_csv(url("https://bcdanl.github.io/data/cces.csv"))

cces <- cces %>% 
  mutate(party = recode(dem,`1`="Democrat",`0`="Republican"))

cong_dat <- cces %>% 
  group_by(year,party) %>%
  summarise( seats =n())
```

## Cooperative Congressional Election Survey {.build .faster}

```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}
p <- ggplot(cong_dat, 
            aes(x = year, y= seats, 
                fill = party)) +
  geom_bar(stat="identity") +
  geom_hline(yintercept = 217) +  # threshold for having a majority of seats in the house.
  scale_fill_manual(values=c("blue","red"))

p
```


## Cooperative Congressional Election Survey {.build .faster}

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(cong_dat, 
            aes(x = year, y= seats, 
                fill = party)) +
  geom_bar(stat="identity") +
  geom_hline(yintercept = 217) +  # threshold for having a majority of seats in the house.
  scale_fill_manual(values=c("blue","red"))

p
```


## gganimate {.build .faster}

- `transition_time()` is intended for time-series data like hours, minutes, days, weeks months, years, etc.

```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}

anim2 <- p + transition_time(year)

anim2
```

## gganimate 

```{r, echo=F, eval=T, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}

anim2 <- p + transition_time(year)

anim2
```



## Cooperative Congressional Election Survey {.build .faster}
- Let's consider a scatter plot of `seniority` against `all_pass`.

  - Variable `seniority` is about how long a member has been in Congress.
  
  - Variable `all_pass` is about how many bills a member passed


## Cooperative Congressional Election Survey {.build .faster}

```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}
p <- ggplot() + geom_jitter(data = filter(cces, 
                            congress==115 & party=="Democrat"),
              aes(x = seniority, y = all_pass,
                  color = party) ) +
  geom_jitter(data = filter(cces, 
                            congress==115 & party=="Republican"),
              aes(x = seniority, y = all_pass,
                  color = party) ) +
  geom_smooth(data = filter(cces, 
                            congress==115 & party=="Democrat"),
              aes(x = seniority, y = all_pass,
                  color = party) ) +
  geom_smooth(data = filter(cces, 
                            congress==115 & party=="Republican"),
              aes(x = seniority, y = all_pass,
                  color = party) ) +
  scale_color_manual(values=c("blue","red"))

p
```


## Cooperative Congressional Election Survey 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot() +
  geom_jitter(data = filter(cces, 
                            congress==115 & party=="Democrat"),
              aes(x = seniority, y = all_pass,
                  color = party) ) +
  geom_jitter(data = filter(cces, 
                            congress==115 & party=="Republican"),
              aes(x = seniority, y = all_pass,
                  color = party) ) +
  geom_smooth(data = filter(cces, 
                            congress==115 & party=="Democrat"),
              aes(x = seniority, y = all_pass,
                  color = party) ) +
  geom_smooth(data = filter(cces, 
                            congress==115 & party=="Republican"),
              aes(x = seniority, y = all_pass,
                  color = party) ) +
  scale_color_manual(values=c("blue","red"))

p
```




## gganimate {.build .faster}

- `transition_layers()` allow for building up a plot layer by layer with an animation.

```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}

anim2 <- p + transition_layers()

anim2
```



## gganimate {.build .faster}

- `enter_*()` and `exit_*()` allow for controlling the entering and exiting in `gganimate`. 

  - `*_fade()` will set the alpha value to zero making the elements fade in/out during the transition.


```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 4, message=FALSE, warning=FALSE}

anim <- ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_boxplot() +
  transition_states(factor(cyl))

# Fade-in, fade-out
anim1 <- anim +
  enter_fade() +
  exit_fade()

anim1
```


## gganimate {.build .faster}

-  `shadow_*()` allows you to retain previous or preview future frames of the animation.

  - `shadow_wake()` shows preceding frames with gradual falloff.
    - `alpha`	is for transparency modification of the wake.
    - `wrap` should the shadow wrap around, so that the first frame will get shadows from the end of the animation.

  - There are also `shadow_mark()`, `shadow_null()`, and `shadow_trail()`.
  
```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 2.5, message=FALSE, warning=FALSE}
p <- ggplot(cong_dat,
                aes(x = year, y = seats, fill = party)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 217) +
  scale_fill_manual(values = c("blue","red"))


anim3 <- p + transition_time(year) +
  shadow_wake(wake_length = 1,
              alpha = TRUE,
              wrap = TRUE)

anim3

```



## gganimate {.build .faster}
  
```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 2.5, message=FALSE, warning=FALSE}
p <- ggplot(cong_dat,
                aes(x = year, y = seats, fill = party)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 217) +
  scale_fill_manual(values = c("blue","red"))


anim3 <- p + transition_time(year) +
  shadow_wake(wake_length = 1,
              alpha = FALSE,
              wrap = FALSE)

anim3

```



## gganimate {.build .faster}

-  `anim_save()` saves saves the animation object as a `*.gif` file.
  
```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 2.5, message=FALSE, warning=FALSE}
anim_save("test.gif", animation = anim3)

```


## plotly {.build .faster}

-  `plotly` is an alternative to `ggplot` with a relatively easy to learn syntax for generating many of the same kinds of plots.

- `plotly` is mainly for the interactive figures of visualization.

  - We will focus only on `plotly::ggplotly()`. 
  
  - `ggplotly()` interacts with ggplot figures to make those figures interactive.
  
```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 2.5, message=FALSE, warning=FALSE}
# Make some noisily increasing data
dat <- data.frame(cond = rep(c("A", "B"), each = 10),
                  xvar = 1:20 + rnorm(20, sd=3),
                  yvar = 1:20 + rnorm(20, sd=3))

p <- ggplot(dat, aes(x = xvar, y = yvar)) +
  geom_point(shape=1)      # Use hollow circles
fig <- ggplotly(p)
fig
```





## plotly {.build .faster}
- Like `gganimate`, `plotly` can animate ggplot figures by creating frames of the animation based on some value in your data.

- Consider the static relationship between seniority and legislative effectiveness using CCES data.

  -  Variable `les` is a score for legislative effectiveness.
    - It measures how productive members of Congress are at making laws.
    
  
```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 2.5, message=FALSE, warning=FALSE}
p <- ggplot(cces, aes(x = seniority, y = les,
                      color = party))+
  geom_point()+
  scale_color_manual(values=c("blue","red")) +
  labs(x = "Seniority", y = "Leg. Effectiveness")

p
```



## plotly {.build .faster}
- Now consider the **dynamic** relationship between seniority and legislative effectiveness using CCES data.

  - `frame` parameter in the `aes()` function animates the figure.
  
```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 2.5, message=FALSE, warning=FALSE}
p <- ggplot(cces, aes(x = seniority, y = les,
                      color = party,
                      frame = year) ) + #this animates the figure
  geom_point()+
  scale_color_manual(values = c("blue","red")) +
  labs(x = "Seniority", y = "Leg. Effectiveness")

p1 <- ggplotly(p)
p1
```





## plotly {.build .faster}
- Now consider the **dynamic** relationship between seniority and legislative effectiveness using CCES data.

  - `ids` parameter in the `aes()`function smooths out the figure by adding the object constancy to the figure.
  
    - `ids` causes points to move from their points in the previous frame rather than being redrawn from scratch every single time.
  
  
```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 2.5, message=FALSE, warning=FALSE}
p <- ggplot(cces, aes(x = seniority, y = les,
                      color = party,
                      frame = year,
                      ids = thomas_name) ) + # this adds the object constancy
  geom_point()+
  scale_color_manual(values = c("blue","red")) +
  labs(x = "Seniority", y = "Leg. Effectiveness")

p1 <- ggplotly(p)
p1
```





## plotly {.build .faster}
- -  `htmlwidgets::saveWidget()` saves the ggplotly object as an `*.html` file.

  
```{r, echo=T, eval=F, fig.align='center', fig.height=2.5, fig.width= 2.5, message=FALSE, warning=FALSE}
htmlwidgets::saveWidget(p1, "fig.html")
```

