---
title: "DANL 200: Introduction to Data Analytics<br>Homework Assignment 2"
author: "Byeong-Hak Choe"
editor: visual
---

```{r setup, include = F}
library(tidyverse)
library(gapminder)
library(skimr)   # a better summary of data.frame
library(scales)  # scales for ggplot
library(ggthemes)  # additional ggplot themes
library(hrbrthemes) # additional ggplot themes and color pallets
library(lubridate)
library(ggridges)
theme_set(theme_minimal()) # setting the minimal theme for ggplot
# setting default chunk options
knitr::opts_chunk$set(
	eval = F,
	echo = T,
	message = FALSE,
	warning = FALSE
)
```

Load R packages you need for Homework Assignment 2

```{r}
library(tidyverse)
library(skimr)
```

# Question 1

Load the DataFrfame for Question 1.

```{r}
path <- 'https://bcdanl.github.io/data/GHG_emissions_by.csv'
ghg_emissions <- read_csv(path)
```

## Q1a

Provide both ggplot code and a simple comment to describe the yearly trend of GHG emissions for each sector.

```{r}

```

<br><br>

## Q1b

Provide both ggplot code and a simple comment to describe the yearly trend of United States of America's GHG emissions for each sector.

```{r}

```

<br><br>

## Q1c

For each party, calculate the yearly percentage change in GHG emissions for each sector.

```{r}

```

<br><br>

## Q1d

Which party has reduced GHG emissions most from 1990 level to 2017 level in terms of the percentage change in GHG emissions?

```{r}

```

<br><br>

## Q1e

Which sector has reduced GHG emissions most from 1990 level to 2017 level in terms of the percentage change in GHG emissions?

```{r}

```


<br><br><br>

# Question 2

Load the DataFrfame for Question 2.

```{r}
path <- 'https://bcdanl.github.io/data/climate_finance.csv'
climate_finance <- read_csv(path)
```

## Q2a

How many `parties` have `provided` or `disbursed` positive funding `contributions` to other countries or regions for their `adaptation` projects for every single `year` from 2011 to 2018?

```{r}

```

<br><br>

## Q2b

For each `party`, calculate the total funding `contributions` that were `disbursed` or `provided` for `mitigation` projects for each `year`.

```{r}

```

<br><br>

## Q2c

For each `party`, calculate the ratio between `adaptation` contribution and `mitigation` contribution for each type of `Status` for each `year`.

```{r}

```

<br><br>

## Q2d

Provide both ggplot code and a simple comment to describe the distribution of the ratio between `adaptation` contribution and `mitigation` contribution, which is calculated in Q2c.

```{r}

```

<br><br>

## Q2e

Provide both ggplot code and a simple comment to describe how the distribution of `Contribution` varies by `Type of support` and `Status`.

```{r}

```
