---
title: "DANL 210: Data Preparation and Management"
subtitle: "DANL 210 - Homework Assignment 1<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---



# Direction for Homework Assignment 1

- Write a Python code to answer each question.

- Make at least some simple comment (`# ...`) in each question.

- Use your working directory with the subfolder, `data`, so that the relative pathname of CSV files in the subfolder `data` is sufficient to import the CSV files.


- Import all the Python libraries you need here.
```{python, eval = F}
import pandas as pd
```



<br><br><br>

# Question 1
Consider the probability of survival in `titanic_1.csv`.
```{python, eval = F}
titanic_1 = pd.read_csv("data/titanic_1.csv")
```


## Q1a
Calculate the simple difference between the probability of survival when passengers are first-class and the probability of survival when they are not.

```{r}

```

<br><br>

## Q1b

How much does the probability of survival increase for first-class passengers relative to those who are not first-class passengers?

```{r}

```

<br><br>

## Q1c

Consider the probability of survival in `titanic_2.csv`.

```{python, eval = F}
titanic_2 = pd.read_csv("data/titanic_2.csv")
```


After stratifying on gender and age, what happens to the difference in the probabilities of survival between first-class passengers and non-first-class passengers.

Explain in your own words what stratifying on gender and age did for this difference in probabilities of survival between first-class passengers and non-first-class passengers.

```{r}

```


<br><br><br>

# Question 2

Consider the data file, NHL-1617.csv.

```{python, eval = F}
nhl1617 = pd.read_csv('data/NHL-1617.csv')
```



## Q2a
How many players have been recorded?

```{r}

```

<br><br>



## Q2b
A column points (`P`) is missing in the data. The number of points of a player is defined as the sum of his goals (`G`) and assists (`A`). Add the point column `P` to your DataFrame.

```{r}

```

<br><br>


## Q2c
Who is the top scorer in terms of points?

```{r}

```

<br><br>

## Q2d
How many Russian (non-goalie) players had some ice time in there 2016/2017 regular season? Hint: Nationality of a player can be found in "Nat". Russians are indicated by "RUS".

```{r}

```

<br><br>


## Q2e
What are their names? 

```{r}

```

<br><br>


## Q2f
Who performed best among the Russian players in terms of points (`P`)?

```{r}

```

<br><br>


## Q2g
How many points (`P`) did he have?

```{r}

```

<br><br>

## Q2h
How well did he perform in the entire league? Put differently, what was his rank in terms of points?

```{r}

```

<br><br>


## Q2i
Find the top ten scorers (in terms of points) and print them including their number of point and their respective team.

```{r}

```

<br><br>


## Q2j
What are the three countries with the most players originating from?

```{r}

```

<br><br>




<br><br><br>

# Question 3
Consider the data file, `coal_production_county.csv`.
```{python, eval = F}
coal = pd.read_csv("data/coal_production_county.csv")
```


## Q3a
For each type of mine, calculate the total coal production for each pair of state-year.

```{r}

```


<br><br>


## Q3b
Find the top 5 coal-producing states for each year.

```{r}

```


<br><br>


## Q3c
Visualize the yearly trend of the total coal production from each type of mine.

```{r}

```
