---
title: "DANL 200: Introduction to Data Analytics<br>Homework Assignment 1"
author: "Byeong-Hak Choe"
editor: visual
---

```{r setup, include = F}
library(tidyverse)
library(gapminder)
library(skimr)   # a better summary of data.frame
library(scales)  # scales for ggplot
library(ggthemes)  # additional ggplot themes
library(hrbrthemes) # additional ggplot themes and color pallets
library(lubridate)
library(ggridges)
theme_set(theme_minimal()) # setting the minimal theme for ggplot
# setting default chunk options
knitr::opts_chunk$set(
	eval = T,
	echo = F,
	message = FALSE,
	warning = FALSE
)
```

# Direction for Homework Assignment 1

## Files
- Step 1. Download the compressed ZIP file, `ny_colleges.zip`, from the Homework Assignment 1 in our Canvas. 

- Step 2. Extract the file, `ny_colleges.zip`, so that you can access the file, `NY_colleges.csv`, for Homework Assignment 1. 

<br><br>

## About `NY_colleges.csv`
- `NY_colleges.csv` is university-year level data about
  - (1) median debt of college students, 
  - (2) net price of college
  - (3) number of college students, 
  - (4) college majors
  - with a variety of segmentation.

  - A description of each variable in `NY_colleges.csv` is provided in `ny_colleges.yaml`, which we can open in RStudio or other text editors.

  - A description of values of each variable in `NY_colleges.csv` is provided in columns, `VARIABLE_NAME`, `VALUE`, and `LABEL`, in `ny_colleges_vars.xlsx`, which we can open in Microsoft Excel.

  - Using Ctrl + F (cmd + F for mac users) would be useful to find the variable description.
  - I recommend you to copy and paste the description of the variable you use in your R code to your R script with a comment (# ...).

  - From the file, `ny_colleges_vars.xlsx`, we can find that
    - `CIP**BACHL`, whose value is either `0`, `1`, or `2`, indicates whether a university offers a bachelor's degree in SOME MAJOR (**).
    - 0:	Program not offered
    - 1:	Program offered
    - 2:	Program offered through an exclusively distance-education program


<br><br><br>

# Question 1

## Q1a
Read the data file, `NY_colleges.csv`, as the data.frame object with the name, `NY_colleges`, using (1) the `read_csv()` function and (2) the absolute path name of the file `NY_colleges.csv` from your local hard disk drive in your laptop.

```{r}

```

<br><br>

## Q1b

What are the mean, median1, minimum, maximum, and standard deviation for each of the following variables,
  - (1) average net price of public institution;
  - (2) average net price of private institution;
  - (3) median debt for students who have completed;
  - (4) median debt for students who have not completed?

```{r}

```

<br><br>

## Instruction for Q1c-Q1i

From Q1c to Q1g, provide both (1) ggplot codes and (2) a couple of sentences to answer the questions.


<br><br>

## Q1c

Compare between public institutions and private institutions in terms of the distribution of average net price. 

```{r}

```


<br><br>

## Q1d

Compare between public institutions and private institutions in terms of the relationship between (1) average net price and (2) median debt for students who have completed.

```{r}

```


<br><br>

## Q1e

How does the relationship between (1) average net price and (2) median debt for students who have completed vary by levels of family income?

```{r}

```

<br><br>

## Q1f

Compare between public institutions and private institutions in terms of the proportion of high-income students.

```{r}

```

<br><br>

## Q1g

Describe how the number of students in a college varies by whether or not a college offers bachelor's degree in Business, Management, Marketing, and Related Support Services.

```{r}

```


<br><br>

## Q1h

Describe how the median debt for students who have completed varies by whether or not a college offers bachelor's degree in Business, Management, Marketing, and Related Support Services.

```{r}

```


## Q1i

Describe how the yearly trend of the median debt for students varies by levels of family income.

```{r}

```

