---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Lecture 15 Class Exercises - Answers<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```

### Loading R packages for Class Exercises

```{r, eval = T, echo = T, message=F}
library(tidyverse)
```

<br><br>


# Tweets with #climatechange
The following data is for Class Exercises:

```{r, echo = T, eval = T, message=F}
NY_CC_tweets <- read_csv(
  'https://bcdanl.github.io/data/NY_tweets_cc_no_contents.csv')
```


```{r, result = 'asis', echo = F, message = F, warning = F}
rmarkdown::paged_table(NY_CC_tweets) 
```

<br><br>

## Variable Description

- `id_user`: a unique identification number for a Twitter user whom retweeted to a tweet with *#climatechange* or *#globalwarming*.
- `id_city`: a unique identification number for a city.
- `FIPS`: a unique identification number for a county.

Each row represents an observation of a retweet to a tweet with *#climatechange* or *#globalwarming*.

Each row includes a Twitter user's geographic information at city or county levels (variables `FIPS`, `county`, `city`) as well as information about timing when a Twitter user retweeted (variables `year`, `month`, `day`, `hour`, `minute`, `second`). 


## Q1a
How many Twitter users retweeted on the date, January 1, 2017?

```{r, echo = T, eval = F}
# 1. using filter(), select(), and distinct()
Q1a <- NY_CC_tweets %>% 
  filter(year == 2017, month == 1, day == 1) %>% 
  select(id_user) %>% 
  distinct()  
# 19


# 2. using filter() and summarise() with n_distinct()
Q1a <- NY_CC_tweets %>% 
  filter(year == 2017, month == 1, day == 1) %>% 
  summarise(n_users = n_distinct(id_user))
# 19


# 3. using filter(), select() and mutate() with n() and distinct()
Q1a <- NY_CC_tweets %>% 
  filter(year == 2017, month == 1, day == 1) %>% 
  select(id_user, year, month, day) %>% 
  distinct() %>% 
  mutate( n_users = n() ) %>% 
  select( year, month, day, n_users ) %>% 
  distinct()
# 19
```

<br><br>

## Q1b
Which city is with the third highest number of retweets on the date, December 1, 2017?

```{r, echo = T, eval = F}
# 1. filter(), group_by(), 
# mutate() with n(), select(), distinct(), and arrange()
Q1b <- NY_CC_tweets %>% 
  filter(year == 2017, month == 12, day == 1) %>% 
  group_by(id_city) %>% 
  mutate(n_rt = n()) %>% 
  select(id_city, city, n_rt) %>% 
  distinct() %>% 
  arrange(-n_rt)
# brooklyn, ny 


# 2. filter(), group_by(), 
# summarise() with n(), and arrange()
Q1b <- NY_CC_tweets %>% 
  filter(year == 2017, month == 12, day == 1) %>% 
  group_by(id_city, city) %>% 
  summarize(n_rt = n()) %>% 
  arrange(-n_rt)
# brooklyn, ny 


# 3. filter(), group_by(), 
# summarise() with n(), ungroup(), and mutate() with dense_rank()
Q1b <- NY_CC_tweets %>% 
  filter(year == 2017, month == 12, day == 1) %>% 
  group_by(id_city, city) %>% 
  summarize(n_rt = n()) %>%
  ungroup() %>%  # dense_rank() calculates rankings within a group.
  mutate(ranking = dense_rank(desc(n_rt))) %>% 
  filter(ranking == 3)
# brooklyn, ny
```

```{r, result = 'asis', echo = F, message = F, warning = F}
Q1b <- NY_CC_tweets %>% 
  filter(year == 2017, month == 12, day == 1) %>% 
  group_by(id_city, city) %>% 
  summarize(n_rt = n()) %>% 
  arrange(-n_rt)
rmarkdown::paged_table(Q1b) 
```
<br><br>

## Q1c
For each year, find the top 5 Twitter users in NY state in terms of the number of retweets they made in NY state. In which city and county do these users live in?

```{r, echo = T, eval = F}
# 1. group_by(), mutate() with n(), 
# group_by(), mutate() with dense_rank(), 
# filter(), select(), distinct(), and arrange()
Q1c <- NY_CC_tweets %>% 
  group_by( id_user, year ) %>% 
  mutate( n_rt = n() ) %>% # the number of retweets for each user for each year
  group_by( year ) %>%  # dense_rank() calculates rankings within a group.
  mutate( n_rt_rank = dense_rank( desc(n_rt) ) ) %>% 
  filter( n_rt_rank <= 5 ) %>% 
  select(-(month:second)) %>% # Need to remove from 'month' to 'second' to use distinct().
  distinct() %>% 
  arrange(year, n_rt_rank)


# 2. group_by(), summarise() with n(), 
# filter(), arrange(), and head()
Q1c_2012 <- NY_CC_tweets %>% 
  group_by( id_user, year, city, county ) %>% 
  summarise( n_rt = n() ) %>% 
  filter( year == 2012 ) %>% 
  arrange(-n_rt) %>% 
  head(5)

Q1c_2013 <- NY_CC_tweets %>% 
  group_by( id_user, year, city, county ) %>% 
  summarise( n_rt = n() ) %>% 
  filter( year == 2013 ) %>% 
  arrange(-n_rt) %>% 
  head(5)

Q1c_2014 <- NY_CC_tweets %>% 
  group_by( id_user, year, city, county ) %>% 
  summarise( n_rt = n() ) %>% 
  filter( year == 2014 ) %>% 
  arrange(-n_rt) %>% 
  head(5)

Q1c_2015 <- NY_CC_tweets %>% 
  group_by( id_user, year, city, county ) %>% 
  summarise( n_rt = n() ) %>% 
  filter( year == 2015 ) %>% 
  arrange(-n_rt) %>% 
  head(5)

Q1c_2016 <- NY_CC_tweets %>% 
  group_by( id_user, year, city, county ) %>% 
  summarise( n_rt = n() ) %>% 
  filter( year == 2016 ) %>% 
  arrange(-n_rt) %>% 
  head(5)

Q1c_2017 <- NY_CC_tweets %>% 
  group_by( id_user, year, city, county ) %>% 
  summarise( n_rt = n() ) %>% 
  filter( year == 2017 ) %>% 
  arrange(-n_rt) %>% 
  head(5)


# 4. group_by(), mutate() with n(), 
# filter(), select(), unique(), arrange(), and head()
Q1c_2012 <- NY_CC_tweets %>% 
  group_by( id_user, year ) %>% 
  mutate( n_rt = n() ) %>% 
  filter( year == 2012 ) %>% 
  select(year, id_user, n_rt, city, county) %>% 
  unique() %>% 
  arrange(-n_rt) %>% 
  head(5)

Q1c_2013 <- NY_CC_tweets %>% 
  group_by( id_user, year ) %>% 
  mutate( n_rt = n() ) %>% 
  filter( year == 2013 ) %>% 
  select(year, id_user, n_rt, city, county) %>% 
  unique() %>% 
  arrange(-n_rt) %>% 
  head(5)

Q1c_2014 <- NY_CC_tweets %>% 
  group_by( id_user, year ) %>% 
  mutate( n_rt = n() ) %>% 
  filter( year == 2014 ) %>% 
  select(year, id_user, n_rt, city, county) %>% 
  unique() %>% 
  arrange(-n_rt) %>% 
  head(5)

Q1c_2015 <- NY_CC_tweets %>% 
  group_by( id_user, year ) %>% 
  mutate( n_rt = n() ) %>% 
  filter( year == 2015 ) %>% 
  select(year, id_user, n_rt, city, county) %>% 
  unique() %>% 
  arrange(-n_rt) %>% 
  head(5)

Q1c_2016 <- NY_CC_tweets %>% 
  group_by( id_user, year ) %>% 
  mutate( n_rt = n() ) %>% 
  filter( year == 2016 ) %>% 
  select(year, id_user, n_rt, city, county) %>% 
  unique() %>% 
  arrange(-n_rt) %>% 
  head(5)

Q1c_2017 <- NY_CC_tweets %>% 
  group_by( id_user, year ) %>% 
  mutate( n_rt = n() ) %>% 
  filter( year == 2017 ) %>% 
  select(year, id_user, n_rt, city, county) %>% 
  unique() %>% 
  arrange(-n_rt) %>% 
  head(5)


# cf. I do not require using functions, 
# but here I am showing you how functions make code simpler
top_users_yr <- function(yr){
  NY_CC_tweets %>% 
    group_by( id_user, year ) %>% 
    mutate( n_rt = n() ) %>% 
    filter( year == yr) %>% 
    select(year, id_user, n_rt, city, county) %>% 
    unique() %>% 
    arrange(-n_rt)
}

for (i in 2012:2017){
  print(head(top_users_yr(i), n = 5))
}
```

```{r, result = 'asis', echo = F, message = F, warning = F}
Q1c <- NY_CC_tweets %>% 
  group_by( id_user, year ) %>% 
  mutate( n_rt = n() ) %>% # the number of retweets for each user for each year
  group_by( year ) %>%  # dense_rank() calculates rankings within a group.
  mutate( n_rt_rank = dense_rank( desc(n_rt) ) ) %>% 
  filter( n_rt_rank <= 5 ) %>% 
  select(-(month:second)) %>% # Need to remove from 'month' to 'second' to use distinct().
  distinct() %>% 
  arrange(year, n_rt_rank)

rmarkdown::paged_table(Q1c) 
```
<br><br>

## Q1d
Summarize the data set into the data frame with county and month levels of retweets with the following variables:

- `FIPS`, `county`, `year`, `month`;
- `n_retweets`: the number of retweets in `year` *YYYY* and `month` *MM* from `county` `C`.

The `unique()` or `distinct()` functions can be used to keep only unique/distinct rows from a data frame.

```{r, echo = T, eval = F}
# 1. group_by() and summarize() with n()
Q1d <- NY_CC_tweets %>% 
  group_by(FIPS, county, year, month) %>% 
  summarise(n_retweets = n())

# 2. count() and rename()
Q1d <- NY_CC_tweets %>% 
  count(FIPS, county, year, month) %>% 
  rename(n_retweets = n)

# 3. group_by(), mutate() with n(), 
# select(), arrange() and distinct()
Q1d <- NY_CC_tweets %>% 
  group_by(FIPS, county, year, month) %>% 
  mutate(n_retweets = n()) %>% 
  select(FIPS, county, year, month, n_retweets) %>% 
  arrange(FIPS, county, year, month) %>% 
  distinct()

Q1d <- NY_CC_tweets %>% 
  mutate(dum = 1) %>% 
  group_by(FIPS, county, year, month) %>% 
  mutate(n_retweets = sum(dum)) %>% 
  select(FIPS, county, year, month, n_retweets) %>% 
  arrange(FIPS, county, year, month) %>% 
  distinct()

```

```{r, result = 'asis', echo = F, message = F, warning = F}
Q1d <- NY_CC_tweets %>% 
  group_by(FIPS, county, year, month) %>% 
  summarise(n_retweets = n())

rmarkdown::paged_table(Q1d) 
```
<br><br>


## Q1e
Describe the relationship between the number of retweets and `county` using ggplot. Make a simple comment on your plot.

```{r, echo = T, eval = F}
Q1e <- NY_CC_tweets %>% 
  group_by(county) %>% 
  summarize(n_retweets = n())
```


```{r, result = 'asis', echo = F, message = F, warning = F}
Q1e <- NY_CC_tweets %>% 
  group_by(county) %>% 
  summarize(n_retweets = n())

rmarkdown::paged_table(Q1e) 
```


```{r, echo = T, eval = F}
ggplot(data = Q1e) +
  geom_col(aes(x = n_retweets, y = county)) 
# Most retweets about #climatechange came from New York county.
```

```{r, echo = F, eval = T}
ggplot(data = Q1e) +
  geom_col(aes(x = n_retweets, y = county)) 
# Most retweets about #climatechange came from New York county.
```

<br><br>

```{r, echo = T, eval = F}
ggplot(data = Q1e) +
  geom_col(aes(x = n_retweets, 
               y = reorder(county, n_retweets) ))
# reorder(CATEGORICAL_VAR, CONTINUOUS_VAR) reorders 
# CATEGORICAL_VAR based on a values of CONTINUOUS_VAR
```

```{r, echo = F, eval = T}
ggplot(data = Q1e) +
  geom_col(aes(x = n_retweets, 
               y = reorder(county, n_retweets) ))
# reorder(CATEGORICAL_VAR, CONTINUOUS_VAR) reorders 
# values of CATEGORICAL_VAR 
# based on values of CONTINUOUS_VAR.
```