---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Homework Assignment 2<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```

### Loading R packages for Homework Assignment 2

```{r, eval = F, echo = T}
library(tidyverse)
```

<br><br>


# Question 1
Read the data file, `NY_school_enrollment_socioecon.csv`, as the data.frame object with the name, `NY_school_enrollment_socioecon`, using (1) the `read_csv()` function and (2) its URL, `https://bcdanl.github.io/data/NY_school_enrollment_socioecon.csv`.

```{r, echo = T, eval = F}
url <- 'https://bcdanl.github.io/data/NY_school_enrollment_socioecon.csv'
NY_school_enrollment_socioecon <- read_csv(url)
```



<br><br>

For description of variables in `NY_school_enrollment_socioecon`, refer to the file, `ny_school_enrollment_socioecon_description.zip`, which is in the Files section in our Canvas web-page. (I recommend you to extract the zip file, and then read the file, ny_school_enrollment_socioecon_description.csv, using Excel or Numbers.)


<br><br>

## Q1a
Look up the meaning of variables `d01_018`, `d01_021`, `d01_024`, and `d01_027`, from the file, *ny_school_enrollment_socioecon_description.csv*.

- Find the top 5 counties in terms of the value of variable `d01_018` for each year.

- Find the top 5 counties in terms of the value of variable `d01_021` for each year.

- Find the top 5 counties in terms of the value of variable `d01_024` for each year.

- Find the top 5 counties in terms of the value of variable `d01_027` for each year.

- Make a comment on the result.

```{r}

```




<br><br>

## Q1b 
Look up the meaning of variables `c02_010`, `c04_010`, and `c06_010` from the file, *ny_school_enrollment_socioecon_description.csv*.

- Find the top 5 counties in terms of the value of variable `c02_010` for each year.

- Find the top 5 counties in terms of the value of variable `c04_010` for each year.

- Find the top 5 counties in terms of the value of variable `c06_010` for each year.

- Make a comment on the result.

```{r}

```





<br><br>

## Q1c
Look up the meaning of variable `pincp` from the file, *ny_school_enrollment_socioecon_description.csv*.

Provide both (1) one ggplot code with `geom_line()` and (2) a couple of sentences to describe the yearly trend of `pincp` for Livingston county and top 4 counties in terms of the mean value of `pincp` over the years, `2015`-`2020`.

```{r}

```




<br><br>

## Q1d
Create the variable of the growth rate of `pincp`, where the growth rate of `pincp` is defined as:

$$ (\text{Growth rate of } \texttt{pincp} \text{ in year } \texttt{Y}) 
 = \frac{( \texttt{pincp} \text{ in year } \texttt{Y} ) - ( \texttt{pincp} \text{ in year } \texttt{Y-1} )}{(\texttt{pincp} \text{ in year } \texttt{Y-1} )} $$
Provide both (1) one ggplot code with `geom_line()` and (2) a couple of sentences to describe the yearly trend of the growth rate of `pincp` for Livingston county and top 5 counties in terms of the mean value of the growth rate of `pincp` over the years, `2016`-`2020`.

```{r}

```




<br><br>

## Q1e
Look up the meaning of variable `poverty_hs` and `poverty_bachelor_or_higher` from the file, *ny_school_enrollment_socioecon_description.csv*.

- Summarize the mean value of `poverty_hs` and `poverty_bachelor_or_higher` from years `2015` to `2020` for each county.

- Provide both (1) a ggplot code with `geom_point()` and (2) a couple of sentences to describe the mean value of `poverty_hs`.

- Provide both (1) a ggplot code with `geom_point()` and (2) a couple of sentences to describe the mean value of `poverty_bachelor_or_higher`.

```{r}

```



<br><br><br>


# Question 2
Read the data file, `beer_markets.csv`, as the data.frame object with the name, `beer_markets`, using (1) the `read_csv()` function and (2) its URL, `https://bcdanl.github.io/data/beer_markets.csv`.

```{r, echo = T, eval = F}
beer_markets <- read_csv(
  'https://bcdanl.github.io/data/beer_markets.csv'
)
```

<br><br>

### Description of variables in the data file, `beer_markets.csv`

Each observation in `beer_markets.csv` is a household-level record for one transaction of beer.

<br>

- `hh`: an identifier of the household;
- `X_purchase_desc`: details on the purchased item;
- `quantity`: the number of items purchased;
- `brand`: Bud Light, Busch Light, Coors Light, Miller Lite, or Natural Light;
- `spent`: total dollar value of purchase;
- `beer_floz`: total volume of beer, in fluid ounces;
- `price_per_floz`: price per fl.oz. (i.e., beer spent/beer floz);
- `container`: the type of container;
- `promo`: Whether the item was promoted (coupon or otherwise);
- `market`: Scan-track market (or state if rural);
- demographic data, including gender, marital status, household income, class of work, race, education, age, the size of household, and whether or not the household has a microwave or a dishwasher.



<br><br>


## Q2a
- Find the top 5 markets in terms of the total volume of beer.
- Find the top 5 markets in terms of the total volume of *BUD LIGHT*.
- Find the top 5 markets in terms of the total volume of *BUSCH LIGHT*.
- Find the top 5 markets in terms of the total volume of *COORS LIGHT*.
- Find the top 5 markets in terms of the total volume of *MILLER LITE*.
- Find the top 5 markets in terms of the total volume of *NATURAL LIGHT*.



```{r}

```


<br><br>



## Q2b
- For households that purchased *BUD LIGHT*, what fraction of households did purchase only *BUD LIGHT*?

- For households that purchased *BUSCH LIGHT*, what fraction of households did purchase only *BUSCH LIGHT*?

- For households that purchased *COORS LIGHT*, what fraction of households did purchase only *COORS LIGHT*?

- For households that purchased *MILLER LITE*, what fraction of households did purchase only *MILLER LITE*?

- For households that purchased *NATURAL LIGHT*, what fraction of households did purchase only *NATURAL LIGHT*?

- Which beer brand does have the largest base of loyal consumers?



```{r}

```



<br><br>

## Q2c
- Calculate the number of beer transactions for each household. 
- Calculate the fraction of each beer brand for each household.


