---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Midterm Exam - Example Answers<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```


### Loading R packages for the Midterm Exam

```{r, eval = T, echo = T, warning = F, message = F}
library(tidyverse)
library(skimr)
```

<br><br>


# Question 1

For Question 1, run the following function to read the `county_data.csv` file:

```{r, echo = T, eval = T, warning = F, message = F}
county_data <- read_csv(
  'https://bcdanl.github.io/data/county_data.csv'
)
```

<br>

You need to provide the absolute path name for the file, `dominick_oj_q1a.csv` to the above `read_csv()` function to read the file.

### Variable Description

- `id`: FIPS State and County code
- `name`: State or County Name
- `state`: State abbreviation
- `census_region`: Census region
 
- `pop_dens`: Population density per square mile, 2014 estimate 
- `pct_aa`: Percent African American population, 2014 estimate
 
- `pop`: Population, 2014 estimate
- `female`: Female persons, percent, 2013
- `caucasian`: Caucasian alone, percent, 2013
- `african_american`: African American alone, percent, 2013
- `travel_time`: Mean travel time to work (minutes), workers age 16+, 2009-2013
- `land_area`: Land area in square miles, 2010
- `hh_income`: Median household income, 2009-2013

- `fips`: FIPS code
- `votes_dem_2016`: Provisional count of Democratic votes in the 2016 Presidential election.
- `votes_gop_2016`: Provisional count of Republican votes in the 2016 Presidential election.
- `total_votes_2016`: Provitional count of votes cast in the 2016 Presidential election.
- `partywinner12`: Winning party, 2012 Presidental Election.



## Q1a. 
Move the column 'fips' to the first and remove column 'id'.


```{r, echo = T, eval = T, message = F, warning= F}
q1a <- county_data %>% 
  select(fips, everything()) %>% 
  select(-id)
```



<br><br>

## Q1b
Provide both (1) ggplot codes and (2) a simple comment to describe the probability distribution of `african_american`. 
```{r, echo = T, eval = T, message = F, warning= F}
ggplot(county_data) + 
  geom_density(aes(x = african_american))
```

<br>

- The distribution of the variable `african_american` is right-skewed.

- It ranges from 0.00% to 85.00%.

- The values around 2-3% are most likely for the value for `african_american`
in US counties.


<br><br>

## Q1c 
Provide both (1) ggplot codes and (2) a simple comment to describe how the probability distribution of `african_american` varies by `census_region`.
```{r, echo = T, eval = T, message = F, warning= F}
ggplot(county_data) + 
  geom_density(aes(x = african_american)) +
  facet_grid( . ~ census_region)
```
- The variable `african_american` are likely to be higher in `South` and `Northeast` of `census_region` than `Midwest` and `West` of `census_region`.


## Q1d
Provide both (1) ggplot codes and (2) a simple comment to describe the relationship between `travel_time` and `hh_income`.

```{r, echo = T, eval = T, warning = F, message= F}
ggplot(county_data,
       aes(x = travel_time, y = hh_income)) +
  geom_hex() + 
  geom_smooth() + geom_smooth(method = lm, color = 'red') 
```


- `travel_time` and `hh_income` are negatively associated overall.
  - They are negatively associated initially, and the relationship switches to be negative around the 22 minute of `travel_time`.


## Q1e
Provide both (1) ggplot codes and (2) a simple comment to describe how the relationship between `travel_time` and `hh_income` varies by `pop_dens`.
```{r, echo = T, eval = T, warning = F, message= F}
ggplot(county_data,
       aes(x = travel_time, y = hh_income)) +
  geom_point(alpha = .1) + 
  geom_smooth() + geom_smooth(method = lm, color = 'red')  +
  facet_grid(.~pop_dens)
```

- For low level of `pop_dens`, `travel_time` and `hh_income` are negatively associated.

- For high level of `pop_dens`, `travel_time` and `hh_income` are positively associated.

- The relationship between `travel_time` and `hh_income` seems to become more positive as `pop_dens` increases.


## Q1f
Provide both (1) ggplot codes and (2) a simple comment to describe how the relationship between `travel_time` and `hh_income` varies by `pop_dens` and `census_region`.
```{r, echo = T, eval = T, warning = F, message= F, fig.width=8, fig.height=12}
ggplot(county_data,
       aes(x = travel_time, y = hh_income)) +
  geom_point(alpha = .25) + 
  geom_smooth() + geom_smooth(method = lm, color = 'red')  +
  facet_grid(census_region~pop_dens)
```


- Overall, the relationship described in Q1e holds across `census_region`.


<br><br><br>

# Question 2

For Question 2, run the following R command to read the music data file.

```{r, echo = T, eval = T, warning= F, message= F}
spotify_all <- read_csv('https://bcdanl.github.io/data/spotify_all.csv')
spotify_all
```

<br>

## Q2a 
Find the ten most popular song. Who are artists for those ten most popular song?

```{r, echo = T, eval = T, warning= F, message= F}
q2a <- spotify_all %>% 
  count(artist_name, track_name) %>% 
  arrange(-n) %>% 
  head(10)

q2a
```

- This example assumes that the most popular song is the song---a combination of `artist_name` and `track_name`---that most frequently appears in the data.frame `spotify_all`. 



<br><br>

## Q2b
- Find the five most popular artist. 
- What is the most popular song for each of the five most popular artist?

```{r, echo = T, eval = T, warning= F, message= F}
q2b <- spotify_all %>% 
  group_by(artist_name) %>% 
  mutate(n_popular_artist = n()) %>% 
  ungroup() %>% 
  mutate( artist_ranking = dense_rank( desc(n_popular_artist) ) ) %>% 
  filter( artist_ranking <= 5) %>% 
  group_by(artist_name, track_name) %>% 
  mutate(n_popular_track = n()) %>% 
  group_by(artist_name) %>% 
  mutate(track_ranking = dense_rank( desc(n_popular_track) ) ) %>% 
  filter( track_ranking <= 2) %>%   # I just wanted to see the top two tracks for each artist
  select(artist_name, artist_ranking, n_popular_artist, track_name, track_ranking, n_popular_track) %>% 
  distinct() %>% 
  arrange(artist_ranking, track_ranking)

q2b
```

- This example assumes that the most popular artist is the artist---the value of `artist_name`---that most frequently appears in the data.frame `spotify_all`. 

- This example assumes that the most popular song is the song---a combination of `artist_name` and `track_name`---that most frequently appears in the data.frame `spotify_all`. 

<br><br>

## Q2c
Provide both (1) ggplot codes and (2) a couple of sentences to describe the relationship between `pos` and the ten most popular artists.

```{r, echo = T, eval = T, warning= F, message= F, fig.height= 10}
q2c <- spotify_all %>% 
  group_by(artist_name) %>% 
  mutate(n_popular_artist = n()) %>% 
  ungroup() %>% 
  mutate( artist_ranking = dense_rank( desc(n_popular_artist) ) ) %>% 
  filter( artist_ranking <= 10) 
  
# boxplot
ggplot(q2c,
       aes(x = pos, y = fct_reorder(artist_name, -artist_ranking)) ) +
  geom_boxplot() +
  stat_summary(
    fun = mean,
    color = 'red'
  )

# density
ggplot(q2c) +
  geom_density(aes(x = pos)) + 
  facet_grid(fct_reorder(artist_name, artist_ranking) ~ .  , switch = "y") +
  theme(strip.text.y.left = element_text(angle = 0))

# histogram
ggplot(q2c) +
  geom_histogram(aes(x = pos), binwidth = 1) + 
  facet_grid(fct_reorder(artist_name, artist_ranking) ~ .  , switch = "y") +
  theme(strip.text.y.left = element_text(angle = 0))
```


- The relationship between `pos` and the ten most popular artists can be described by how the distribution of `pos` varies across the ten most popular artists.

  - The distribution of `pos` does not seem to vary a lot across the ten most popular artists.
  
  - Anything noticeable can be mentioned.

<br><br>

## Q2d
Create the data frame with `pid`-`artist` level of observations with the following four variables:

- `pid`: playlist id
- `playlist_name`: the name of the playlist
- `artist`: the name of the track's primary artist, which appears only once within a playlist
- `n_artist`: the number of occurrences of artist within a playlist

```{r, echo = T, eval = T, warning= F, message= F}
q2d <- spotify_all %>% 
  count(pid, playlist_name, artist_name) %>% 
  rename(n_artist = n) %>% 
  arrange(pid, -n_artist, artist_name)

q2d
```

<br><br><br>


# Question 3

## Q3a
- Download the compressed file, `ca_housing.zip`, from the Files section in our Canvas web-site. 

- Extract the file, `ca_housing.zip`, so that you can use the file, `california_housing.csv`. 

- Read the data file, `california_housing.csv`, as the data.frame object with the name, `ca_housing`, using (1) the `read_csv()` function and (2) the absolute path name of the file, `california_housing.csv`, from your local hard disk drive in your laptop.

```{r, echo = T, eval = T, warning= F, message= F}
ca_housing <- read_csv(
  '/Users/byeong-hakchoe/Google Drive/suny-geneseo/teaching-materials/lecture-data/california_housing.csv'
)

ca_housing
```

<br><br>

## Q3b. 
Report the mean, median, minimum, maximum, and standard deviation for the variable, `medianHouseValue`, in the data.frame, ca_housing.


```{r, echo = T, eval = T, warning= F, message= F}
skim(ca_housing$medianHouseValue)
```


<br><br>

## Q3c. 
Calculate the correlation between `housingMedianAge` and `medianHouseValue`.
```{r, echo = T, eval = T, warning= F, message= F}
cor(ca_housing$housingMedianAge, ca_housing$medianHouseValue)
```
