---
title: "DANL 200: Introduction to Data Analytics"
subtitle: "DANL 200 - Homework Assignment 5 - Example Answers<br><br>" 
author: 
  - "Byeong-Hak Choe<br>"
date: "`r Sys.Date()`<br><br>"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    highlight: espresso
---
```{r klippy, echo=FALSE, include=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)

```

### Loading R packages for Homework Assignment 4

```{r, eval = T, echo = T, warning = F, message = F}
library(tidyverse)
library(lubridate)
library(stargazer)
library(broom)
```

<br><br>


# Question 1
The following data is for Question 1.

```{r, echo = T, eval = T, warning = F, message= F}
beer_markets <- read_csv(
  'https://bcdanl.github.io/data/beer_markets.csv'
)
```

<br>

- Variable description
  - `hh`: an identifier of the purchasing household;
  - `_purchase_desc`: details on the purchased item;
  - `quantity`: the number of items purchased;
  - `brand`: Bud Light, Busch Light, Coors Light, Miller Lite, or Natural Light;
  - `spent`: total dollar value of purchase;
  - `beer_floz`: total volume of beer, in fluid ounces;
  - `price_per_floz`: price per fl.oz. (i.e., beer spent/beer floz);
  - `container`: the type of container;
  - `promo`: Whether the item was promoted (coupon or otherwise);
  - `market`: Scan-track market (or state if rural);
  - demographic data, including gender, marital status, household income, class of work, race, education, age, the size of household, and whether or not the household has a microwave or a dishwasher.


<br><br>

## Q1a
Create the data.frame that keeps all the observations whose value of `container` is either 'CAN' or 'NON REFILLABLE BOTTLE' in the given data.frame, `beer_markets`.


```{r, eval = T, echo = T, warning = F, message = F}
table(beer_markets$container)
beer_markets <- filter(beer_markets, 
                       container == 'CAN' | 
                       container =='NON REFILLABLE BOTTLE')

```

<br><br>

## Q1b 
Create the data.frame that has factor-type variables of `container` and `market` using the resulting data.frame from Q1a.

```{r, eval = T, echo = T, warning = F, message = F}
beer_markets <- beer_markets %>% 
  mutate(container = factor(container),
         market = factor(market))

levels(beer_markets$container)
levels(beer_markets$market)

table(beer_markets$container)
table(beer_markets$market)
```

<br><br>

## Q1c
Create the data.frame that has a factor-type variable of `market` whose reference level is "BUFFALO-ROCHESTER" using the resulting data.frame from Q1b.

```{r, eval = T, echo = T, warning = F, message = F}
beer_markets$market <- relevel(beer_markets$market, 
                               "BUFFALO-ROCHESTER")
```


<br><br>

## Q1d
Split the resulting data.frame of Q1c into training and testing data.frames such that approximately 67% of observations in the resulting data.frame of Q1c belong to the training data.frame and the rest observations belong to the testing data.frame.

```{r, eval = T, echo = T, warning = F, message = F}
set.seed(1234)
gp <- runif(nrow(beer_markets))
train <- filter(beer_markets, gp > .33)  # training data
test <- filter(beer_markets, gp <= .33)  # test data
```


<br><br>

## Q1e
Consider the three linear regression models with the following three different formulas:

```{r, eval = T, echo = T, warning = F, message = F}
formula_1 <- log(price_per_floz) ~ market + container + brand + log(beer_floz)

formula_2 <- log(price_per_floz) ~ market + container + brand * log(beer_floz)

formula_3 <- log(price_per_floz) ~ market + container + brand * promo * log(beer_floz)

```


Conduct exploratory data analysis regarding the model with `formula_3`.

```{r, eval = T, echo = T, warning = F, message = F}
sum_train <- skimr::skim(
  select(train,
         price_per_floz, beer_floz, 
         container, brand, promo, market)
)

library(GGally)
ggpairs( select(train,
                price_per_floz, beer_floz, 
                container, brand, promo)  )

```
- More visualization should be considered.

<br><br>

## Q1f
- Train the three linear regression models using the training data.frame from Q1d with the three different formulas, `formula_1`, `formula_2`, and `formula_3`, provided in Q1e.

- What are the main goals of these linear regression models? Provide the summary of the result for each linear regression model.

- Create the data.frame of all the beta estimates for all variables and their t-statistics and p-values from each model.

```{r, results='asis', eval = T, echo = T, warning = F, message = F}
model_1 <- lm(formula_1, data = train)
model_2 <- lm(formula_2, data = train)
model_3 <- lm(formula_3, data = train)

# summary(model_1)
# summary(model_2)
# summary(model_3)

library(stargazer)
stargazer(model_1, model_2, model_3, type = 'html')
stargazer(model_1, model_2, model_3, type = 'html', 
          omit = c("market", "container", "Constant"))

b1 <- coef(model_1)
b2 <- coef(model_2)
b3 <- coef(model_3)

library(broom)
df_b1 <- tidy(model_1)
df_b2 <- tidy(model_2)
df_b3 <- tidy(model_3)
```

-  The goal of these linear regression is to answer the following questions:
  - (1) How are the explanatory variables---sales volume, promotion, and brand characteristics (e.g, consumer loyalty, beer taste)---related with the beer price?
  - (2) Given the linear relationship between the outcome and the explanatory variables, what are the predicted beer prices for unseen data?

- Intuition behind the model:
  - I choose the pricing model. The beer industry is close to oligopoly, which means that each beer company has enough market power to set the price level above its competitive level of price.
  - Each beer company sets the price of a brand's beer based on the degree of consumers' loyalty, sales volume, promotion, container type, and market characteristics (as well as various consumer characteristics).


## Q1g
Interpret the beta estimates of the following variables from the model with formula_3.
  - (1) `marketALBANY`
  - (2) `marketEXURBAN NY`
  - (3) `marketRURAL NEW YORK`
  - (4) `marketSURBURBAN NY`
  - (5) `marketSYRACUSE`
  - (6) `marketURBAN NY`          

```{r, eval = T, echo = T, warning = F, message = F}
q1g_df_b3 <- df_b3 %>% 
  filter(str_detect(term, "marketALBANY") |
           str_detect(term, "marketEXURBAN NY") |
           str_detect(term, "marketRURAL NEW YORK") |
           str_detect(term, "marketSURBURBAN NY") |
           str_detect(term, "marketSYRACUSE") |
           str_detect(term, "marketURBAN NY") ) %>% 
  mutate(sinificance = ifelse(p.value <= .1, T, F)) %>% 
  select(sinificance, everything())
  # beta for marketRURAL NEW YORK is not statistically significant
```

```{r, eval = T, echo = T, warning = F, message = F}
b3['marketALBANY']  # ***
exp(b3['marketALBANY'])  # 1.036511 
```

- (1) All else being equal, being in `ALBANY` relative to being in `BUFFALO-ROCHESTER` is associated with an increase in `price_per_floz` by 3.65%.

```{r, eval = T, echo = T, warning = F, message = F}
b3['marketEXURBAN NY']  # ***
exp(b3['marketEXURBAN NY'])
```

- (2) All else being equal, being in `EXURBAN NY` relative to being in `BUFFALO-ROCHESTER` is associated with an increase in `price_per_floz` by 19.83%.

```{r, eval = T, echo = T, warning = F, message = F}
b3['marketRURAL NEW YORK']  # there is no star
```

- (3) We cannot reject the null hypothesis that the beta coefficient for the variable `marketRURAL NEW YORK` is zero.

```{r, eval = T, echo = T, warning = F, message = F}
b3['marketSURBURBAN NY'] # ***
exp(b3['marketSURBURBAN NY'])
```

- (4) All else being equal, being in `SURBURBAN NY` relative to being in `BUFFALO-ROCHESTER` is associated with an increase in `price_per_floz` by 5.80%

```{r, eval = T, echo = T, warning = F, message = F}
b3['marketSYRACUSE'] # ***
exp(b3['marketSYRACUSE'])
1 - exp(b3['marketSYRACUSE'])
```

- (5) All else being equal, being in `SYRACUSE` relative to being in `BUFFALO-ROCHESTER` is associated with a decrease in `price_per_floz` by 4.61%.


```{r, eval = T, echo = T, warning = F, message = F}
b3['marketURBAN NY'] # ***
exp(b3['marketURBAN NY'])
```

- (6) All else being equal, being in `URBAN NY` relative to being in `BUFFALO-ROCHESTER` is associated with an increase in `price_per_floz` by 17.97%.


## Q1h
- Across the three models in Q1f, how is the percentage change in the price of beer sensitive to the percentage change in the volume of beer purchases for each brand? 

- In the model with `formula_3`, how does such sensitivity vary by `promo`?

```{r, eval = T, echo = T, warning = F, message = F}
q1h_df_b1 <- tidy(model_1) %>% 
  filter( !str_detect(term, "market"),
          !str_detect(term, "container") ) %>% 
  mutate(sinificance = ifelse(p.value <= .1, T, F)) %>% 
  select(sinificance, everything())

q1h_df_b2 <- tidy(model_2) %>% 
  filter( !str_detect(term, "market"),
          !str_detect(term, "container") ) %>% 
  mutate(sinificance = ifelse(p.value <= .1, T, F)) %>% 
  select(sinificance, everything()) 

q1h_df_b3 <- tidy(model_3) %>% 
  filter( !str_detect(term, "market"),
          !str_detect(term, "container")  ) %>% 
  mutate(sinificance = ifelse(p.value <= .1, T, F)) %>% 
  select(sinificance, everything())
```

```{r, eval = T, echo = T, warning = F, message = F}
# Model 1
q1h_df_b1
(sensitivity1_ALL <- b1['log(beer_floz)'])
```

- All else being equal, 1% increase in `beer_floz` is associated with a decrease in `price_per_floz` by 0.141%.

```{r, eval = T, echo = T, warning = F, message = F}
# Model 2
q1h_df_b2

## BUD LIGHT
(sensitivity2_BUD <- b2['log(beer_floz)'])
```

- All else being equal, an increase in `BUD LIGHT` `beer_floz` by 1% is associated with a decrease in its `price_per_floz` by 0.146%.

```{r, eval = T, echo = T, warning = F, message = F}
q1h_df_b2
## BUSCH LIGHT
(sensitivity2_BUSCH <- 
    b2['log(beer_floz)'] + b2['brandBUSCH LIGHT:log(beer_floz)'])
```

- All else being equal, an increase in `BUSCH LIGHT` `beer_floz` by 1% is associated with a decrease in its `price_per_floz` by 0.153%.

```{r, eval = T, echo = T, warning = F, message = F}
q1h_df_b2
## COORS LIGHT 
(sensitivity2_COORS <- b2['log(beer_floz)'])
```

- All else being equal, an increase in `COORS LIGHT` `beer_floz` by 1% is associated with a decrease in its `price_per_floz` by 0.146%.

- cf) There are no stars (or dots) next to `b2['brandCOORS LIGHT:log(beer_floz)']`, meaning that it is not statistically significant.

- So we cannot reject the null hypotheses that the beta coefficient for the variable, `brandCOORS LIGHT:log(beer_floz)` is zero.

```{r, eval = T, echo = T, warning = F, message = F}
q1h_df_b2
## MILLER LITE
(sensitivity2_MILLER <- 
    b2['log(beer_floz)'] + b2['brandMILLER LITE:log(beer_floz)'])
```

- All else being equal, an increase in `MILLER LITE` `beer_floz` by 1% is associated with a decrease in its `price_per_floz` by 0.167%.

```{r, eval = T, echo = T, warning = F, message = F}
q1h_df_b2
## NATURAL LIGHT
(sensitivity2_NATURAL <- 
    b2['log(beer_floz)'] + b2['brandNATURAL LIGHT:log(beer_floz)'])
```

- All else being equal, an increase in `NATURAL LIGHT` `beer_floz` by 1% is associated with a decrease in its `price_per_floz` by 0.092%.


```{r, eval = T, echo = T, warning = F, message = F}
# Model 3 
q1h_df_b3
## BUD LIGHT
(sensitivity3_BUD_NOpromo <- b3['log(beer_floz)'])
(sensitivity3_BUD_promo <- b3['log(beer_floz)'])
```
- All else being equal, an increase in `BUD LIGHT` sales by 1% with or without `promo` is associated with a decrease in its `price_per_floz` by 0.141% 

- cf) `b3['promoTRUE:log(beer_floz)']` is not statistically significant.

```{r, eval = T, echo = T, warning = F, message = F}
q1h_df_b3
## BUSCH LIGHT
(sensitivity3_BUSCH_NOpromo <- 
    b3['log(beer_floz)'] + b3['brandBUSCH LIGHT:log(beer_floz)'])

(sensitivity3_BUSCH_promo <- 
    b3['log(beer_floz)'] + b3['brandBUSCH LIGHT:log(beer_floz)'] + 
  b3['brandBUSCH LIGHT:promoTRUE:log(beer_floz)'])
```

- All else being equal, an increase in `BUSCH LIGHT` sales by 1% without `promo` is associated with a decrease in its `price_per_floz` by 0.156%.

- All else being equal, an increase in `BUSCH LIGHT` sales by 1% with `promo` is associated with a decrease in its `price_per_floz` by 0.115%.

```{r, eval = T, echo = T, warning = F, message = F}
q1h_df_b3
## COORS LIGHT 
(sensitivity3_COORS_NOpromo <- b3['log(beer_floz)'])

(sensitivity3_COORS_promo <- 
    b3['log(beer_floz)']+ b3['brandCOORS LIGHT:promoTRUE:log(beer_floz)'])
```

- All else being equal, an increase in `COORS LIGHT` sales by 1% without `promo` is associated with a decrease in its `price_per_floz` by 0.141%.

- All else being equal, an increase in `COORS LIGHT` sales by 1% with `promo` is associated with a decrease in its `price_per_floz` by 0.117%.

```{r, eval = T, echo = T, warning = F, message = F}
q1h_df_b3
## MILLER LITE
(sensitivity3_MILLER_NOpromo <- 
    b3['log(beer_floz)'] + b3['brandMILLER LITE:log(beer_floz)'])

(sensitivity3_MILLER_promo <- 
    b3['log(beer_floz)'] + b3['brandMILLER LITE:log(beer_floz)'] + 
  b3['brandMILLER LITE:promoTRUE:log(beer_floz)'])
```

- All else being equal, an increase in `MILLER LITE` sales by 1% without `promo` is associated with a decrease in its `price_per_floz` by 0.165%.

- All else being equal, an increase in `MILLER LITE` sales by 1% with `promo` is associated with a decrease in its `price_per_floz` by 0.135%.

```{r, eval = T, echo = T, warning = F, message = F}
q1h_df_b3
## NATURAL LIGHT
(sensitivity3_NATURAL_NOpromo <- 
    b3['log(beer_floz)'] + b3['brandNATURAL LIGHT:log(beer_floz)'])

(sensitivity3_NATURAL_promo <- 
    b3['log(beer_floz)'] + b3['brandNATURAL LIGHT:log(beer_floz)'] + 
  b3['brandNATURAL LIGHT:promoTRUE:log(beer_floz)'])
```

- All else being equal, an increase in `NATURAL LIGHT` sales by 1% without `promo` is associated with a decrease in its `price_per_floz` by 0.099%.

- All else being equal, an increase in `NATURAL LIGHT` sales by 1% with `promo` is associated with a decrease in its `price_per_floz` by 0.046%.

```{r, eval = T, echo = T, warning = F, message = F}
# all elasticities in one data.frame:
sensitivity <- data.frame(
  brand = c('BUD LIGHT', 'BUSCH LIGHT', 'COORS LIGHT', 'MILLER LITE', 'NATURAL LIGHT'),
  model1 = round(sensitivity1_ALL, digits = 4),
  model2 = round(c(sensitivity2_BUD, sensitivity2_BUSCH, sensitivity2_COORS, sensitivity2_MILLER, sensitivity2_NATURAL), digits = 4),
  model3_NOpromo = round(c(sensitivity3_BUD_NOpromo, sensitivity3_BUSCH_NOpromo, sensitivity3_COORS_NOpromo, sensitivity3_MILLER_NOpromo, sensitivity3_NATURAL_NOpromo), digits = 4),
  model3_promo = round(c(sensitivity3_BUD_promo, sensitivity3_BUSCH_promo, sensitivity3_COORS_promo, sensitivity3_MILLER_promo, sensitivity3_NATURAL_promo), digits = 4)
)

sensitivity
```

<br>

- In `model_1`, the volume sensitivity to changes in price does not vary by `brand`.


- In `model_2`, the volume sensitivity to changes in price varies by `brand`.
  - `MILLER LITE` is the most sensitive one.
  - `NATURAL LIGHT` is the least sensitive one.
  - `BUD LIGHT` and `COORS LIGHT` have the same sensitivity.
  
  
- In `model_3`,  the brand-specific volume sensitivity to changes in price can vary by `promo`.
  - For all beer brands except for `BUD LIGHT`, `promo == TRUE` made their demands less sensitive!
  
  

## Q1i
- For each model in Q1f, draw a residual plot.
- On average, are the prediction correct? Are there systematic errors?
- Which model do you prefer most and why?

```{r, eval = T, echo = T, warning = F, message = F}
test <- test %>% 
  mutate( pred_1 = predict(model_1, newdata = test),
          pred_2 = predict(model_2, newdata = test),
          pred_3 = predict(model_3, newdata = test) )


# residual plot
p1 <- ggplot(data = test, aes(x = pred_1, 
                        y = log(price_per_floz) - pred_1 )) +
  geom_point(color = 'grey60', alpha = .1) +
  geom_smooth(color = 'blue') +
  geom_hline(aes(yintercept = 0), color = 'red') +  # y = 0
  theme_minimal()
p1

p2 <- ggplot(data = test, aes(x = pred_2, 
                        y = log(price_per_floz) - pred_2 )) +
  geom_point(color = 'grey60', alpha = .1) +
  geom_smooth(color = 'blue') +
  geom_hline(aes(yintercept = 0), color = 'red') +  # y = 0
  theme_minimal()
p2

p3 <- ggplot(data = test, aes(x = pred_3, 
                        y = log(price_per_floz) - pred_3 )) +
  geom_point(color = 'grey60', alpha = .1) +
  geom_smooth(color = 'blue') +
  geom_hline(aes(yintercept = 0), color = 'red') +  # y = 0
  theme_minimal()
p3
```
- When answering the question---Which model do you prefer most and why?---, we can consider intuition behind formulas across models and the significance of beta estimates.

- Which model do you prefer most and why?
  - I prefer model 3.
  - The beta estimates for promo-related interaction terms except for `BUD LIGHT` are statistically significant.
  - Intuitively, promotion affects purchasing decision of consumers who are sufficiently sensitive to changes in beer volumes.
  - So, `promo` is likely to be a relevant variable to explain the variation of the beer price.

