---
title: "Teaching Presentation"
author: Byeong-Hak Choe
date: February 4, 2022
output: 
  ioslides_presentation:
    transition: 0
    widescreen: TRUE
    smaller: FALSE
    smart: TRUE
---

# Price elasticity of demand for orange juice


## Price elasticity of demand {.build .faster}

- Price elasticity of demand  measures how sensitive the quantity demanded is to its price.

&nbsp;
<center>
$\texttt{Price elasticity of demand} = \dfrac{\texttt{% Change in quantity demanded}}{\texttt{% Change in price}}$
</center>

&nbsp;

- The concept of the elasticity plays an important role in industries.

  - Marketers need to understand how consumers are sensitive to fluctuations in price.

## Loading R packages and orange juice data {.build .faster}
```{r setup, eval=TRUE, echo=TRUE, fig.align='center', out.height="75%"}
library(tidyverse)      # ggplot and more
theme_set(theme_minimal())  # minimal theme for ggplot
library(skimr)          # a better summary of data
library(stargazer)      # regression tables
library(moderndive)     # geom_parallel_slopes()
library(gridExtra)      # sub figures
```

- We will use weekly sales data for orange juice (OJ) from Dominick’s grocery stores in the 1990s.

```{r data, eval=TRUE, echo=TRUE, fig.align='center', out.height="75%"}
oj <- read.table(
  'https://bcecon.github.io/dominick_oj.csv',
  sep = ',',
  header = TRUE,
  stringsAsFactor = TRUE
)
```




# Descriptive statistics

## OJ data {.build .faster}
```{r ds, eval=TRUE, echo=TRUE, fig.align='center', out.height="75%"}
head(oj)
levels(oj$brand)
```


## Descriptive statistics {.build .faster}
```{r skim, include=FALSE}
skimmed <- skim(oj)
skimmed <- skimmed %>% 
  select(-factor.ordered, -complete_rate)
skimmed <- as_tibble(skimmed)
skimmed_chr <- skimmed[1,2:5]
skimmed_chr[1,4] <- "dominicks: 9649,  minute.maid: 9649,  tropicana: 9649"
skimmed_numeric <- skimmed[2:4,c(2:3,5:13)]
skimmed_numeric[,4:10] <- round(skimmed_numeric[,4:10], 2)
skimmed_numeric <- skimmed_numeric %>% 
  select(-factor.top_counts)
```

```{r skim_chr, eval=TRUE, echo=TRUE, fig.align='center', out.height="75%"}
knitr::kable(skimmed_chr, "simple",   # from skimr::skim() 
             col.names = c('Variable', 'n_missing', 'n_Categories', 'Counts'))  
```


## Descriptive statistics

```{r skim_n, eval=TRUE, echo=TRUE, fig.align='center', out.height="75%"}
knitr::kable(skimmed_numeric, "simple",   # from skimr::skim()
             col.names = c('Variable', 'n_missing', 'Mean', 'SD',
                           'Min', 'Q1', 'Median','Q3', 'Max', 'Hist'))
```

# Visualizing OJ data

## Distribution of OJ prices {.build .faster}

- It is better to use a logarithmic scale when percent change matters.

```{r plyr, fig.align='center', include=FALSE, out.height="75%"}
library(plyr)
mu <- ddply(oj, "brand", summarise, grp.mean=mean(log(price)))
detach("package:plyr", unload=TRUE)
library(tidyverse)
```

```{r v_price, eval=TRUE, echo=TRUE, fig.align='center', fig.height=3.75}

ggplot() + 
  geom_boxplot(data = oj, aes( y = log(price), x = brand, color = brand )) +
  geom_point(data = mu, aes(y = grp.mean, x = brand)) + 
  theme(axis.text.x = element_blank(), 
        axis.ticks = element_blank())
```


## Demand for OJ {.build .faster}
```{r v_demand, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}
ggplot(data = oj, aes( x = log(price), y = log(sales), 
                       color = brand )) + 
  geom_point( alpha = .25 ) 
```


# Regression models

## Regression model with brand dummies {.build .faster}
- The regression model for price elasticity of OJ demand can be written as:

<center>
$\log(\texttt{sales}) = \alpha_{\texttt{brand}} + \beta\log(\texttt{price}) + \epsilon$
</center>

&nbsp;

- Here $\alpha_{\texttt{brand}}$ is shorthand for a separate indicator for each OJ brand:

<center>
$\alpha_{\texttt{brand}} = \alpha_{\texttt{d}}1_{\texttt{[brand=dominicks]}}+\alpha_{\texttt{m}}1_{\texttt{[brand=minut.emaid]}}+\alpha_{\texttt{t}}1_{\texttt{[brand=tropicana]}}$
</center>

&nbsp;

- Here $\beta$ measures the price elasticity of OJ demand:

<center>
$\beta = \dfrac{\Delta\log(\texttt{sales})}{\Delta\log(\texttt{price})}\\\hspace{.25cm}  = \texttt{(% change in sales)} / \texttt{(% change in price)}$
</center>



## Run the model 

```{r reg1, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4, results="asis"}

reg_oj1 <- lm(log(sales) ~ log(price) + brand, 
             data=oj)
stargazer(reg_oj1, type = "html", omit = c("Constant"))
```


## Run the model 

```{r reg1_r, eval=TRUE, echo=FALSE, fig.align='center', fig.height=4, results="asis"}

reg_oj1 <- lm(log(sales) ~ log(price) + brand, 
             data=oj)
stargazer(reg_oj1, type = "html", omit = c("Constant"))
```


## Regression lines {.build .faster}
```{r reg1_vi, include=FALSE}
reg_lines1 <- ggplot(data = oj, aes(x = log(price), y = log(sales), 
                      color = brand )) +
  geom_point(size = .75, alpha = 0.15) +
  geom_parallel_slopes() + theme(aspect.ratio=3/4, legend.position="top")
```

```{r reg1_v, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4}

ggplot(data = oj, aes(x = log(price), y = log(sales), 
                      color = brand )) +
  geom_point(size = .75, alpha = 0.15) +
  geom_parallel_slopes()

```


## Regerssion model with interaction terms {.build .faster}

- How does consumer price sensitivity change across brands?

&nbsp;

- The new regression model can be written as:

<center>
$\log(\texttt{sales}) = \alpha_{\texttt{brand}} + \beta_{\texttt{brand}}\log(\texttt{price}) + \epsilon$
</center>

&nbsp;

- Here $\beta_{\texttt{brand}}$ is shorthand for the brand-specific price elasticity:

<center>
$\beta_{\texttt{brand}} = \hspace{.1cm}\beta_{\texttt{d}}1_{\texttt{[brand=dominicks]}}+\beta_{\texttt{m}}1_{\texttt{[brand=minutemaid]}}+\beta_{\texttt{t}}1_{\texttt{[brand=tropicana]}}\hspace{.1cm}$
</center>


## Run the model

```{r reg2, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4}

reg_oj2 <- lm(log(sales) ~ log(price)*brand, 
             data=oj)
stargazer(reg_oj1, reg_oj2, type = "html", omit = c("Constant"))
```

## Result {.smaller}

```{r reg2_r, eval=TRUE, echo=FALSE, fig.align='center', out.height="50%", results="asis"}

reg_oj2 <- lm(log(sales) ~ log(price)*brand, 
             data=oj)
stargazer(reg_oj1, reg_oj2, type = "html", omit = c("Constant"))
```


## Regression lines {.build .faster}

```{r reg2_v, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}

reg_lines2 <- ggplot(data = oj, aes(x = log(price), y = log(sales), color = brand )) +
  geom_point(size = .75, alpha = 0.15) + geom_smooth(method = lm, se=FALSE) +
   theme(aspect.ratio=3/4, legend.position="top")
grid.arrange(reg_lines1, reg_lines2, nrow = 1)   # sub-figures
```



## Advertisement effect {.build .faster}

- How does consumer price sensitivity change with advertisement?

&nbsp;

- The new regression model can be written as:

<center>
$\log(\texttt{sales}) = \alpha_{\texttt{brand,feat}} + \beta_{\texttt{brand,feat}}\log(\texttt{price}) + \epsilon$
</center>

&nbsp;

- Here $\beta_{\texttt{brand,feat}}$ measures the brand-feat-specific price elasticity:

<center>
$\beta_{\texttt{brand,feat}} = \hspace{.3cm}(\hspace{.1cm} \beta_{\texttt{d,no_ad}}1_{\texttt{[brand=dominicks]}}+\beta_{\texttt{m,no_ad}}1_{\texttt{[brand=minutemaid]}}\\\hspace{5.25cm}+\beta_{\texttt{t,no_ad}}1_{\texttt{[brand=tropicana]}}\hspace{.1cm})\times1_{\texttt{[feat=0]}}\\   \hspace{4.25cm} + (\hspace{.1cm}\beta_{\texttt{d,ad}}1_{\texttt{[brand=dominicks]}}+\beta_{\texttt{m,ad}}1_{\texttt{[brand=minutemaid]}}\\\hspace{5.25cm}+\beta_{\texttt{t,ad}}1_{\texttt{[brand=tropicana]}}\hspace{.1cm})\times1_{\texttt{[feat=1]}}$
</center>

## Run the model

```{r reg3, eval=FALSE, echo=TRUE, fig.align='center', fig.height=4}

reg_oj3 <- lm(log(sales) ~ log(price)*brand*feat, 
             data=oj)
stargazer(reg_oj1, reg_oj2, reg_oj3, type = "html", omit = c("Constant"))
```

## Result {.smaller}

```{r reg3_r, eval=TRUE, echo=FALSE, fig.align='center', out.height="50%", results="asis"}

reg_oj3 <- lm(log(sales) ~ log(price)*brand*feat, 
             data=oj)
stargazer(reg_oj1, reg_oj2, reg_oj3, type = "html", omit = c("Constant"))
```

## Regression lines {.build .faster}
```{r feat, fig.align='center', include=FALSE, out.height="75%"}
feat_names <- c(`0` = "no ad", `1` = "ad")
```

```{r reg3_v, eval=TRUE, echo=TRUE, fig.align='center', fig.height=4, message=FALSE, warning=FALSE}

ggplot(data = oj, aes(x = log(price), y = log(sales), color = brand )) +
  geom_point(size = .75, alpha = 0.15) +
  geom_smooth(method=lm) +
  facet_grid( . ~ feat, labeller = as_labeller(feat_names))

```


## Price elasticities of OJ {.build .faster}

```{r elas_t, eval=TRUE, echo=FALSE, fig.align='center', out.height="75%"}
coef1 <-  coef(reg_oj1)
coef2 <-  coef(reg_oj2)
coef3 <-  coef(reg_oj3)
elasticity <- data.frame(
  "dominicks" = c(coef1['log(price)'],coef2['log(price)'],coef3['log(price)'],coef3['log(price)']+coef3['log(price):feat']),
  "minute.maid" = c(coef1['log(price)'],
                coef2['log(price)']+coef2['log(price):brandminute.maid'], coef3['log(price)']+coef3['log(price):brandminute.maid'], coef3['log(price)']+coef3['log(price):brandminute.maid']+coef3['log(price):feat']+coef3['log(price):brandminute.maid:feat']),
  "tropicana" = c(coef1['log(price)'],
                coef2['log(price)']+coef2['log(price):brandtropicana'], coef3['log(price)']+coef3['log(price):brandtropicana'], coef3['log(price)']+coef3['log(price):brandtropicana']+coef3['log(price):feat']+coef3['log(price):brandtropicana:feat'])
  )
row.names(elasticity) <- c("Model 1", "Model 2", "Model 3 w/ no ad", "Model 3 w/ ad")

knitr::kable(round(elasticity,1), "simple")   
```

&nbsp;

- Which model is better?

  - Which model is more realistic?
  
  - Why does being featured always lead to more price sensitivity?



## The relationship between ad and brand

```{r ad_brand_tb, fig.align='center', include=FALSE, out.height="75%"}
by_brand <- group_by(oj, brand, feat)
salestable <- summarize(by_brand, 
                        sales_tot = 
                          sum(sales, na.rm = TRUE)
                        )
salestable <- salestable %>% group_by(feat) %>% 
  mutate(sales_prop = sales_tot/sum(sales_tot))
```

```{r ad_barnd_v, eval=TRUE, echo=TRUE, fig.align='center', fig.height=3.5}
ggplot(data = salestable) +
  geom_bar(mapping = aes(x = factor(feat), y = sales_prop, 
                         fill = brand), stat = "identity") + 
  xlab("feat") + ylab("proportion of sales") +
  scale_x_discrete(breaks=c(0, 1), labels=c("no ad", "ad"))
```



## Residual plot and model fit {.build .faster}

- Which model is better?

  - On average, are the predictions correct?

  - Are there systematic errors in prediction?
  
- Predicted log of sales can be calculated as follows:
```{r pred, eval=TRUE, echo=TRUE, fig.align='center', fig.height=3.5}
oj$predLogSales1 <- predict(reg_oj1, newdata = oj)  # Model 1
oj$predLogSales2 <- predict(reg_oj2, newdata = oj)  # Model 2
oj$predLogSales3 <- predict(reg_oj3, newdata = oj)  # Model 3
```

## Residual plot for model 1

```{r resid1_v, eval=TRUE, echo=TRUE, fig.align='center', fig.height=3.775, message=FALSE, warning=FALSE}
ggplot(data = oj, aes(x = predLogSales1, y = log(sales) - predLogSales1 )) +
  geom_point(alpha = 0.2, color = "orange") + geom_smooth(color = "darkblue") +
  geom_line(aes(x = predLogSales1, y = 0), color = "red", linetype = 2) +
  xlab("prediction") + ylab("residual error (actual - prediction)")
```



## Residual plot for model 2 

```{r resid2_v, eval=TRUE, echo=TRUE, fig.align='center', fig.height=3.775, message=FALSE, warning=FALSE}
ggplot(data = oj, aes(x = predLogSales2, y = log(sales) - predLogSales2 )) +
  geom_point(alpha = 0.2, color = "orange") + geom_smooth(color = "darkblue") +
  geom_line(aes(x = predLogSales2, y = 0), color = "red", linetype = 2) +
  xlab("prediction") + ylab("residual error (actual - prediction)")
```



## Residual plot for model 3 

```{r resid3_v, eval=TRUE, echo=TRUE, fig.align='center', fig.height=3.775, message=FALSE, warning=FALSE}
ggplot(data = oj, aes(x = predLogSales3, y = log(sales) - predLogSales3 )) +
  geom_point(alpha = 0.2, color = "orange") + geom_smooth(color = "darkblue") +
  geom_line(aes(x = predLogSales3, y = 0), color = "red", linetype = 2) +
  xlab("prediction") + ylab("residual error (actual - prediction)")
```

