---
title: "DANL 310 Lecture 17"
subtitle: "R Markdown Basics and Tables"
author: "Byeong-Hak Choe"
institute: "SUNY Geneseo"
date: "April 14, 2022"
output:
  xaringan::moon_reader:
    css: 
      - default
      - css/nhsr.css
      - css/nhsr-fonts.css
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: googlecode
      highlightLines: true
      highlightLanguage: ["r"]
      countIncrementalSlides: false
      ratio: "16:9"
    includes:
      after_body: [css/insert-logo.html]
---

```{r setup, include = FALSE}
library(knitr)
library(tidyverse)
library(NHSRtheme)
# set default options
opts_chunk$set(echo = FALSE,
               fig.width = 7.252,
               fig.height = 4,
               comment = "#",
               dpi = 300)

knitr::knit_engines$set("markdown")

xaringanExtra::use_tile_view()
xaringanExtra::use_panelset()
xaringanExtra::use_clipboard()
xaringanExtra::use_webcam()
xaringanExtra::use_broadcast()
xaringanExtra::use_share_again()
xaringanExtra::style_share_again(
  share_buttons = c("twitter", "linkedin", "pocket")
)


xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         #<<
  mute_unhighlighted_code = TRUE  #<<
)

# uncomment the following lines if you want to use the NHS-R theme colours by default
# scale_fill_continuous <- partial(scale_fill_nhs, discrete = FALSE)
# scale_fill_discrete <- partial(scale_fill_nhs, discrete = TRUE)
# scale_colour_continuous <- partial(scale_colour_nhs, discrete = FALSE)
# scale_colour_discrete <- partial(scale_colour_nhs, discrete = TRUE)
```

class: title-slide, left, bottom

# `r rmarkdown::metadata$title`
----
## **`r rmarkdown::metadata$subtitle`**
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$date`


---
class: inverse, center, middle

# R Markdown Basics
<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
---


# Code chunks and inline R code

- You should define the object for the number in the code chunck, so that you can use that number in the narratives of the project document.


### <p style="color:#327ec6">R Markdown Code</p>

```r
library(tidyverse)
oj <- read_csv('https://bcecon.github.io/dominick_oj.csv')
nvars <- format(round(ncol(oj), 0), 
                nsmall=0, 
                big.mark=",") 
nobs <- format(round(nrow(oj), 0), 
                nsmall=0, 
                big.mark=",") 
```

The number of variables is <code>&grave;r nvars&grave;</code>; the number of observations is <code>&grave;r nobs&grave;</code>.



### <p style="color:#327ec6">Output</p>

```{r, message=F}
library(tidyverse)
oj <- read_csv(
  'https://bcecon.github.io/dominick_oj.csv')
nvars <- format(round(ncol(oj), 0), 
                nsmall=0, big.mark=",") 
nobs <- format(round(nrow(oj), 0), 
               nsmall=0, big.mark=",") 
```

The number of variables is `r nvars`; the number of observations is `r nobs`.

Roses are <span style="color: red;">red</span>, violets are <span style="color: blue;">blue</span>.

---


# Font color

- The Markdown syntax has no built-in method for changing text colors. We can use HTML syntax to change the formatting of words:



### <p style="color:#327ec6">R Markdown Code</p>

```r
Roses are <span style="color: red;">red</span>, violets are <span style="color: blue;">blue</span>.
```

### <p style="color:#327ec6">Output</p>

Roses are <span style="color: red;">red</span>, violets are <span style="color: red;">blue</span>.

---


# Spacing

- To add a space between paragraphs or lines, you can use `<br>`.


.pull-left[
### <p style="color:#327ec6">R Markdown Code</p>

```r
One
<br>

Two
```
]

.pull-right[
### <p style="color:#327ec6">Output</p>

One
<br>

Two
]


---

# YAML

- Here is an example of the YAML metadata for an HTML document from R Markdown:

.pull-left[
### <p style="color:#327ec6">R Markdown YAML Example</p>

```r
title: "TITLE_OF_YOUR_PROJECT"
subtitle: "SUBTITLE_OF_YOUR_PROJECT<br><br>" 
author: 
  - "AUTHOR_1"
  - "AUTHOR_2"
  - <br>
date: "`r Sys.Date()`<br><br>"

output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: united
    highlight: tango
```
]

.pull-right[
### <p style="color:#327ec6"> `theme` </p>
- `theme` specifies the Bootstrap theme to use for the page (themes are drawn from the Bootswatch theme library). 
  - Valid themes include `default`, `bootstrap`, `cerulean`, `cosmo`, `darkly`, `flatly`, `journal`, `lumen`, `paper`, `readable`, `sandstone`, `simplex`, `spacelab`, `united`, and `yeti.`

  - Pass `null` for no theme (in this case you can use the `css` parameter to add your own styles).



]


---


# YAML

- Here is an example of the YAML metadata for an HTML document from R Markdown:

.pull-left[
### <p style="color:#327ec6">R Markdown YAML Example</p>

```r
title: "TITLE_OF_YOUR_PROJECT"
subtitle: "SUBTITLE_OF_YOUR_PROJECT<br><br>" 
author: 
  - "AUTHOR_1"
  - "AUTHOR_2"
  - <br>
date: "`r Sys.Date()`<br><br>"

output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: united
    highlight: tango
```
]

.pull-right[
### <p style="color:#327ec6"> `highlight` </p>

- `highlight` specifies the syntax highlighting style. 

  - Supported styles include `default`, `tango`, `pygments`, `kate`, `monochrome`, `espresso`, `zenburn`, `haddock`, `breezedark`, and `textmate.` 
  
  - Pass `null` to prevent syntax highlighting.

]



---

class: inverse, center, middle

# Tables for Descriptive Statisics
<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
