---
title: "DANL 310 Lecture 10"
author: "Byeong-Hak Choe"
date: March 1, 2022
output: 
  ioslides_presentation:
    transition: 0
    widescreen: TRUE
    smaller: FALSE
    smart: TRUE
---
# Draw maps


## R packages

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
library(tidyverse)
library(socviz)
library(ggthemes)
theme_set(theme_minimal())
```

## Map U.S. state-level data {.build .faster}

- The `socviz::election` dataset has various measures of the vote and vote shares by state. 


```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
election %>% select(state, total_vote,
                    r_points, pct_trump, party, census) %>%
    sample_n(5)

```

- We don’t have to represent spatial data spatially.


## Map U.S. state-level data {.build .faster}

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
party_colors <- c("#2E74C0", "#CB454A")  # Hex color codes for Dem Blue and Rep Red
p0 <- ggplot(data = subset(election, st %nin% "DC"),
             mapping = aes(x = r_points,
                           y = reorder(state, r_points),
                           color = party))

p1 <- p0 + geom_vline(xintercept = 0, color = "gray30") +
    geom_point(size = 2)

p2 <- p1 + scale_color_manual(values = party_colors)

p3 <- p2 + scale_x_continuous(breaks = c(-30, -20, -10, 0, 10, 20, 30, 40),
                              labels = c("30\n (Clinton)", "20", "10", "0",
                                         "10", "20", "30", "40\n(Trump)"))

p3 + facet_wrap(~ census, ncol=1, scales="free_y") +
     guides(color = "none") + labs(x = "Point Margin", y = "") +
     theme(axis.text=element_text(size=8))
```


## Map U.S. state-level data

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
# Hex color codes for Dem Blue and Rep Red
party_colors <- c("#2E74C0", "#CB454A") 

p0 <- ggplot(data = subset(election, st %nin% "DC"),
             mapping = aes(x = r_points,
                           y = reorder(state, r_points),
                           color = party))

p1 <- p0 + geom_vline(xintercept = 0, color = "gray30") +
    geom_point(size = 2)

p2 <- p1 + scale_color_manual(values = party_colors)

p3 <- p2 + scale_x_continuous(breaks = c(-30, -20, -10, 0, 10, 20, 30, 40),
                              labels = c("30\n (Clinton)", "20", "10", "0",
                                         "10", "20", "30", "40\n(Trump)"))

p3 + facet_wrap(~ census, ncol=1, scales="free_y") +
     guides(color = "none") + labs(x = "Point Margin", y = "") +
     theme(axis.text=element_text(size=8))

```

## Map U.S. state-level data {.build .faster}
- Let us get a data frame of the US state map

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
us_states <- map_data("state")
head(us_states)
dim(us_states)
```



## Map U.S. state-level data {.build .faster}
```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            mapping = aes(x = long, y = lat,
                          group = group))

p + geom_polygon(fill = "white", color = "black")
```

- A map is a set of lines drawn in the right order on a grid.


## Map U.S. state-level data 
```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            mapping = aes(x = long, y = lat,
                          group = group))

p + geom_polygon(fill = "white", color = "black")
```



## Map U.S. state-level data {.build .faster}

- Let's `fill` the map.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            aes(x = long, y = lat,
                group = group, fill = region))

p + geom_polygon(color = "gray90", size = 0.1) + guides(fill = FALSE)
```



## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            aes(x = long, y = lat,
                group = group, fill = region))

p + geom_polygon(color = "gray90", size = 0.1) + guides(fill = FALSE)
```




## Map U.S. state-level data {.build .faster}
- Let’s deal with the projection. 
  - By default, the map is plotted using the venerable *Mercator* projection.
  - The *Albers* projection is nicer. 
 
```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            mapping = aes(x = long, y = lat,
                          group = group, fill = region))

p + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    guides(fill = FALSE)
```

- We can transform the default projection used by `geom_polygon()`, via the `coord_map()` function.

  - The *Albers* projection requires two latitude parameters, `lat0` and `lat1`. 

## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            mapping = aes(x = long, y = lat,
                          group = group, fill = region))

p + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    guides(fill = FALSE)
```





## Map U.S. state-level data {.build .faster}
- Let’s get the `election` data on to the map 
 
```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
election$region <- tolower(election$state)
us_states_elec <- left_join(us_states, election)
```

- In the map data, `us_states`, the state names (in a variable named `region`) are in lower case. 

- Here we can create a variable in the `election` data frame to correspond to this, using the `tolower()` function to convert the state names.

- It is important to know your data and variables well enough to check that they have merged properly.

  - FIPS code is useful in joining the data.



## Map U.S. state-level data {.build .faster}
 
```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
            aes(x = long, y = lat,
                group = group, fill = party))

p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

```

- We use our `party` colors for the `fill`.


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
            aes(x = long, y = lat,
                group = group, fill = party))

p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 
```




## Map U.S. state-level data {.build .faster}
 
```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p2 <- p1 + scale_fill_manual(values = party_colors) +
    labs(title = "Election Results 2016", fill = NULL)

p2 + theme_map() 


```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p2 <- p1 + scale_fill_manual(values = party_colors) +
    labs(title = "Election Results 2016", fill = NULL)

p2 + theme_map() 

```


## Map U.S. state-level data {.build .faster}
- To the `fill` aethetic, let’s try a continuous measure, such as the percentage of the vote received by Donald Trump (`pct_trump`). 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
             mapping = aes(x = long, y = lat, group = group, fill = pct_trump))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p1 + labs(title = "Trump vote") + theme_map() + labs(fill = "Percent")

```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
             mapping = aes(x = long, y = lat, group = group, fill = pct_trump))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p1 + labs(title = "Trump vote") + theme_map() + labs(fill = "Percent")

```




## Map U.S. state-level data {.build .faster}

- Blue is not the color we want here.

- The color gradient runs in the wrong direction.

- Let's fix these problems using `scale_fill_gradient()`:

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p2 <- p1 + scale_fill_gradient(low = "white", high = "#CB454A") +
        labs(title = "Trump vote") 
p2 + theme_map() + labs(fill = "Percent")

```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p2 <- p1 + scale_fill_gradient(low = "white", high = "#CB454A") +
        labs(title = "Trump vote") 
p2 + theme_map() + labs(fill = "Percent")

```


## Map U.S. state-level data {.build .faster}

- For election results, we might prefer a gradient that diverges from a midpoint. 
  - The `scale_gradient2()` function gives us a blue-red spectrum that passes through white by default. 
  - We can also re-specify the mid-level color along with the `high` and `low` colors. 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
             mapping = aes(x = long, y = lat, group = group, fill = d_points))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p2 <- p1 + scale_fill_gradient2() + labs(title = "Winning margins") 
p2 + theme_map() + labs(fill = "Percent")

```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
             mapping = aes(x = long, y = lat, group = group, fill = d_points))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p2 <- p1 + scale_fill_gradient2()

p2 + theme_map() + labs(fill = "Percent", title = "Winning margins")

```


## Map U.S. state-level data {.build .faster}

- From the `scale_gradient2()` function, we can also re-specify the mid-level color along with the `high` and `low` colors. 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p3 <- p1 + scale_fill_gradient2(low = "red", 
                                mid = scales::muted("purple"),
                                high = "blue", 
                                breaks = c(-25, 0, 25, 50, 75)) 
p3 + theme_map() + labs(fill = "Percent", title = "Winning margins")

```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p3 <- p1 + scale_fill_gradient2(low = "red", 
                                mid = scales::muted("purple"),
                                high = "blue", 
                                breaks = c(-25, 0, 25, 50, 75)) 
p3 + theme_map() + labs(fill = "Percent", title = "Winning margins")

```





## Map U.S. state-level data {.build .faster}

- If you take a look at the gradient scale for this first “purple America” map, `p3`, you’ll see that it extends very high on the Blue side.
  - This is because Washington DC is included in the data.
  
- If we omit Washington DC, we’ll see that our color scale shifts.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = subset(us_states_elec,
                           region %nin% "district of columbia"),
             aes(x = long, y = lat, group = group, fill = d_points))

p3 <- p1 + scale_fill_gradient2(low = "red", 
                                mid = scales::muted("purple"),
                                high = "blue", 
                                breaks = c(-25, 0, 25, 50, 75)) 
p3 + theme_map() + labs(fill = "Percent", title = "Winning margins", caption = "DC is omitted.")
```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = subset(us_states_elec,
                           region %nin% "district of columbia"),
             aes(x = long, y = lat, group = group, fill = d_points))

p3 <- p1 + scale_fill_gradient2(low = "red", 
                                mid = scales::muted("purple"),
                                high = "blue", 
                                breaks = c(-25, 0, 25, 50, 75)) 
p3 + theme_map() + labs(fill = "Percent", title = "Winning margins", caption = "DC is omitted.")

```



## America’s ur-choropleths

- Choropleth maps display divided geographical areas or regions that are colored, shaded or patterned in relation to a data variable.

- County-level US choropleth maps can be aesthetically pleasing, because of the added detail they bring to a national map. 

- The county-level datasets (`county_map` and `county_data`) are included in the `socviz` library.

  - The county map data frame, `county_map`, has been processed a little in order to transform it to an *Albers* projection, and also to relocate (and re-scale) Alaska and Hawaii.




## America’s ur-choropleths {.build .faster}

```{r skim, include=FALSE}
county_map %>% sample_n(5)

county_data %>%
    select(id, name, state, pop_dens, pct_black) %>%
    sample_n(5)

county_full <- left_join(county_map, county_data, by = "id")
```

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
county_map %>% sample_n(5)

county_data %>%
    select(id, name, state, pop_dens, pct_black) %>%
    sample_n(5)

county_full <- left_join(county_map, county_data, by = "id")
```

- The `id` field is the FIPS code for the county.

- `pop_dens` is population density. 

- `pct_black` is percent of African-American population.

- We merge the data frames using the shared FIPS `id` column.



## Map U.S. county-level data {.build .faster} 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = county_full,
            mapping = aes(x = long, y = lat,
                          fill = pop_dens, 
                          group = group))
p1 <- p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()
p1
```

-  `p1` object produces a legible map, but by default it chooses an unordered categorical layout. 

  - This is because the `pop_dens` variable is not ordered.
  - `pop_dens` is an un-ordered discrete variable.
  - `summary(county_full$pop_dens)`
  
- The use of `coord_equal()` makes sure that the relative scale of our map does not change even if we alter the overall dimensions of the plot.


## Map U.S. county-level data

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = county_full,
            mapping = aes(x = long, y = lat,
                          fill = pop_dens, 
                          group = group))
p1 <- p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()
p1
```


## Map U.S. county-level data {.build .faster} 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p2 <- p1 + scale_fill_brewer(palette = "Blues",
                             labels = c("0-10", "10-50", "50-100", "100-500","500-1,000", "1,000-5,000", ">5,000"))

p2 + labs(fill = "Population per\nsquare mile") +
    theme_map() +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")
```

- We can manually supply the right sort of scale using the `scale_fill_brewer()` function, together with a nicer set of labels. 

- We can also use the `guides()` function to make sure each element of the key in the *legend* appears on the same row. 



## Map U.S. county-level data

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p2 <- p1 + scale_fill_brewer(palette = "Blues",
                             labels = c("0-10", "10-50", "50-100", "100-500","500-1,000", "1,000-5,000", ">5,000"))

p2 + labs(fill = "Population per\nsquare mile") +
    theme_map() +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")
```




## Map U.S. county-level data {.build .faster} 

- We can now do exactly the same thing for our map of percent African-American population by county. 
  - `pct_black` is an un-ordered factor variable.
  - `summary(county_full$pct_black)`

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = county_full,
            mapping = aes(x = long, y = lat, fill = pct_black, 
                          group = group))
p1 <- p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

p2 <- p1 + scale_fill_brewer(palette="Greens")

p2 + labs(fill = "US Population, Percent Black") +
    guides(fill = guide_legend(nrow = 1)) + 
    theme_map() + theme(legend.position = "bottom")

```


## Map U.S. county-level data

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = county_full,
            mapping = aes(x = long, y = lat, fill = pct_black, 
                          group = group))
p1 <- p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()
p2 <- p1 + scale_fill_brewer(palette="Greens")

p2 + labs(fill = "US Population, Percent Black") +
    guides(fill = guide_legend(nrow = 1)) + 
    theme_map() + theme(legend.position = "bottom")

```







## Map U.S. county-level data {.build .faster} 

-  Let's draw two new county-level choropleths.
    
  - We have a `pop_dens6` variable that divides the population density into six categories.

  - We will map the color scale to the value of variable.

```{r, include=FALSE}
orange_pal <- RColorBrewer::brewer.pal(n = 6, name = "Oranges")
orange_pal

orange_rev <- rev(orange_pal)
orange_rev
```

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
orange_pal <- RColorBrewer::brewer.pal(n = 6, name = "Oranges")
orange_pal

orange_rev <- rev(orange_pal)
orange_rev
```


- We use the `RColorBrewer::brewer.pal()` function to manually create two palettes.
  - The `brewer.pal()` function produces evenly-spaced color schemes. 

- We use the `rev()` function to reverse the order of a color vector.
  
  
## Map U.S. county-level data {.build .faster} 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
pop_p <- ggplot(data = county_full,
            mapping = aes(x = long, y = lat,
                          fill = pop_dens6, 
                          group = group))

pop_p1 <- pop_p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

pop_p2 <- pop_p1 + scale_fill_manual(values = orange_pal)

pop_p2 + labs(title = "Population Density",
              fill = "People per square mile") +
    theme_map() + theme(legend.position = "bottom")

```



## Map U.S. county-level data

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
pop_p <- ggplot(data = county_full,
            mapping = aes(x = long, y = lat,
                          fill = pop_dens6, 
                          group = group))

pop_p1 <- pop_p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

pop_p2 <- pop_p1 + scale_fill_manual(values = orange_pal)

pop_p2 + labs(title = "Population Density",
              fill = "People per square mile") +
    theme_map() + theme(legend.position = "bottom")

```



  
## Map U.S. county-level data {.build .faster} 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}

pop_p2_rev <- pop_p1 + scale_fill_manual(values = orange_rev)

pop_p2_rev + labs(title = "Reverse-coded Population Density",
              fill = "People per square mile") +
    theme_map() + theme(legend.position = "bottom")

```



## Map U.S. county-level data

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
pop_p2_rev <- pop_p1 + scale_fill_manual(values = orange_rev)

pop_p2_rev + labs(title = "Reverse-coded Population Density",
              fill = "People per square mile") +
    theme_map() + theme(legend.position = "bottom")

```



## Map U.S. county-level data {.build .faster} 

- Let's consider a county map of a continuous variable, such as `per_gop_2016`.
  - Check `class(county_full$per_gop_2016)`.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
gop_p <- ggplot(data = county_full,
            mapping = aes(x = long, y = lat,
                          fill = per_gop_2016, 
                          group = group))

gop_p1 <- gop_p + geom_polygon(color = "gray70", size = 0.05) + coord_equal()
gop_p1
```



## Map U.S. county-level data

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
gop_p <- ggplot(data = county_full,
            mapping = aes(x = long, y = lat,
                          fill = per_gop_2016, 
                          group = group))

gop_p1 <- gop_p + geom_polygon(color = "gray70", size = 0.05) + coord_equal()
gop_p1
```





## Map U.S. county-level data {.build .faster} 

- For a continuous variable, we can use `scale_fill_gradient()`, `scale_fill_gradient2()`, or `scale_fill_gradient2()` function:
  - `scale_fill_gradient()` produces a two-color gradient.
  - `scale_fill_gradient2()` produces a three-color gradient with specified midpoint.
  - `scale_fill_gradientn()` produces an n-color gradient.

- For `scale_fill_gradient2()`, choose the value and color for `midpoint` carefully.


## Map U.S. county-level data {.build .faster} 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
gop_p2 <- gop_p1 + scale_fill_gradient2( 
    low = '#2E74C0',  # from party_colors for DEM
    mid = '#FFFFFF',  # transparent white
    high = '#CB454A',  # from party_colors for GOP
    na.value = "grey50",
    midpoint = .5)

gop_p2 + labs(title = "US Presidential Election 2016",
              fill = "Trump vote share") +
    theme_map() + theme(legend.position = "bottom")

```



## Map U.S. county-level data

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
gop_p2 <- gop_p1 + scale_fill_gradient2( 
    low = '#2E74C0',  # from party_colors for DEM
    mid = '#FFFFFF',  # transparent white
    high = '#CB454A',  # from party_colors for GOP
    na.value = "grey50",
    midpoint = .5)   # set the midpoint value

gop_p2 + labs(title = "US Presidential Election 2016",
              fill = "Trump vote share") +
    theme_map() + theme(legend.position = "bottom")

```



## Statebins {.build .faster} 
- As an alternative to state-level choropleths, we can consider `statebins`.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
# install.packages("statebins")
library(statebins)
p <- ggplot( data = election, 
             mapping = aes( state = state, fill = pct_trump ) )
p1 <- p +  geom_statebins(lbl_size = 5,
                          border_col = "grey90", border_size = 1)
p2 <- p1 + labs(fill = "Percent Trump") +
  coord_equal() +
  theme_statebins( legend_position = c(.45, 1) ) +
  theme( legend.direction="horizontal" )

p2 + scale_fill_gradient2( 
    low = '#2E74C0',  # from party_colors for DEM
    mid = '#FFFFFF',  # transparent white
    high = '#CB454A',  # from party_colors for GOP
    na.value = "grey50",
    midpoint = 50)   # set the midpoint value
```



## Statebins

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
library(statebins)
p <- ggplot(data = election, 
            mapping = aes(state=state, fill=pct_trump))
p1 <- p +  geom_statebins(lbl_size = 5,
                          border_col = "grey90",
                          border_size = 1)
p2 <- p1 + labs(fill = "Percent Trump") +
  coord_equal() +
  theme_statebins( legend_position = c(.45, 1) ) +
  theme( legend.direction="horizontal" )

p2 + scale_fill_gradient2( 
    low = '#2E74C0',  # from party_colors for DEM
    mid = '#FFFFFF',  # transparent white
    high = '#CB454A',  # from party_colors for GOP
    na.value = "grey50",
    midpoint = 50)   # set the midpoint value
```






## Statebins {.build .faster} 

- Let's remove DC and use `scale_fill_gradient()`.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = subset(election, st != "DC")  , 
            mapping = aes(state=state, fill=pct_clinton)) 
p1 <- p + geom_statebins(lbl_size = 5,
                         border_col = "grey90",
                         border_size = 1)
p2 <- p1 + labs(fill = "Percent Clinton") +
  coord_equal() +
  theme_statebins( legend_position = c(.45, 1) ) +
  theme( legend.direction="horizontal" )

p2 + scale_fill_gradient( 
    low = '#FFFFFF',  # transparent white
    high = '#2E74C0',  # from party_colors for DEM
    na.value = "grey50")   # set the midpoint value
```


## Statebins

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = subset(election, st != "DC"), 
            mapping = aes(state=state, fill=pct_clinton)) 
p1 <- p + geom_statebins(lbl_size = 5,
                         border_col = "grey90",
                         border_size = 1)
p2 <- p1 + labs(fill = "Percent Clinton") +
  coord_equal() +
  theme_statebins( legend_position = c(.45, 1) ) +
  theme( legend.direction="horizontal" )

p2 + scale_fill_gradient( 
    low = '#FFFFFF',  # transparent white
    high = '#2E74C0',  # from party_colors for DEM
    na.value = "grey50")   # set the midpoint value
```




## Statebins {.build .faster} 

- Let's use `scale_fill_manual()` to `fill` color by `party`.

- `legend_position` allows for adjusting a coordinate for the legend position.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = election  , 
            mapping = aes(state=state, fill=party)) 
p1 <- p + geom_statebins(lbl_size = 5,
                         border_col = "grey90",
                         border_size = 1)
p2 <- p1 + labs(fill = "Winner") +
  coord_equal() +
  theme_statebins( legend_position = c(.25, 1) ) +
  theme( legend.direction="horizontal",
         legend.title = element_text(size=30),
         legend.text = element_text(size=30) )

p2 + scale_fill_manual( values = c(Republican = "darkred", 
                                   Democratic = "royalblue"))

```


## Statebins

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = election  , 
            mapping = aes(state=state, fill=party)) 
p1 <- p + geom_statebins(lbl_size = 5,
                         border_col = "grey90",
                         border_size = 1)
p2 <- p1 + labs(fill = "Winner") +
  coord_equal() +
  theme_statebins( legend_position = c(.25, 1) ) +
  theme( legend.direction="horizontal",
         legend.title = element_text(size=30),
         legend.text = element_text(size=30) )

p2 + scale_fill_manual( values = c(Republican = "darkred", 
                                   Democratic = "royalblue"))

```




## Statebins {.build .faster} 

- Let's discretize a continuous variable using `scale_fill_gradient()` with `breaks`.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = election  , 
            mapping = aes(state = state, fill=pct_trump)) 
p1 <- p + geom_statebins(lbl_size = 5,
                         border_col = "grey90",
                         border_size = 1)
p2 <- p1 + labs(fill = "Percent Trump") +
  coord_equal() +
  theme_statebins( legend_position = c(.2, 1) ) +
  theme( legend.direction="horizontal")

p2 + scale_fill_gradient(breaks = c(5, 21, 41, 48, 57),
                         labels = c("< 5", "5-21", "21-41", "41-58", "> 57"),
                         low = "#f9ecec", high = "#CB454A") +
  guides(fill = guide_legend())

```



## Statebins

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = election  , 
            mapping = aes(state = state, fill=pct_trump)) 
p1 <- p + geom_statebins(lbl_size = 5,
                         border_col = "grey90",
                         border_size = 1)
p2 <- p1 + labs(fill = "Percent Trump") +
  coord_equal() +
  theme_statebins( legend_position = c(.2, 1) ) +
  theme( legend.direction="horizontal")

p2 + scale_fill_gradient(breaks = c(5, 21, 41, 48, 57),
                         labels = c("< 5", "5-21", "21-41", "41-58", "> 57"),
                         low = "#f9ecec", high = "#CB454A") +
  guides(fill = guide_legend())

```


## Small-multiple maps {.build .faster}
- Sometimes we have geographical data with repeated observations over time.
  - A common case is to have a country- or state-level measure observed over a period of years (Panel data).
  
- Let's consider consider the poverty rate determined by level of educational attainment in NY.

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
NY_socioecon_geo_poverty <- read.table(
  'https://bcecon.github.io/NY_socioecon_geo_poverty.csv',
  sep = ',',
  header = TRUE,
  stringsAsFactor = TRUE
)

library(viridis)
```

- The `viridis` colors run in low-to-high sequences and combines perceptually uniform colors with easy-to-see, easily-contrasted hues along their scales. 

  - The `scale_fill_viridis_c()` function is for continuous data.
  - The `scale_fill_viridis_d()` function is for discrete data.



## Small-multiple maps {.build .faster}
```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = NY_socioecon_geo_poverty,
            mapping = aes(x = long, y = lat, group = group, 
                          fill = c04_058 ))
  
p1 <- p + geom_polygon(color = "grey", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p2 <- p1 + scale_fill_viridis_c(option = "plasma") + theme_map() 

p2 + facet_wrap(~ year, ncol = 3) +
    theme(legend.position = "bottom",
          strip.background = element_blank()) +
    labs(fill = "Poverty rate in NY (%)",
         title = "Poverty rate for the male population 25 years and over \nfor whom the highest educational attainment is bachelor's degree")
```

- We facet the maps just like any other small-multiple with `facet_wrap()`. 


## Small-multiple maps

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = NY_socioecon_geo_poverty,
            mapping = aes(x = long, y = lat, group = group, 
                          fill = c04_058 ))
  
p1 <- p + geom_polygon(color = "grey", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p2 <- p1 + scale_fill_viridis_c(option = "plasma")

p2 + theme_map() + facet_wrap(~ year, ncol = 3) +
    theme(legend.position = "bottom",
          strip.background = element_blank()) +
    labs(fill = "Poverty rate in NY (%)",
         title = "Poverty rate for the male population 25 years and over for whom the highest educational attainment is bachelor's degree")
```





