---
title: "DANL 310 Lecture 4"
subtitle: "DANL 310: Data Visualization and Presentation"
author: "Byeong-Hak Choe"
institute: "SUNY Geneseo" 
date: "Febraury 2, 2023"
output:
  xaringan::moon_reader:
    css: 
      - default
      - css/nhsr.css
      - css/nhsr-fonts.css
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: googlecode
      highlightLines: true
      highlightLanguage: ["r"]
      countIncrementalSlides: false
      ratio: "16:9"
    includes:
      after_body: [css/insert-logo.html]
---

```{r setup, include = FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(NHSRtheme)
library(fontawesome)
# set default options
opts_chunk$set(echo = T, eval = F,
               fig.width = 7.252,
               fig.height = 4,
               comment = "#",
               dpi = 300)

knitr::knit_engines$set("markdown")

xaringanExtra::use_tile_view()
xaringanExtra::use_panelset()
xaringanExtra::use_clipboard()
xaringanExtra::use_webcam()
xaringanExtra::use_broadcast()
xaringanExtra::use_share_again()
xaringanExtra::style_share_again(
  share_buttons = c("twitter", "linkedin", "pocket")
)


xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         #<<
  mute_unhighlighted_code = TRUE  #<<
)

# uncomment the following lines if you want to use the NHS-R theme colours by default
# scale_fill_continuous <- partial(scale_fill_nhs, discrete = FALSE)
# scale_fill_discrete <- partial(scale_fill_nhs, discrete = TRUE)
# scale_colour_continuous <- partial(scale_colour_nhs, discrete = FALSE)
# scale_colour_discrete <- partial(scale_colour_nhs, discrete = TRUE)
```

class: title-slide, left, bottom

# `r rmarkdown::metadata$title`
----
## **`r rmarkdown::metadata$subtitle`**
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$date`

---
class: inverse, center, middle

# Get Started with `ggplot` visualization
<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>



---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Make a Plot </p>

- Before we start visualzation, set the R packages and default chunk options using `include = F` chunk option.
````{verbatim}
```{r setup, include = F}
library(tidyverse)
library(skimr)   # a better summary of data.frame
library(scales)  # scales for ggplot
library(ggthemes)  # additional ggplot themes
theme_set(theme_minimal()) # setting the minimal theme for ggplot

# setting default chunk options
knitr::opts_chunk$set(echo = T, eval = T,
                      message=F, warning = F) 
```
````


---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> ggplot Themes </p>

- Use `theme()` if you want to tweak the display of an existing theme.  

- The following lists basic `ggplot` themes:  

theme            | theme
-----------------|-----------------
theme_grey()     | theme_gray()
theme_bw()       | theme_linedraw()
theme_light()    | theme_dark()
theme_minimal()  | theme_classic()
theme_void()     | theme_test()



---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> ggplot Themes </p>

- `ggthemes` package provides the following themes:
  - `theme_economist()`, `theme_wsj()`, `theme_fivethirtyeight()`, `theme_gdocs()`, `theme_map()`


- `hrbrthemes` package provides the following themes:
  - `theme_ipsum()`, `theme_ipsum_rc()`, `theme_ft_rc()`, and more.
  
  
---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> `gapminder` data </p>

- The `gapminder` package include the `gapminder` data frame.

```{r}
library(gapminder)
gapminder
skim(gapminder)

```



---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Mappings link data to things we see  </p>

- First, we tell the core `ggplot` function what our data is and what our aesthetic mapping is:
```{r}
p <- ggplot(data = gapminder)
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
```


- We can then add the geometric object (`geom_*()`) to the `ggplot()`.

```{r}
p + geom_point() 
```





---
# Get Started with `ggplot` visualization

*1*. Tell the `ggplot()` function what our **tidy** data is. 
  - `ggplot(data = ...)`


*2*. Tell `ggplot()` what relationships we want to see.
  - The `ggplot(mapping = aes(...))` 
  - `p <- ggplot(data = ... , aes(...))`
  
  
*3*. Tell `ggplot()` how we want to see the relationships in our data.
  - Choose a `geom_*()`.
  
  
*4*. Layer on `geoms` as needed, by adding them to the `p` object one at a time.


*5*. Use some additional functions to adjust scales, labels, tick marks, titles, etc.



---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Build your plots layer by layer </p>

```{r}
p + geom_point() + geom_smooth() 
p + geom_point() + geom_smooth(method = "lm") 
p + geom_point() +
    geom_smooth(method = "lm")
```


```{r}
p + geom_point() +
    geom_smooth(method = "gam")
```


---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Mapping aesthetics vs setting them </p>

- In the following, we map `color` to `continent` in the `aes()` function:
```{r}
p1 <-  ggplot(data = gapminder,
             mapping = aes(x = gdpPercap, y = lifeExp,
                           color = continent))
```

- We then add geoms.
```{r}
p1 + geom_point() +
     geom_smooth(method = "lm") 
```




---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Mapping aesthetics vs setting them </p>

- In the following, we map `color` to `"purple"` outside of the `aes()` function:

```{r}
p2 <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, y = lifeExp))
```

- We then add geoms, manually set aethetics, and then adjust scales.
```{r}
p2 + geom_point(color = "purple") +
     geom_smooth(method = "lm")
```



---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Mapping aesthetics vs setting them </p>

- Here are another exmaples that set aesthetic mapping manually outside `aes()` function and inside `geom_*()` function.
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))

p + geom_point(color = "purple") +
    geom_smooth(method = "loess") 

p + geom_point(alpha = 0.3) +
    geom_smooth(color = "orange", se = F, size = 8, method = "lm") 
```


---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Mapping aesthetics vs setting them </p>

- Let's add `scale_*()` and `labs()` functions to `ggplot()`

```{r}
p + geom_point(alpha = 0.3) +
    geom_smooth(method = "gam") +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapminder.")
```

- **Q**. Is it really the best way to display this country-year data? 
  - What are we gaining and losing by ignoring the temporal and country-level structure of the data? How could we do better? 

---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Aesthetics can be mapped per geom </p>

- We can have different aesthetic mappings across different geoms:
```{r}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(mapping = aes(color = continent)) +
    geom_smooth(method = "loess")  +
    scale_x_continuous(trans = scales::log_trans())  # natural log
```



---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Save your work </p>

- We can use `ggsave()` to save ggplot output as a *.png or *.pdf file.
```{r}
ggsave(filename = "my_figure.png")
```

- We can put our recent plot into an object called p_out and then tell `ggave()` that we want to save that object.

```{r}
p_out <- p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
ggsave("my_figure.pdf", plot = p_out)
```


---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Save your work </p>


- We can set the size of ggplot figue:

```{r, message=FALSE, warning=FALSE}
ggsave("lifexp_vs_gdp_gradient.pdf",
       plot = p_out, height = 8, width = 10, units = "in")
```




---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Work a little more with `gapminder` data </p>

- What happens when you put the `geom_smooth()` function before `geom_point()` instead of after it? 
  - What does this tell you about how the plot is drawn? Think about how this might be useful when drawing plots.
  


---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Work a little more with `gapminder` data </p>

- Change the mappings in the `aes()` function so that you plot Life Expectancy against population (`pop`) rather than per capita GDP. 
  - What does that look like? 
  - What does it tell you about the unit of observation in the dataset?
  


---
# Get Started with `ggplot` visualization
### <p style="color:#00449E"> Work a little more with `gapminder` data </p>
- What happens if you map color to `year` instead of `continent`? 
  - Is the result what you expected? 
  - Think about what class of object year is. 
  - Remember you can get a quick look at the top of the data, which includes some shorthand information on the class of each variable, by typing `gapminder`.
  - Instead of mapping `color = year`, what happens if you try `color = factor(year)`?


---
# Show the right number
### <p style="color:#00449E"> Grouped data and the `group` aesthetic </p>

- Let's get a line plot that draws the trajectory of life expectancy over time for each country in the `gapminder` data.frame. 

```{r}
p <- ggplot(data = gapminder,
            mapping = aes( x = year,
                           y = gdpPercap ) )

p + geom_line() 
```

- What happened?

- `geom_line()` joins up all the lines for each particular year in the order they appear in the dataset.

---
# Show the right number
### <p style="color:#00449E"> Grouped data and the `group` aesthetic </p>


- Without group related parameters, `ggplot()` does not know that the yearly observations in the data are grouped by country.

```{r}
p <- ggplot(data = gapminder,
            mapping = aes( x = year,
                           y = gdpPercap ) )

p + geom_line( aes( group = country ) ) 
```



---
# Show the right number
### <p style="color:#00449E"> Grouped data and the `group` aesthetic </p>


- The `group` aesthetic is usually only needed when the grouping information we need to tell `ggplot()` about is not built-in to the variables being mapped.

  - How about `color` aesthetic, instead of `group`?

```{r}
p <- ggplot(data = gapminder,
            mapping = aes( x = year,
                           y = gdpPercap ) )

p + geom_line( aes( color = country ) ) 
```

---
# Show the right number
### <p style="color:#00449E"> Facet to make small multiples </p>
- Making a “small multiple” plot by faceting data based on a caterigorical variable allows a lot of information to be presented compactly, and in a consistently comparable way. 
  - `facet_wrap( VAR1 ~ . )` or `facet_wrap( . ~ VAR1 )`
  - `facet_grid( VAR1 ~ . )`: row-wise split
  - `facet_grid( . ~ VAR1 )`: colum-wise split
  - `facet_grid( VAR1 ~ VAR2 )` 

```{r}
p + geom_line( aes( group = country ) ) + 
  facet_wrap(~ continent)
```


---
# Show the right number
### <p style="color:#00449E"> Facet to make small multiples </p>

- Let's have all the facetted plots in a single row:
```{r}
p + geom_line(color="gray70", aes(group = country)) +
    geom_smooth(size = 1.1, method = "loess", se = FALSE) +
    facet_[?] +
    scale_y_log10(labels=scales::dollar) + 
    theme(axis.text.x = element_text(angle = 45),
          axis.title.x = element_text(margin = margin(t = 25))) +
    labs(x = "Year", y = "GDP per capita",
         title = "GDP per capita on Five Continents")
```


---
# Show the right number
### <p style="color:#00449E"> The 2016 General Social Survey data </p>

- The `socviz` package includes the `gss_sm` data frame.
  - `gss_sm` is a dataset containing an extract from the 2016 General Social Survey.

```{r}
install.packages("socviz")
library(socviz)
?gss_sm
glimpse(gss_sm)
skimr::skim(gss_sm)
gss_sm <- gss_sm
```


---
# Show the right number
### <p style="color:#00449E"> Facet to make small multiples </p>

- Describe the relationship between the age of the respondent and the number of children they have using a scatterplot and a fitted curve.

```{r}
p <- ggplot(data = gss_sm,
            mapping = aes( x = age, y = childs ))

p + [?](alpha = 0.2) +
    [?]() 
```


---
# Show the right number
### <p style="color:#00449E"> Facet to make small multiples </p>

- Describe how the relationship between the age of the respondent and the number of children they have varies by `sex` and `race`.

```{r}
p <- ggplot(data = gss_sm,
            mapping = aes( x = age, y = childs ))

p + geom_point(alpha = 0.2) +
    geom_smooth() +
    [?]
```



---
# Show the right number
### <p style="color:#00449E"> Facet to make small multiples </p>

- The `facet_grid()` function is best used when you cross-classify some data by two categorical variables.

  - e.g., the relationship between the age and the number of children by sex and race


```{r}
p <- ggplot(data = gss_sm,
            mapping = aes( x = age, y = childs ))

p + geom_point(alpha = 0.2) +
    geom_smooth() +
    facet_grid(sex ~ race + degree)
```


---
# Show the right number
### <p style="color:#00449E"> Geoms can transform data </p>

- Let's plot a bar char:

```{r}
p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion))
p + geom_bar()
```


---
# Show the right number
### <p style="color:#00449E"> Geoms can transform data </p>


- Where does count come from? 

  - Bar charts, histograms, and frequency polygons bin your data and then plot bin counts, the number of points that fall in each bin.

  - Smoothers fit a model to your data and then plot predictions from the model.

  - Boxplots compute a robust summary of the distribution and then display a specially formatted box.




---
# Show the right number
### <p style="color:#00449E"> Geoms can transform data </p>


- If we want a chart of relative frequencies rather than counts, we will need to get the `prop` statistic instead.

- Our call to statistic from the `aes()` function generically looks like this: `<mapping> = <..statistic..>`. 

```{r}
p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop..))
```

- What happened?



---
# Show the right number
### <p style="color:#00449E"> Geoms can transform data </p>

- We need to tell ggplot to ignore the x-categories when calculating denominator of the proportion, and use the total number observations instead.

  - To do so we specify `group = 1` inside the `aes()` call. 

```{r}
p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop.., group = 1)) 
```


---
# Show the right number
### <p style="color:#00449E"> Geoms can transform data </p>

- Let’s look at another question from the survey. The `gss_sm` data contains a religion variable derived from a question asking:

  - “What is your religious preference? Is it Protestant, Catholic, Jewish, some other religion, or no religion?”

```{r}
table(gss_sm$religion)
```

---
# Show the right number
### <p style="color:#00449E"> Geoms can transform data </p>

```{r}
p <- ggplot(data = gss_sm,
            mapping = aes(x = religion, color = religion))
p + geom_bar()
```

- If we map religion to `color`, only the border lines of the bars will be assigned colors, and the insides will remain gray.


---
# Show the right number
### <p style="color:#00449E"> Geoms can transform data </p>

-  If the gray bars look boring and we want to fill them with color instead, we can map the religion variable to `fill` in addition to mapping it to `x`.

```{r}
p <- ggplot(data = gss_sm,
            mapping = aes(x = religion, fill = religion))
p + geom_bar() + guides( fill = "none" )
```

-  If we set `guides(fill = "none")`, the legend about the `fill` mapping is removed.


---
# Show the right number
### <p style="color:#00449E"> Frequency plots the slightly awkward way </p>

- A more appropriate use of the `fill` aesthetic with `geom_bar()` is to cross-classify two categorical variables. 

  - The default output of such `geom_bar()` is a stacked bar chart, with counts on the y-axis.


```{r}
p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion, fill = religion))
p + geom_bar(position = "fill")
```
- An alternative choice is to set the position argument to `"fill"`.
  - It is to compare proportions across groups.



---
# Show the right number
### <p style="color:#00449E"> Frequency plots the slightly awkward way </p>

- We can use `position = "dodge"` to make the bars within each region of the country appear side by side.

```{r}
p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion, fill = religion))
p + geom_bar(position = "dodge",
             mapping = aes(y = ..prop..))

```

- What happened?

---
# Show the right number
### <p style="color:#00449E"> Frequency plots the slightly awkward way </p>

- In this case we should consider grouping variable, `religion`, so we map `religion` to the `group` aesthetic.

```{r}
p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion, fill = religion))
p + geom_bar(position = "dodge",
             mapping = aes(y = ..prop.., group = religion))
```


---
# Show the right number
### <p style="color:#00449E"> Frequency plots the slightly awkward way </p>

- How can we have a proportional bar chart such that the sum of all bars in each bigregion is 1?

  - There are various ways to do so, and faceting is one of them.
  - The proportions are calculated within each panel, which is the breakdown we wanted.


```{r}
p <- ggplot(data = gss_sm,
            mapping = aes(x = religion))
p + geom_bar(position = "dodge",
             mapping = aes(y = ..prop.., group = bigregion)) +
    facet_wrap(~ bigregion, ncol = 1)
```



---
# Show the right number
### <p style="color:#00449E"> Histograms and density plots </p>

- ggplot comes with a dataset, `midwest`, containing information on counties in several midwestern states of the USA. 

```{r}
?midwest
glimpse(midwest)
skim(midwest)
view(midwest)
```

---
# Show the right number
### <p style="color:#00449E"> Histograms and density plots </p>

- By default, the `geom_histogram()` function will choose a bin size for us based on a rule of thumb.

```{r}
p <- ggplot(data = midwest,
            mapping = aes(x = area))
p + geom_histogram()
```



---
# Show the right number
### <p style="color:#00449E"> Histograms and density plots </p>

- When drawing histograms it is worth experimenting with `bins` and also optionally the `origin` of the x-axis. 
```{r}
p <- ggplot(data = midwest,
            mapping = aes(x = area))
p + geom_histogram(bins = 10)
```



