---
title: "Lecture 15"
subtitle: "DANL 200: Introduction to Data Analytics"
author: "Byeong-Hak Choe"
institute: "SUNY Geneseo"
date: "October 20, 2022"
output:
  xaringan::moon_reader:
    css: 
      - default
      - css/nhsr.css
      - css/nhsr-fonts.css
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: googlecode
      highlightLines: true
      highlightLanguage: ["r"]
      countIncrementalSlides: false
      ratio: "16:9"
    includes:
      after_body: [css/insert-logo.html]
---

```{r setup, include = FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(NHSRtheme)
library(fontawesome)

# set default options
opts_chunk$set(echo = FALSE,
               fig.width = 7.252,
               fig.height = 4,
               comment = "#",
               dpi = 300)

knitr::knit_engines$set("markdown")

xaringanExtra::use_tile_view()
xaringanExtra::use_panelset()
xaringanExtra::use_clipboard()
xaringanExtra::use_webcam()
xaringanExtra::use_broadcast()
xaringanExtra::use_share_again()
xaringanExtra::style_share_again(
  share_buttons = c("twitter", "linkedin", "pocket")
)


xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         #<<
  mute_unhighlighted_code = F  #<<
)

# uncomment the following lines if you want to use the NHS-R theme colours by default
# scale_fill_continuous <- partial(scale_fill_nhs, discrete = FALSE)
# scale_fill_discrete <- partial(scale_fill_nhs, discrete = TRUE)
# scale_colour_continuous <- partial(scale_colour_nhs, discrete = FALSE)
# scale_colour_discrete <- partial(scale_colour_nhs, discrete = TRUE)
```


class: title-slide, left, bottom

# `r rmarkdown::metadata$title`
----
## **`r rmarkdown::metadata$subtitle`**
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$date`



---
# Exam Policies

During the exam, ...
- Students can read any books, paper sheets, PDF files on PDF reader apps, R Scripts or text files on RStudio.
  - Students are not allowed to use other apps, for example, Google Docs, Google Sheets, Microsoft Word, Microsoft Excel, Microsoft PowerPoint, Numbers, Pages, Keynotes, Notes.
- Students can visit (1) the class website (https://bcdanl.github.io) and (2) the Canvas website for the course.
  - Students are not allowed to visit any other websites, such as Google, Stack Overflow, and Gmail.
- Students can communicate with Byeong-Hak Choe.
  - Students are not allowed to communicate with other students.


---
class: inverse, center, middle

# Data Transformation
<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
  

---
# Data Transformation
### <p style="color:#00449E"> `dplyr` basics </p>

- `filter(<DATA.FRAME>, LOGICAL_STATEMENT, !is.na(VARIABLE), ...)`
- `arrange(<DATA.FRAME>, VARIABLE1, desc(VARIABLE2), -VARIABLE3, ...)`
- `select(<DATA.FRAME>, VARIABLE, -VARIABLE2, starts_with("abc"), contains("time"), everything(), A:Z, ...)`
- `rename(<DATA.FRAME>, <NEW_VARIABLE> = <EXISTING_VARIABLE>)`
- `mutate(<DATA.FRAME>, <NEW_VARIABLE> =  ...)`
- `summarise(<DATA.FRAME>, <NEW_VARIABLE> = ...)`
- `group_by(<DATA.FRAME>, VARIABLE1, VARIABLE2, ...)`


- The first argument of the functions above is a data frame, unless the pipe operator ( `%>%` ) is used.
- The subsequent arguments describe what to do with the data frame, using the variable names.
- The result is a new data frame.




---
class: inverse, center, middle

# `mutate()`
<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---
# Add new variables with `mutate()`


- `mutate()` is useful to add new columns that are functions of existing columns.


```{r, echo = T, eval = F}
flights_sml <- select(flights, 
                      year:day, 
                      ends_with("delay"), 
                      distance, 
                      air_time )

mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60 )

mutate(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours )
```



---
# Add new variables with `mutate()`
- If we only want to keep the new variables, use `transmute()`.


```{r, echo = T, eval = F}
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```



---
# Add new variables with `mutate()`
### <p style="color:#00449E"> Useful creation functions

- There are many functions for creating new variables that we can use with `mutate()`.

- Arithmetic operators: `+`, `-`, `*`, `/`, `^`
- Modular arithmetic: `%/%` (integer division) and `%%` (remainder).


```{r, echo = T, eval = F}
transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```


---
# Add new variables with `mutate()`

- There are many functions for creating new variables that we can use with `mutate()`.

- Offsets: `lead()` and `lag()` allow us to refer to leading or lagging values. 

- Cumulative and rolling aggregates: R provides functions for running sums, products, mins and maxes: `cumsum()`, `cumprod()`, `cummin()`, `cummax()`, and `cummean()`


```{r, echo = T, eval = F}
(x <- 1:10)
lag(x)
lead(x)
cumsum(x)
cummean(x)
```




---
# Add new variables with `mutate()`
### <p style="color:#00449E"> Useful creation functions


.panelset[
.panel[.panel-name[rank functions]

- There are many functions for creating new variables that we can use with `mutate()`.


- Ranking: `min_rank()`, `dense_rank()`, `row_number()`, , `percent_rank()`, and `cume_dist()`.



```{r, echo = T, eval = F}
rank_me <- tibble( x = c(10, 5, 1, 5, 5, NA) )
rank_me <- mutate(rank_me,
                  id_x = row_number(x),
                  x_min_rank = min_rank(x),
                  x_dense_rank = dense_rank(x) )
```
]


.panel[.panel-name[definitions of rank functions]

- `row_number()` assigns each element a unique value.

- `min_rank()` and `dense_rank()` assign tied values the same rank.

  - `min_rank()` function assigns a rank equal to the number of values less than that tied value plus one.

  - `dense_rank()` function assigns a rank equal to the number of distinct values less than that tied value plus one. 

]

]

---
# Add new variables with `mutate()`
### <p style="color:#00449E"> Useful creation functions

- To create new variables based on a condition, we can use the `if_else()` and `ifelse()` functions. 
  - `if_else(<condition>, <if TRUE>, <else>)`
  - `ifelse(<condition>, <if TRUE>, <else>)`


```{r, echo = T, eval = F}
flight_season <-mutate(flights,
    summer_month = if_else(month %in% c(6, 7, 8), TRUE, FALSE))

flight_season <-mutate(flights,
    summer_month = if_else(month %in% c(6, 7, 8), TRUE, 0))

flight_season <-mutate(flights,   # ifelse() is less restrictive
    summer_month = ifelse(month %in% c(6, 7, 8), TRUE, 0))
```



---
# Add new variables with `mutate()`
### <p style="color:#00449E"> Exercises
 
- Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

- Compare `air_time` with `arr_time - dep_time`. What do you expect to see? What do you see? What do you need to do to fix it?

- Compare `dep_time`, `sched_dep_time`, and `dep_delay.` How would you expect those three numbers to be related?

- Find the 10 most delayed flights using a ranking function. How do you want to handle ties? 



---
class: inverse, center, middle

# `summarize()`
<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
  


---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> 

- `summarize()` collapses a data frame to a single row.
- `na.rm` argument removes the missing values prior to computation when using `summarize()` or other aggregate functions.
- `summarize()` is not terribly useful unless we pair it with `group_by()`.


```{r, echo = T, eval = F}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))

by_day <- group_by(flights, year, month, day)
summarize(by_day, delay = mean(dep_delay, na.rm = TRUE))
```




---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Example

- Explore the relationship between the mean level `distance` and the mean level `arr_delay` for each value of `dest`.


```{r, echo = T, eval = F}
by_dest <- group_by(flights, dest)
delay <- summarize(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)

delay <- filter(delay, count > 20, dest != "HNL")

ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```






---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Combining multiple operations with the pipe (`%>%`)

- Explore the relationship between the distance and average delay for each location using the pipe (`%>%`).


```{r, echo = T, eval = F}
delays <- flights %>% 
  group_by(dest) %>% 
  summarize( count = n(),
             dist = mean(distance, na.rm = TRUE),
             delay = mean(arr_delay, na.rm = TRUE) ) %>% 
  filter(count > 20, dest != "HNL")
```

- A good way to pronounce %>% when reading code is “then.”
- A shortcut for %>%  is Ctrl+Shift+M  (Command+Shift+M for Mac).




---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Missing values

- What happens if we don’t set `na.rm`?


```{r, echo = T, eval = F}
flights %>% 
  group_by(year, month, day) %>% 
  summarize(mean = mean(dep_delay))
  
flights %>% 
  group_by(year, month, day) %>% 
  summarize(mean = mean(dep_delay, na.rm = TRUE))
```




---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Missing values

- Suppose missing values represent cancelled flights.


```{r, echo = T, eval = F}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarize(mean = mean(dep_delay))
```




---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Counts

- Let’s look at the planes (identified by their tail number (`tailnum`)) that have the highest average delays


```{r, echo = T, eval = F}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarize(
    delay = mean(arr_delay)
  )

ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```




---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Counts

- When doing any aggregation, it’s always a good idea to include either a count (`n()`), or a count of non-missing values (`sum(!is.na(x))`).


```{r, echo = T, eval = F}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarize( delay = mean(arr_delay, na.rm = TRUE),
             n = n() )

ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)

delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)
```




---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Useful summary functions

- Measures of location: `mean(x)` and `median(x)`. 

- Measures of spread: `sd(x)`, `IQR(x)`. 



```{r, echo = T, eval = F}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarize( avg_delay1 = mean(arr_delay),
             avg_delay2 = mean(arr_delay[arr_delay > 0]) ) # the average positive delay
# Brackets[] lets us filter (or subset) data from a column
# Why is distance to some destinations more variable than to others?
not_cancelled %>% 
  group_by(dest) %>% 
  summarize(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```





---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Useful summary functions


- Measures of rank: `min(x)`, `quantile(x, 0.25)`, `max(x)`, `range(x)`.
- Measures of position: `first(x)`, `nth(x, 2)`, `last(x)`.


- When do the first and last flights leave each day?

.panelset[

.panel[.panel-name[(1)]

```{r, echo = T, eval = F}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarize(first = min(dep_time), 
            last = max(dep_time) )
```
]


.panel[.panel-name[(2)]
```{r, echo = T, eval = F}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarize(first_dep = first(dep_time),last_dep = last(dep_time) )
```
]

.panel[.panel-name[(3)]

```{r, echo = T, eval = F}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r) )
```
]

]





---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Useful summary functions

- Measures of count: `sum(!is.na(x))`, `n_distinct(x)`,  `count(x)`, and `count(x, wt = VARIABLE)`.

```{r, echo = T, eval = F}
# Which destinations have the most carriers?
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
  
  
not_cancelled %>% 
  count(dest)

not_cancelled %>% 
  count(tailnum, wt = distance)
```




---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Useful summary functions

- We can compute the height of bars in a histogram by combining `dplyr::count()` and `ggplot2::cut_width()`:

```{r, echo = T, eval = F}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)

diamonds %>% 
  count(cut_width(carat, 0.5))
```





---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Useful summary functions


- Counts and proportions of logical values: `sum(x > 10)`, `mean(y == 0)`.




```{r, echo = T, eval = F}
# How many flights left before 5am? 
# (these usually indicate delayed flights from the previous day)
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarize(n_early = sum(dep_time < 500))

# What proportion of flights are delayed by more than an hour?
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarize(hour_prop = mean(arr_delay > 60))
```





---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Grouping by multiple variables


- When you group by multiple variables, each summary peels off one level of the grouping.




```{r, echo = T, eval = F}
daily <- group_by(flights, year, month, day)

(per_day   <- summarize(daily, flights = n()))

(per_month <- summarize(per_day, flights = sum(flights)))

(per_year  <- summarize(per_month, flights = sum(flights)))
```





---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Ungrouping


- If you need to remove grouping, and return to operations on ungrouped data, use `ungroup()`.



```{r, echo = T, eval = F}
daily %>% 
  ungroup() %>%             # no longer grouped by date
  summarize(flights = n())  # all flights
```





---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Exercises


- Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

  - A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.

  - A flight is always 10 minutes late.

  - A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.

  - 99% of the time a flight is on time. 1% of the time it’s 2 hours late.

Q. Which is more important: arrival delay or departure delay?




---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Exercises

- Come up with another approach that will give you the same output as `not_cancelled` `%>%` `count(dest)` and `not_cancelled %>% count(tailnum, wt = distance)` (without using `count()`).

- Our definition of cancelled flights (`is.na(dep_delay) | is.na(arr_delay)` ) is slightly suboptimal. Why? Which is the most important column?

- Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?



---
# Grouped summaries with `summarize()`
### <p style="color:#00449E"> Exercises

- Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about `flights %>% group_by(carrier, dest) %>% summarize(n())`)

- What does the sort argument to `count()` do. When might you use it?




---
class: inverse, center, middle

# Grouped mutates
<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
  

---
# Grouped mutates (and filters)
### <p style="color:#00449E"> 

- Grouping is most useful in conjunction with `summarize()`, but you can also do convenient operations with `mutate()` and `filter()`.

.panelset[

.panel[.panel-name[worst]
- Find the worst members of each group:

```{r, echo = T, eval = F}
flights %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)
```
]


.panel[.panel-name[threshold]
- Find all groups bigger than a threshold:

```{r, echo = T, eval = F}
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 17250)
popular_dests
```
]

.panel[.panel-name[standardization]
- Standardize to compute per group metrics:

```{r, echo = T, eval = F}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```
]

]

---
# Grouped mutates (and filters)
### <p style="color:#00449E"> Exercises

- Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.

- Which plane (`tailnum`) has the worst on-time record?

- What time of day should you fly if you want to avoid delays as much as possible?

- For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.








---
# Grouped mutates (and filters)
### <p style="color:#00449E"> Exercises

- Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using `lag()`, explore how the delay of a flight is related to the delay of the immediately preceding flight.

- Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time of a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?



---
# Grouped mutates (and filters)
### <p style="color:#00449E"> Exercises

- Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.

- For each plane, count the number of flights before the first delay of greater than 1 hour.




---
# Data Transformation
### <p style="color:#00449E"> Class Exercises

The following data is for the Exercises:

```{r, echo = T, eval = F}
NY_CC_tweets <- read_csv(
  'https://bcdanl.github.io/data/NY_tweets_cc_no_contents.csv')
```




