---
title: "DANL 310 Lecture 09"
author: "Byeong-Hak Choe"
date: February 24, 2022
output: 
  ioslides_presentation:
    transition: 0
    widescreen: TRUE
    smaller: FALSE
    smart: TRUE
---
# Work with models; Draw maps


## R packages

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
library(tidyverse)
library(socviz)
library(gapminder)
library(skimr)
library(ggrepel)
library(ggthemes)
library(broom)
theme_set(theme_minimal())
```

## Work with models {.build .faster}

- Let us consider the linear regression of `lifeExp` on `gdpPercap`, `pop`, and `continent`.

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
out <- lm(formula = lifeExp ~ log(gdpPercap) + pop + continent,
          data = gapminder)

summary(out)
str(out)
```

- Try `out$coefficients`, `out$residuals`, and `out$fitted.values`.

## Tidy model objects with `broom` {.build .faster}
- The `broom` package to help us do a lot of things with our model output.

- `broom` can tidily extract three kinds of information. 
  - We can see *component-level* information about aspects of the model itself, such as **coefficients** and **t-statistics**. 
  - We can obtain *observation-level* information about the model's connection to the underlying data, such as **fitted values** and **residuals**. 
  - We can get *model-level* information that summarizes the fit as a whole, such as **F-statistic**, **model deviance**, or **r-squared**. 

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
library(broom)
```




## Get component-level statistics with `tidy()`{.build .faster}
- The `tidy()` function takes a model object and returns a data frame of component-level information.

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
out_comp <- tidy(out)
out_comp %>% round_df()
```





## Get component-level statistics with `tidy()`{.build .faster}
- We can work with `tidy()` to make plots in a familiar way, and much more easily.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(out_comp, mapping = aes(x = term,
                                    y = estimate))

p + geom_point() + coord_flip() 
```



## Get component-level statistics with `tidy()`
```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(out_comp, mapping = aes(x = term,
                                    y = estimate))

p + geom_point() + coord_flip() 
```



## Get component-level statistics with `tidy()`{.build .faster}
- We can tell `tidy()` to calculate confidence intervals for the estimates, using R's `conf.int = TRUE` option.

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
out_conf <- tidy(out, conf.int = TRUE)
out_conf %>% round_df()
```




## Get component-level statistics with `tidy()`{.build .faster}
- We can tell `tidy()` to calculate confidence intervals for the estimates, using R's `conf.int = TRUE` option.

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
out_conf <- subset(out_conf, term != "(Intercept)")
out_conf$nicelabs <- prefix_strip(out_conf$term, "continent")
```

- The `socviz::prefix_strip()` strips a series of characters from the beginning of a character vector.


## Get component-level statistics with `tidy()` {.build .faster}
- We can use `geom_pointrange()` to make a figure that displays some information about our confidence in the variable estimates

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(out_conf, mapping = aes(x = reorder(nicelabs, estimate),
                                    y = estimate, ymin = conf.low, ymax = conf.high))
p + geom_pointrange() + coord_flip() + labs(x="", y="OLS Estimate")

```


## Get component-level statistics with `tidy()`

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(out_conf, mapping = aes(x = reorder(nicelabs, estimate),
                                    y = estimate, ymin = conf.low, ymax = conf.high))
p + geom_pointrange() + coord_flip() + labs(x="", y="OLS Estimate")

```


## Get observation-level statistics with `augment()`

- The values returned by `augment()` are all statistics calculated at the level of the original observations. 

- Working from a call to `augment()` will return a data frame with all the original observations used in the estimation of the model, together with columns like the following:
  - `.fitted` — The fitted values of the model.
  - `.se.fit` — The standard errors of the fitted values.
  - `.resid` — The residuals; `.std.resid` — The standardized residuals.
  - `.hat` — The diagonal of the hat matrix.
  - `.sigma` — An estimate of residual standard deviation when the corresponding observation is dropped from the model.
  - `.cooksd` — Cook’s distance, a common regression diagnostic on how the observation is influential in the estimation;
  

## Get observation-level statistics with `augment()` {.build .faster}
```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
out_aug <- augment(out)
head(out_aug) 
head(out_aug) %>% round_df()
```


## Get observation-level statistics with `augment()` {.build .faster}
- The new columns created by `augment()` can be used to create some standard regression plots. 
  - We can draw the residual plot (the residuals versus the fitted values). 
  
```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = out_aug,
            mapping = aes(x = .fitted, y = lifeExp - .fitted))
p + geom_point()

```


## Get observation-level statistics with `augment()` {.build .faster}
```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = out_aug,
            mapping = aes(x = .fitted, y = lifeExp - .fitted))
p + geom_point()

```



## Get model-level statistics with `glance()` {.build .faster}
- `glance()` returns the information typically presented at the bottom of a model's `summary()` output.

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
glance(out) %>% round_df()
```


## Grouped analysis and list columns {.build .faster}

- Consider the linear regression model for the relationship between life expectancy and GDP by continent, for each year in the data.


```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
eu77 <- gapminder %>% filter(continent == "Europe", year == 1977)
fit <- lm(lifeExp ~ log(gdpPercap), data = eu77)
summary(fit)
```


## Grouped analysis and list columns {.build .faster}

- With `dplyr` and `broom` we can do the grouped analysis for every continent-year slice of the data in a compact and tidy way.

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
out_le <- gapminder %>%
    group_by(continent, year) %>%
    nest()
out_le
```

- `nest()` creates a list-column of data frames.
  -  The list column, `data`, contains a small table of data corresponding to each continent-year group.




## Grouped analysis and list columns {.build .faster}

- We can look at specific group, for example, “Europe 1977”, if we like, by filtering the data and then unnesting the list column.

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
out_le %>% filter(continent == "Europe" & year == 1977) %>% 
  unnest()
```



## Grouped analysis and list columns {.build .faster}

- List columns are useful because we can act on them in a compact and tidy way.

```{r skim, include=FALSE}
if("package:maps" %in% search()) detach("package:maps", unload=TRUE) 
```

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
fit_ols <- function(df) {
    lm(lifeExp ~ log(gdpPercap), data = df)
}

out_le <- gapminder %>%
    group_by(continent, year) %>%
    nest() %>% 
    mutate(model = map(data, fit_ols)) 

out_le
```

- `map()` function applies a function to each element of a vector.



## Grouped analysis and list columns {.build .faster}

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
fit_ols <- function(df) {
    lm(lifeExp ~ log(gdpPercap), data = df)
}

out_tidy <- gapminder %>%
    group_by(continent, year) %>%
    nest() %>% 
    mutate(model = map(data, fit_ols),
           tidied = map(model, tidy)) %>%
    unnest(tidied) %>%
    filter(term != "(Intercept)" &
           continent != "Oceania")
```




## Grouped analysis and list columns {.build .faster}
- We now have tidy regression output with an estimate of the association between log GDP per capita and life expectancy *for each year*, *within continents*.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = out_tidy,
            mapping = aes(x = year, y = estimate,
                          ymin = estimate - 2*std.error,
                          ymax = estimate + 2*std.error,
                          group = continent, color = continent))

p + geom_pointrange(position = position_dodge(width = 1)) +
    scale_x_continuous(breaks = unique(gapminder$year)) + 
    theme(legend.position = "top") +
    labs(x = "Year", y = "Estimate", color = "Continent")
```


## Grouped analysis and list columns

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = out_tidy,
            mapping = aes(x = year, y = estimate,
                          ymin = estimate - 2*std.error,
                          ymax = estimate + 2*std.error,
                          group = continent, color = continent))

p + geom_pointrange(position = position_dodge(width = 1)) +
    scale_x_continuous(breaks = unique(gapminder$year)) + 
    theme(legend.position = "top") +
    labs(x = "Year", y = "Estimate", color = "Continent")
```

## Map U.S. state-level data {.build .faster}

- The `socviz::election` dataset has various measures of the vote and vote shares by state. 


```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
election %>% select(state, total_vote,
                    r_points, pct_trump, party, census) %>%
    sample_n(5)

```

- We don’t have to represent spatial data spatially.


## Map U.S. state-level data {.build .faster}

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
party_colors <- c("#2E74C0", "#CB454A")  # Hex color codes for Dem Blue and Rep Red
p0 <- ggplot(data = subset(election, st %nin% "DC"),
             mapping = aes(x = r_points,
                           y = reorder(state, r_points),
                           color = party))

p1 <- p0 + geom_vline(xintercept = 0, color = "gray30") +
    geom_point(size = 2)

p2 <- p1 + scale_color_manual(values = party_colors)

p3 <- p2 + scale_x_continuous(breaks = c(-30, -20, -10, 0, 10, 20, 30, 40),
                              labels = c("30\n (Clinton)", "20", "10", "0",
                                         "10", "20", "30", "40\n(Trump)"))

p3 + facet_wrap(~ census, ncol=1, scales="free_y") +
     guides(color = "none") + labs(x = "Point Margin", y = "") +
     theme(axis.text=element_text(size=8))
```


## Map U.S. state-level data

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
# Hex color codes for Dem Blue and Rep Red
party_colors <- c("#2E74C0", "#CB454A") 

p0 <- ggplot(data = subset(election, st %nin% "DC"),
             mapping = aes(x = r_points,
                           y = reorder(state, r_points),
                           color = party))

p1 <- p0 + geom_vline(xintercept = 0, color = "gray30") +
    geom_point(size = 2)

p2 <- p1 + scale_color_manual(values = party_colors)

p3 <- p2 + scale_x_continuous(breaks = c(-30, -20, -10, 0, 10, 20, 30, 40),
                              labels = c("30\n (Clinton)", "20", "10", "0",
                                         "10", "20", "30", "40\n(Trump)"))

p3 + facet_wrap(~ census, ncol=1, scales="free_y") +
     guides(color = "none") + labs(x = "Point Margin", y = "") +
     theme(axis.text=element_text(size=8))

```

## Map U.S. state-level data {.build .faster}
- Let us get a data frame of the US state map

```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
us_states <- map_data("state")
head(us_states)
dim(us_states)
```



## Map U.S. state-level data {.build .faster}
```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            mapping = aes(x = long, y = lat,
                          group = group))

p + geom_polygon(fill = "white", color = "black")
```

- A map is a set of lines drawn in the right order on a grid.


## Map U.S. state-level data 
```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            mapping = aes(x = long, y = lat,
                          group = group))

p + geom_polygon(fill = "white", color = "black")
```



## Map U.S. state-level data {.build .faster}

- Let's `fill` the map.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            aes(x = long, y = lat,
                group = group, fill = region))

p + geom_polygon(color = "gray90", size = 0.1) + guides(fill = FALSE)
```



## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            aes(x = long, y = lat,
                group = group, fill = region))

p + geom_polygon(color = "gray90", size = 0.1) + guides(fill = FALSE)
```




## Map U.S. state-level data {.build .faster}
- Let’s deal with the projection. 
  - By default, the map is plotted using the venerable *Mercator* projection.
  - The *Albers* projection is nicer. 
 
```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            mapping = aes(x = long, y = lat,
                          group = group, fill = region))

p + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    guides(fill = FALSE)
```

- We can transform the default projection used by `geom_polygon()`, via the `coord_map()` function.

  - The *Albers* projection requires two latitude parameters, `lat0` and `lat1`. 

## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p <- ggplot(data = us_states,
            mapping = aes(x = long, y = lat,
                          group = group, fill = region))

p + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    guides(fill = FALSE)
```





## Map U.S. state-level data {.build .faster}
- Let’s get the `election` data on to the map 
 
```{r, echo=T, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
election$region <- tolower(election$state)
us_states_elec <- left_join(us_states, election)
```

- In the map data, `us_states`, the state names (in a variable named `region`) are in lower case. 

- Here we can create a variable in the `election` data frame to correspond to this, using the `tolower()` function to convert the state names.

- It is important to know your data and variables well enough to check that they have merged properly.

  - FIPS code is useful in joining the data.



## Map U.S. state-level data {.build .faster}
 
```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
            aes(x = long, y = lat,
                group = group, fill = party))

p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

```

- We use our `party` colors for the `fill`.


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
            aes(x = long, y = lat,
                group = group, fill = party))

p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 
```




## Map U.S. state-level data {.build .faster}
 
```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p2 <- p1 + scale_fill_manual(values = party_colors) +
    labs(title = "Election Results 2016", fill = NULL)

p2 + theme_map() 


```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p2 <- p1 + scale_fill_manual(values = party_colors) +
    labs(title = "Election Results 2016", fill = NULL)

p2 + theme_map() 

```


## Map U.S. state-level data {.build .faster}
- To the `fill` aethetic, let’s try a continuous measure, such as the percentage of the vote received by Donald Trump (`pct_trump`). 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
             mapping = aes(x = long, y = lat, group = group, fill = pct_trump))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p1 + labs(title = "Trump vote") + theme_map() + labs(fill = "Percent")

```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
             mapping = aes(x = long, y = lat, group = group, fill = pct_trump))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p1 + labs(title = "Trump vote") + theme_map() + labs(fill = "Percent")

```




## Map U.S. state-level data {.build .faster}

- Blue is not the color we want here.

- The color gradient runs in the wrong direction.

- Let's fix these problems using `scale_fill_gradient()`:

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p2 <- p1 + scale_fill_gradient(low = "white", high = "#CB454A") +
        labs(title = "Trump vote") 
p2 + theme_map() + labs(fill = "Percent")

```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p2 <- p1 + scale_fill_gradient(low = "white", high = "#CB454A") +
        labs(title = "Trump vote") 
p2 + theme_map() + labs(fill = "Percent")

```


## Map U.S. state-level data {.build .faster}

- For election results, we might prefer a gradient that diverges from a midpoint. 
  - The `scale_gradient2()` function gives us a blue-red spectrum that passes through white by default. 
  - We can also re-specify the mid-level color along with the `high` and `low` colors. 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
             mapping = aes(x = long, y = lat, group = group, fill = d_points))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p2 <- p1 + scale_fill_gradient2() + labs(title = "Winning margins") 
p2 + theme_map() + labs(fill = "Percent")

```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = us_states_elec,
             mapping = aes(x = long, y = lat, group = group, fill = d_points))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

p2 <- p1 + scale_fill_gradient2()

p2 + theme_map() + labs(fill = "Percent", title = "Winning margins")

```


## Map U.S. state-level data {.build .faster}

- From the `scale_gradient2()` function, we can also re-specify the mid-level color along with the `high` and `low` colors. 

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p3 <- p1 + scale_fill_gradient2(low = "red", 
                                mid = scales::muted("purple"),
                                high = "blue", 
                                breaks = c(-25, 0, 25, 50, 75)) 
p3 + theme_map() + labs(fill = "Percent", title = "Winning margins")

```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p3 <- p1 + scale_fill_gradient2(low = "red", 
                                mid = scales::muted("purple"),
                                high = "blue", 
                                breaks = c(-25, 0, 25, 50, 75)) 
p3 + theme_map() + labs(fill = "Percent", title = "Winning margins")

```





## Map U.S. state-level data {.build .faster}

- If you take a look at the gradient scale for this first “purple America” map, `p3`, you’ll see that it extends very high on the Blue side.
  - This is because Washington DC is included in the data.
  
- If we omit Washington DC, we’ll see that our color scale shifts.

```{r, echo=T, eval=F, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = subset(us_states_elec,
                           region %nin% "district of columbia"),
             aes(x = long, y = lat, group = group, fill = d_points))

p3 <- p1 + scale_fill_gradient2(low = "red", 
                                mid = scales::muted("purple"),
                                high = "blue", 
                                breaks = c(-25, 0, 25, 50, 75)) 
p3 + theme_map() + labs(fill = "Percent", title = "Winning margins", caption = "DC is omitted.")
```


## Map U.S. state-level data 

```{r, echo=F, eval=T, fig.align='center', fig.height=5.5, message=FALSE, warning=FALSE}
p0 <- ggplot(data = subset(us_states_elec,
                           region %nin% "district of columbia"),
             aes(x = long, y = lat, group = group, fill = d_points))

p3 <- p1 + scale_fill_gradient2(low = "red", 
                                mid = scales::muted("purple"),
                                high = "blue", 
                                breaks = c(-25, 0, 25, 50, 75)) 
p3 + theme_map() + labs(fill = "Percent", title = "Winning margins", caption = "DC is omitted.")

```



