---
title: "DANL 310 Lecture 16"
subtitle: "Practice Session"
author: "Byeong-Hak Choe"
institute: "SUNY Geneseo"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---
   

# Set Up Your Project and Load Libraries

To begin we must load some libraries we will be using. If we do not load them, R will not be able to find the functions contained in these libraries. The tidyverse includes ggplot and other tools. We also load the socviz and gapminder libraries.

```{r setup, include=FALSE}

## By defult, show code for all chunks in the knitted document,
## as well as the output. To override for a particular chunk
## use echo = FALSE in its options.
knitr::opts_chunk$set(echo = TRUE) 

## Set the default size of figures
knitr::opts_chunk$set(fig.width=8, fig.height=5)  

## Load the libraries we will be using
library(skimr)
library(tidyverse)

```

<br>


# Making a table for summary statistics

Here is the data:
```{r}
library(skimr)

oj <- read.table(
  'https://bcecon.github.io/dominick_oj.csv',
  sep = ',',
  header = TRUE,
  stringsAsFactor = TRUE
)

oj
nvars <- format(round(ncol(oj), 0), nsmall=0, big.mark=",") 
nobs <- format(round(nrow(oj), 0), nsmall=0, big.mark=",") 
```



<br> 
The number of variables is `r nvars`; the number of observations is `r nobs`.

<br>

Here is the making of the table for descriptive statistics for the data using `Hmisc::describe`:



```{r}
Hmisc::describe(oj) %>% Hmisc::html()
```

<br>

Here is the making of the table for descriptive statistics for the data using `skimr::skim`:

```{r}
skim(oj)
```

<br>

Here is the making of the table for descriptive statistics for the data using `rstatix::get_summary_stats`:

```{r}
rstatix::get_summary_stats(oj, type = "full")
```


<br>

Here is the making of the table for descriptive statistics for the data using `dplyr::summarize()`:

```{r, results = 'asis', message=F}
library(papeR)
library(knitr)
kable(summarize(oj, type = "numeric"))
kable(summarize(oj, type = "factor", variables = "brand", cumulative = TRUE))
kable(summarize(oj, type = "numeric", group = "brand", test = FALSE))
```


```{r}
skimmed <- skim(oj)
skimmed <- skimmed %>% 
  select(-factor.ordered, -complete_rate)

skimmed <- as_tibble(skimmed)
skimmed_chr <- skimmed[1,2:5]
skimmed_chr[1,4] <- "dominicks: 9649,  minute.maid: 9649,  tropicana: 9649"

skimmed_numeric <- skimmed[2:4,c(2:3,5:13)]
skimmed_numeric[,4:10] <- round(skimmed_numeric[,4:10], 2)
skimmed_numeric <- skimmed_numeric %>% 
  select(-factor.top_counts)
```

<br>

Here is the summary of the factor variable in `oj`:

```{r skim_chr, eval=TRUE, echo=TRUE, fig.align='center', out.height="75%"}
knitr::kable(skimmed_chr, "simple",   # from skimr::skim() 
             col.names = c('Variable', 'n_Missing', 'n_Categories', 'Counts'))  
```


<br>

Here is the summary of the numeric variables in `oj`:

```{r skim_n, eval=TRUE, echo=TRUE, fig.align='center', out.height="75%"}
knitr::kable(skimmed_numeric, "simple",   # from skimr::skim()
             col.names = c('Variable', 'n_Missing', 'Mean', 'SD',
                           'Min', 'Q1', 'Median','Q3', 'Max', 'Hist'))
```


