---
title: Lecture 1
subtitle: Syllabus, Course Outline, and Introduction
format:
  clean-revealjs:
    self-contained: false
    chalkboard: true
    incremental: true
    code-annotations: hover
    scrollable: false

    # logo: logo-title-slide.png
author:
  - name: Byeong-Hak Choe
    email: bchoe@geneseo.edu
    affiliations: SUNY Geneseo
date: 2024-08-26
execute: 
  eval: true
  echo: true
callout-icon: false

from: markdown+emoji
include-after-body: target-hover.html # effect.html

# bibliography: refs.bib
---


```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)


theme_set(theme_fivethirtyeight()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.5),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.5),
                               margin = margin(0,10,0,0)),
                axis.text.x = element_text(size = rel(1.5)),
                axis.text.y = element_text(size = rel(1.5)),
                strip.text = element_text(size = rel(1.5)),
                legend.position = "top",
                legend.text = element_text(size = rel(1.5)),
                legend.title = element_text(size = rel(1.5))
                )
          )
```


# Instructor {background-color="#1c4982"}

## Instructor
### Current Appointment & Education

- Name: Byeong-Hak Choe.
- Assistant Professor of Data Analytics and Economics, School of Business at SUNY Geneseo.


- Ph.D. in Economics from University of Wyoming.
- M.S. in Economics from Arizona State University.
- M.A. in Economics from SUNY Stony Brook.
- B.A. in Economics & B.S. in Applied Mathematics from Hanyang University at Ansan, South Korea.
  - Minor in Business Administration.
  - Concentration in Finance.
  


## Instructor
### Economics and Data Science

- Choe, B.H., 2021. "Social Media Campaigns, Lobbying and Legislation: Evidence from #climatechange and Energy Lobbies."

- Question: To what extent do social media campaigns compete with fossil fuel lobbying on climate change legislation?

- Data include:
  - 5.0 million tweets with #climatechange/#globalwarming around the globe;
  - 12.0 million retweets/likes to those tweets;
  - 0.8 million Twitter users who wrote those tweets;
  - 1.4 million Twitter users who retweeted or liked those tweets;
  - 0.3 million US Twitter users with their location at a city level;
  - Firm-level lobbying data (expenses, targeted bills, etc.). 



## Instructor
### Economics and Data Science

- Choe, B.H. and Ore-Monago, T., 2024. "Governance and Climate Finance in the Developing World"

- **Climate finance** refers to the financial resources allocated for mitigating and adapting to climate change, including support for initiatives that reduce greenhouse gas emissions and enhance resilience to climate impacts.
  - We focus on transnational financing that rich countries provide poor countries with financial resources, in order to help them adapt to climate change and mitigate greenhouse gas (GHG) emissions.
  - Since the GHG emissions in developing countries are rapidly growing, it is crucial to assess the effectiveness of climate finance.
  - Poor governance can be significant barriers to emissions reductions within these countries.


## Instructor
### Economics and Data Science

- Choe, B.H. and Ore-Monago, T., 2024. "Governance and Climate Finance in the Developing World"

- Data include:
  - Global climate finance data (e.g., donors, recipients, characteristics of climate change projects)
  - World Bank Governance Indicators over the years (e.g., government effectiveness, voice and accountability, political stability and absence of violence/terrorism, regulatory quality, rule of law, control of corruption)
  - Various economic indicators (e.g., trade pattern of low carbon technology products, macroeconomic risks, energy)


## Instructor
### Economics and Data Science

- Choe, B.H., James, Alex and Newbold, Steve, 2025. "Estimating the Value of Statistical Life (VSL) through Big Data"

- VSL is the monetary value associated with reducing the risk of death. 
  - How much value would that be? How can we measure it?




# Syllabus {background-color="#1c4982"}

## Syllabus

### Email, Class & Office Hours

- Email: [bchoe@geneseo.edu](bchoe@geneseo.edu)


- Class Homepage: 
  i. [https://mylearning.suny.edu/d2l/home/1513935](https://mylearning.suny.edu/d2l/home/1513935){target="_blank"}
  i. [http://bcdanl.github.io](http://bcdanl.github.io){target="_blank"}
  
  
- Office: South Hall 301
- Office Hours: 
  - In-person: Mondays and Wednesdays 1:30 P.M.--3:00 P.M.
  - Online: by appointment via email.


## Syllabus

### Course Description

- This course offers a practical introduction to the data analytics process and methods. 
- The goal is to help you unlock the potential of data analysis and enhance your ability to transform data into a powerful decision-making tool. 
- You will develop foundational data analytics skills that will prepare you for a career or further studies in more advanced data analytics topics. 
- The course covers: (1) an introduction to data analytics thinking, (2) exploratory data analysis through data wrangling and visualization, and (3) an introduction to data science, social media analytics, and artificial intelligence (AI). 
- Throughout the course, you will gain hands-on experience with the Python and R programming languages, along with its associated data analysis libraries and AI tools.




## Syllabus

### Course Learning Outcomes

- Grasp the basic principles of data analytics, including data types and data processing.
- Gain introductory experience with programming languages commonly used in data analytics, such as Python or R.
- Develop the ability to create and interpret various types of data visualizations.
- Enhance critical thinking skills by learning to ask relevant questions and draw insights from data.
- Apply data analytics techniques to solve real-world problems in various domains.


## Syllabus

### Reference Materials

- [**R for Data Science** (2nd Edition)](https://r4ds.hadley.nz){target="_blank"} by Hadley Wickham \& Garrett Grolemund.
- [**Python for Data Science**](https://aeturrell.github.io/python4DS){target="_blank"} by Arthur Turrell et. al.
- [**Coding for Economists**](https://aeturrell.github.io/coding-for-economists){target="_blank"} by Arthur Turrell.
- [**Statistical Inference via Data Science: A ModernDive into R and the Tidyverse**](https://moderndive.com){target="_blank"} by Chester Ismay and Albert Y. Kim.



## Syllabus

### Course Requirements
- **Homework**: There will be six homework assignments.
- **Quiz**: There will be in-class quizzes.
- **Class Participation**: You are encouraged to ask questions and discuss topics during the class time.
- **Exam**: There will be two short midterm exams and one comprehensive final exam.
- **Team presentation**: There will be a team presentation.
  
  
  
## Syllabus
### Team Presentations

- Each team will present on data storytelling with visualization. 

- Data storytelling with visualization is the art of communicating complex data insights in a clear, engaging, and impactful way by blending data analysis, visual design, and narrative techniques. 

- It goes beyond simply showing charts and graphs; it involves crafting a compelling story that guides the audience through the data, highlights key findings, and effectively conveys the intended message. 


## Syllabus
### Team Presentations
- Key aspects of data storytelling include:
  - **Clarity**: Make sure the visuals are easy to understand and focus on the most important information.
  - **Structure**: Just like a story, start with an introduction, present the main points, and finish with a conclusion.
  - **Engagement**: Tailor the story to your audience so it is relevant and interesting to them.




## Syllabus

### Class Schedule

- There will be tentatively 42 class sessions.
  - **The midterm exam 1** is scheduled on **October 9, 2024, Wednesday**, during the class time.
  - **The midterm exam 2** is scheduled on **November 8, 2024, Friday**, during the class time.
  - **The team presentation** is scheduled during the first week of December, 2024.
  - **The final exam** is scheduled on **December 16, 2024, Monday, noon--2:00 P.M.**




## Syllabus
### Course Contents

```{r, out.width='100%', fig.align='center'}
#| eval: true
#| echo: false
text_tbl <- data.frame(
  Week = c("1", "2", "3", "4", "5", "6", "7"),
  Contents = c("Course Outlines; Overview",
"Introduction",
"Data Analytics Careers",
"Basic Tools",
"Basic Tools",
"Data Preparation and Management", 
"Data Preparation and Management"),
  `HW Exam` = c("", "", "HW 1", "", "", "HW 2", "Midterm I")
  )


# Create a DT datatable without search box and 'Show entries' dropdown
DT::datatable(text_tbl, rownames = FALSE,
              options = list(
  dom = 't', # This sets the DOM layout without the search box ('f') and 'Show entries' dropdown ('l')
  paging = FALSE, # Disable pagination
  columnDefs = list(list(
    targets = "_all", # Applies to all columns
    orderable = FALSE # Disables sorting
  ))
), callback = htmlwidgets::JS("
  // Change header background and text color
  $('thead th').css('background-color', '#1c4982');
  $('thead th').css('color', 'white');

  // Loop through each row and alternate background color
  $('tbody tr').each(function(index) {
    if (index % 2 == 0) {
      $(this).css('background-color', '#d1dae6'); // Light color for even rows
    } else {
      $(this).css('background-color', '#9fb2cb'); // Dark color for odd rows
    }
  });

  // Set text color for all rows
  $('tbody tr').css('color', 'black');

  // Add hover effect
  $('tbody tr').hover(
    function() {
      $(this).css('background-color', '#607fa7'); // Color when mouse hovers over a row
    }, 
    function() {
      var index = $(this).index();
      if (index % 2 == 0) {
        $(this).css('background-color', '#d1dae6'); // Restore even row color
      } else {
        $(this).css('background-color', '#9fb2cb'); // Restore odd row color
      }
    }
  );
")
)

```




## Syllabus
### Course Contents

```{r, out.width='100%', fig.align='center'}
#| eval: true
#| echo: false
text_tbl <- data.frame(
  Week = c("8", "9", "10", "11"),
  Contents = c(
"Data Visualization",
"Data Visualization",
"Data Visualization",
"Data Visualization"
),
  `HW Exam` = c("", "HW 3", "", "HW 4 & Midterm II")
  )


# Create a DT datatable without search box and 'Show entries' dropdown
DT::datatable(text_tbl, rownames = FALSE,
              options = list(
  dom = 't', # This sets the DOM layout without the search box ('f') and 'Show entries' dropdown ('l')
  paging = FALSE, # Disable pagination
  columnDefs = list(list(
    targets = "_all", # Applies to all columns
    orderable = FALSE # Disables sorting
  ))
), callback = htmlwidgets::JS("
  // Change header background and text color
  $('thead th').css('background-color', '#1c4982');
  $('thead th').css('color', 'white');

  // Loop through each row and alternate background color
  $('tbody tr').each(function(index) {
    if (index % 2 == 0) {
      $(this).css('background-color', '#d1dae6'); // Light color for even rows
    } else {
      $(this).css('background-color', '#9fb2cb'); // Dark color for odd rows
    }
  });

  // Set text color for all rows
  $('tbody tr').css('color', 'black');

  // Add hover effect
  $('tbody tr').hover(
    function() {
      $(this).css('background-color', '#607fa7'); // Color when mouse hovers over a row
    }, 
    function() {
      var index = $(this).index();
      if (index % 2 == 0) {
        $(this).css('background-color', '#d1dae6'); // Restore even row color
      } else {
        $(this).css('background-color', '#9fb2cb'); // Restore odd row color
      }
    }
  );
")
)

```




## Syllabus
### Course Contents

```{r, out.width='100%', fig.align='center'}
#| eval: true
#| echo: false
text_tbl <- data.frame(
  Week = c("12", "13", "14", "15", "16"),
  Contents = c(
"Data Science",
"Social Media Analytics",
"Artificial Intelligence",
"Team Presentations", "Wrap-up"
),
  `HW Exam` = c("", "HW 5", "", "HW 6", "Final")
  )


# Create a DT datatable without search box and 'Show entries' dropdown
DT::datatable(text_tbl, rownames = FALSE,
              options = list(
  dom = 't', # This sets the DOM layout without the search box ('f') and 'Show entries' dropdown ('l')
  paging = FALSE, # Disable pagination
  columnDefs = list(list(
    targets = "_all", # Applies to all columns
    orderable = FALSE # Disables sorting
  ))
), callback = htmlwidgets::JS("
  // Change header background and text color
  $('thead th').css('background-color', '#1c4982');
  $('thead th').css('color', 'white');

  // Loop through each row and alternate background color
  $('tbody tr').each(function(index) {
    if (index % 2 == 0) {
      $(this).css('background-color', '#d1dae6'); // Light color for even rows
    } else {
      $(this).css('background-color', '#9fb2cb'); // Dark color for odd rows
    }
  });

  // Set text color for all rows
  $('tbody tr').css('color', 'black');

  // Add hover effect
  $('tbody tr').hover(
    function() {
      $(this).css('background-color', '#607fa7'); // Color when mouse hovers over a row
    }, 
    function() {
      var index = $(this).index();
      if (index % 2 == 0) {
        $(this).css('background-color', '#d1dae6'); // Restore even row color
      } else {
        $(this).css('background-color', '#9fb2cb'); // Restore odd row color
      }
    }
  );
")
)

```



## Syllabus
### Grading

$$
\begin{align}
(\text{Total Percentage Grade}) =&\;\, 0.05\times(\text{Attendance}) \notag\\
&\,+\, 0.15\times(\text{Quiz & Class Participation})\notag\\ 
& \,+\, 0.15\times(\text{Homework})\notag\\  
&\,+\, 0.15\times(\text{Presentation})\notag\\
& \,+\, 0.50\times(\text{Exam}).\notag
\end{align}
$$



## Syllabus
### Grading

- You are allowed up to 6 absences without penalty. 
  - Send me an email if you have standard excused reasons (illness, family emergency, transportation problems, etc.).

- For each absence beyond the initial six, there will be a deduction of 1% from the Total Percentage Grade.

- The single lowest homework score will be dropped when calculating the total homework score. 
  - Each homework except for the homework with the lowest score accounts for 20% of the total homework score.


## Syllabus
### Grading

$$
\begin{align}
&(\text{Total Exam Score}) \\
=\, &\text{max}\,\left\{0.50\times(\text{Midterm Exam Score}) \,+\, 0.50\times(\text{Final Exam Score})\right.,\notag\\ 
&\qquad\;\,\left.0.25\times(\text{Midterm Exam Score}) \,+\, 0.75\times(\text{Final Exam Score})\right\}.\notag
\end{align}
$$



- The total exam score is the maximum between 
  1. the simple average of the midterm exam score and the final exam score and
  2. the weighted average of them with one-fourth weight on the midterm exam score and three-third weight on the final exam score:


## Syllabus
### Make-up Policy

-  Make-up exams will not be given unless you have either a medically verified excuse or an absence excused by the University.

- If you cannot take exams because of religious obligations, notify me by email at least two weeks in advance so that an alternative exam time may be set.

- A missed exam without an excused absence earns a grade of zero.

- Late submissions for homework assignment will be accepted with a penalty. 

- A zero will be recorded for a missed assignment.


## Syllabus
### Academic Integrity and Plagiarism

-  All homework assignments and exams must be the original work by you. 

- Examples of academic dishonesty include:
  - *representing the work, thoughts, and ideas of another person as your own*
  - *allowing others to represent your work, thoughts, or ideas as theirs*, and
  - *being complicit in academic dishonesty by suspecting or knowing of it and not taking action*.


- Geneseo’s Library offers frequent workshops to help you understand how to **paraphrase**, **quote**, and **cite** outside sources properly. 
  - See [https://www.geneseo.edu/library/library-workshops](https://www.geneseo.edu/library/library-workshops){target="_blank"}.


## Syllabus
### Artificial Intelligence (AI) Policy

- Unless AI tools are explicitly permitted for homework or in-class quizzes, you must complete your work independently. 

- This means you should not use tools like ChatGPT for any aspect of our coursework. 

- Such use is a form of academic dishonesty. Use of such tools is not only cheating, it will also cheat you of the opportunity to learn and develop your own skills. 

- While AI will undoubtedly play important roles in our future society, you will be better able to utilize AI if you have developed your own critical thinking, writing, and analytical skills by doing your own work. 

- If you have any questions about this, please ask.



## Syllabus
### Accessibility

-  The Office of Accessibility will coordinate reasonable accommodations for persons with physical, emotional, or cognitive disabilities to ensure equal access to academic programs, activities, and services at Geneseo.

- Please contact me and the Office of Accessibility Services for questions related to access and accommodations.



## Syllabus
### Career Design

- To get information about career development, you can visit the Career Development Events Calendar ([https://www.geneseo.edu/career_development/events/calendar](https://www.geneseo.edu/career_development/events/calendar){target="_blank"}).

- You can stop by South 112 to get assistance in completing your Handshake Profile [https://app.joinhandshake.com/login](https://app.joinhandshake.com/login){target="_blank"}.
  - Handshake is ranked #1 by students as the best place to find full-time jobs.
  - 50% of the 2018-2020 graduates received a job or internship offer on Handshake.
  - Handshake is trusted by all 500 of the Fortune 500.

