---
output:
  pagedown::html_paged:
    css: ["css/one-column-paged-academic.css",  "https://use.fontawesome.com/releases/v5.7.2/css/all.css"]
    number_sections: false
links-to-footnotes: true
knit: pagedown::chrome_print
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(lubridate)
library(glue)
library(kableExtra)

######## CREATE HELPER FUNCTIONS
# function which takes a name and replaces the first name with initial and dot
replace_first_name_with_initial_and_dot <- function(author_name) {
  str_replace(author_name, "(?<=\\S)\\S+", "\\.")
}

# function for doing this for an entry with a number of authors
replace_first_names_in_list_with_initial_and_dot <- function(authors){
  authors_split <- str_split(authors, ",") %>% 
    map(str_trim)
  
  authors_split[[1]] %>% map_chr(replace_first_name_with_initial_and_dot) %>% 
    paste0(collapse = ", ")
}

# function for adding leading zero's when there's more than 10 entries
padding_function <- function(cv_entries_tibble, letter_to_pad){
  if(nrow(cv_entries_tibble) > 9){
    cv_entries_tibble %>% 
      arrange(date) %>%
  mutate(number = str_pad(row_number(), width = 2, side = "left", pad = "0"),
         number = str_c(letter_to_pad, number)) %>%
  arrange(desc(date))
  } else {
    cv_entries_tibble %>% 
      arrange(date) %>%
  mutate(number = str_pad(row_number(), width = 1, side = "left", pad = "0"),
         number = str_c(letter_to_pad, number)) %>%
  arrange(desc(date))
  }
}

# function for making a year blank when it's the same as the one preceding it

####### READ IN DATA
#### publications #####
publications <- read_csv("_content/publications.csv") %>% 
  mutate(authors_full = map_chr(authors_full, replace_first_names_in_list_with_initial_and_dot)) %>% 
  mutate(authors_full = str_replace(authors_full, "U\\. Lyngs", "*U\\. Lyngs*")) %>%    # make my name italic
  mutate(year = year(date)) %>% 
  mutate(venue_abbrev = if_else(!is.na(venue_abbrev), str_c(venue_abbrev, ": "), "")) # make the abbrevation an empty string if NA; add a colon and a space after if it if there is one


#### publications #####
publications <- read_csv("_content/publications.csv")

#### all other entries #####
cv_entries <- read_csv("_content/cv_entries.csv") %>%
  mutate(year_begin = year(date)) %>%
  mutate(year_end = case_when(
    date_end == "present" ~ date_end,
    !is.na(date_end) ~ str_sub(date_end, 1, 4),
    is.na(date_end) ~ date_end
  )) %>%
  mutate(year = ifelse((is.na(year_end) | year_begin == year_end),
                      year_begin,
                      str_c(year_begin, " --- ", year_end)))


blank_year_when_repeated <- function(cv_entry_tibble){
  cv_entry_tibble %>% 
    mutate(row_number = row_number()) %>% 
    mutate(year = as.character(year),
           year = case_when(
             row_number == 1 ~ year,
             year == lag(year) ~ "",
             year != lag(year) ~ year
           )) %>% 
    select(-row_number)
}


```


```{css}
:root{
  --last-updated: " \00a0 \00a0 \00a0 \00a0 Last updated: April 2022";
}
```



:::::: {.title-and-contact}

::: {.title}
<h1 class="title">Byeong-Hak Choe</h1>
:::

::: {.contact-info}
bchoe@geneseo.edu  \
https://bcecon.github.io
:::

::::::

## Research summary
:::::: {.research-interests}

::: {.spacer}
:::

::: {.interests}
I view myself as an **applied economist** with interest in the field of **environmental economics** and a particular focus on **climate change policy**. Methodologically, I make use of **econometric/machine learning methods and data science tools** to conduct empirical analysis. I also use computational methods to solve **integrated assessment models (IAM) of climate change** as well as theoretical economic models, such as **dynamic contracts**.

:::
::::::

# Current Position
```{r}
cv_entries %>% 
  filter(type == 'current_positions', is.na(exclude)) %>%
  mutate(what = str_c("<span class='cv-entry-title'>", what, "</span><br>", 
                      where)) %>% 
  select(year, what) %>% 
  kbl(padding = 0, col.names = NULL, escape = FALSE) %>%
  kable_styling( bootstrap_options = c("hover"))  %>% 
  column_spec(1, width = "15%")

```


# Education
```{r}
cv_entries %>% 
  filter(type == 'education') %>% 
  mutate(what = str_c( "<span class='cv-entry-title'>", 
                              what, "</span>, ", where,
                              "<br>", 
                       ifelse(is.na(additional_info), "", additional_info) 
                       , "<br>", 
                       ifelse(is.na(additional_info2), "", additional_info2))) %>% 
  select(year, what) %>% 
  kbl(padding = 0, col.names = NULL, escape = FALSE) %>%
  kable_styling( bootstrap_options = c("hover", "condensed"))   %>% 
  column_spec(1, width = "15%")

```



# Teaching
```{r}
cv_entries %>% 
  filter(type == 'teaching') %>% 
  mutate(what = ifelse(!is.na(additional_info), 
    str_c("<span class='cv-entry-title'>", what, "</span><br class='work-break'>", where, "<br>", additional_info),
    str_c("<span class='cv-entry-title'>", what, "</span><br class='work-break'>", where))) %>% 
  select(year, what) %>% 
  kbl(padding = 0, col.names = NULL, escape = FALSE) %>%
  kable_styling( bootstrap_options = c("hover", "condensed"))  %>% 
  column_spec(1, width = "15%") 
  
```


# Service
```{r}
cv_entries %>% 
  filter(type == 'service') %>% 
  mutate(where = ifelse(!is.na(url), str_c("[", where, "](", url, ")"), where)) %>% 
  mutate(what = ifelse(!is.na(additional_info),
                       str_c("<span class='cv-entry-title'>", what,
                      "</span><br class='work-break'>", 
                      where),
                      str_c("<span class='cv-entry-title'>", what,
                      "</span><br class='work-break'>", 
                      where))) %>% 
  arrange(desc(date_end), desc(date)) %>% 
  select(year, what) %>% 
  kbl(padding = 0, col.names = NULL, escape = FALSE) %>%
  kable_styling( bootstrap_options = c("hover", "condensed")) %>% 
  column_spec(1, width = "15%")  

```



# Work Experience
```{r}
cv_entries %>% 
  filter(type == 'work', is.na(website)) %>% 
  mutate(additional_info = replace_na(additional_info, "")) %>% 
  mutate(what = str_c("<span class='cv-entry-title'>", what, "</span><br class='work-break'>", where, "<br>", additional_info)) %>% 
  arrange(desc(date)) %>% 
  select(year, what) %>% 
  kbl(padding = 0, col.names = NULL, escape = FALSE) %>%
  kable_styling( bootstrap_options = c("hover", "condensed"))   %>% 
  column_spec(1, width = "15%")
```


# Professional activities
```{r}
cv_entries %>% 
  filter(type == 'prof-act') %>% 
  mutate(what = str_c("<span class='cv-entry-title'>", what, "</span>, ", where)) %>% 
  select(year_begin, what) %>% 
  kbl(padding = 0, col.names = NULL, align = c("l", "l"), escape = FALSE) %>%
  kable_styling( bootstrap_options = c("hover", "condensed")) %>% 
  column_spec(1, width = "15%")  
```


# Hornors & Awards
```{r}
cv_entries %>% 
  filter(type == 'awards') %>%
  filter(!(website == "n") | is.na(website)) %>% 
  mutate(what = if_else(
    !is.na(where),
    str_c("<span class='cv-entry-title'>", what, "</span>, ", where),
    str_c("<span class='cv-entry-title'>", what, "</span>"))
  ) %>% 
  blank_year_when_repeated() %>% 
  select(year_begin, what) %>% 
  kbl(padding = 0, col.names = NULL, align = c("l", "l"), escape = FALSE) %>%
  kable_styling( bootstrap_options = c("hover", "condensed"))   %>% 
  column_spec(1, width = "15%")
```
 

# Grants and funding
```{r}
cv_entries %>% 
  filter(type == 'minor_grants' | type == 'major_grants',
         is.na(exclude)) %>%
  filter(!(website == "n") | is.na(website)) %>% 
  mutate(what = if_else(
    !is.na(additional_info),
    str_c("<span class='cv-entry-title'>", what, "</span>, ", where, "<br>", additional_info),
    str_c("<span class='cv-entry-title'>", what, "</span>, ", where))
  ) %>% 
  blank_year_when_repeated() %>% 
  select(year, what) %>% 
  kbl(padding = 0, col.names = NULL, align = c("l", "l"), escape = FALSE)  %>%
  kable_styling( bootstrap_options = c("hover", "condensed")) %>% 
  column_spec(1, width = "15%")   
```




# Miscelaneous
## Computing skills
```{r}
cv_entries %>% 
  filter(type == 'technical') %>% 
  mutate(icon = case_when(
    str_detect(what, "data analysis") ~ "fa-bar-chart",
    str_detect(what, "research") ~ "fa-flask",
    str_detect(what, "Web") ~ "fa-laptop"
  )) %>% 
  mutate(what = str_c("<i class='fa fa-lg ", icon, "'></i> &nbsp;<span class='cv-entry-title'>", what, "</span> ", additional_info)) %>% 
  select(type, what) %>% 
  mutate(type = "") %>% 
  kbl(padding = 0, col.names = NULL, escape=FALSE) %>%
  kable_styling( bootstrap_options = c("hover", "condensed")) 
```



## Skills & volunteering
```{r}
cv_entries %>%
  filter(type == 'volunteering') %>%
  mutate(what = str_c("<span class='cv-entry-title'>", what, "</span><br class='work-break'>", where)) %>%
  mutate(what = ifelse(!is.na(additional_info), str_c(what, "<br>", additional_info), what)) %>%
  arrange(desc(date_end)) %>% 
  select(year, what) %>%
  kbl(padding = 0, col.names = NULL, escape = FALSE)  %>%
  kable_styling( bootstrap_options = c("hover", "condensed")) %>% 
  column_spec(1, width = "15%")
```

