---
title: Homework Assignment 1 - Example Answers
subtitle: ""
date: last-modified
from: markdown+emoji
comments: false
code-fold: false

execute: 
  warning: false
  message: false
  fig-width: 9
  fig-height: 7
  fig-align: center
  eval: false
---
```{r}
#| include: false

library(tidyverse)
library(skimr)
library(DT)
library(hrbrthemes)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))

x <- c(5, 7, 6, 9, 100, 8, 5, 7, 6)
q1 <- quantile(x, 0.25)
q3 <- quantile(x, 0.75)
iqr_value <- q3 - q1
```


# R Basics

## Question 1
Compute the weighted mean of the vector `scores <- c(85, 90, 88, 92, 87)` with corresponding weights `weights <- c(1, 2, 1, 1, 3)`.

**Answer**:
```{r}
# Define the scores vector
scores <- c(85, 90, 88, 92, 87)

# Define the corresponding weights vector
weights <- c(1, 2, 1, 1, 3)

# Compute the weighted mean by summing the products of scores and weights and dividing by the sum of weights
weighted_mean <- sum(scores * weights) / sum(weights)
```

Therefore the weighted mean is **88.25**.

<br><br>



## Question 2
Compute the interquartile range (IQR) of the following vector `x <- c(5, 7, 6, 9, 100, 8, 5, 7, 6)` manually, without using the `IQR()` function.

**Answer**:
```{r}
# Define the data vector
x <- c(5, 7, 6, 9, 100, 8, 5, 7, 6)

# Compute the first quartile (Q1)
q1 <- quantile(x, 0.25)

# Compute the third quartile (Q3)
q3 <- quantile(x, 0.75)

# Calculate the interquartile range (IQR)
iqr_value <- q3 - q1
```

Therefore the IQR is **2**.


<br><br>

## Question 3
Detect the outliers in the vector `x <- c(5, 7, 6, 9, 100, 8, 5, 7, 6)` using the 1.5*IQR rule.

**Answer**:
```{r}
# Calculate the lower bound for outliers using Q1 and the IQR
lower_bound <- q1 - 1.5 * iqr_value

# Calculate the upper bound for outliers using Q3 and the IQR
upper_bound <- q3 + 1.5 * iqr_value

# Find any values in the data vector that are below the lower bound (lower outliers)
outliers_lower <- x[x < lower_bound]

# Find any values in the data vector that are above the upper bound (upper outliers)
outliers_upper <- x[x > upper_bound]
```

There is one single value, **100**, that belongs to `outliers_upper`. There is no lower outlier.



<br><br>

## Question 4
Calculate the skewness of the vector `x <- c(3, 5, 8, 12, 14, 15, 18, 20)` without using any external R package. Skewness is defined as:
$$
\text{Skewness} \,=\, \frac{n}{(n-1)(n-2)}\sum\left(\frac{x_{i}-\bar{x}}{s}\right)^{3}
$$
where $s$ is the standard deviation of the vector `x`.


**Answer**:
```{r}
# Define the data vector
x <- c(3, 5, 8, 12, 14, 15, 18, 20)

# Calculate the number of elements (n) in the vector
n <- length(x)

# Compute the mean of the vector
mean_x <- mean(x)

# Compute the standard deviation of the vector
sd_x <- sd(x)

# Calculate skewness using the formula provided
skewness <- ( n / ( (n-1) * (n-2) ) ) * sum( ( (x - mean_x)/sd_x )^3 )
```

Therefore `skewness` is approximately `-0.2337`.

<br><br>

## Question 5
Calculate `mode_v`, the mode of a numeric vector `v <- c(2, 3, 5, 5, 6, 7, 3)` using the `mfv()` function provided by the R package `modeest`.

**Answer**:
```{r}
# Define the vector of data
v <- c(2, 3, 5, 5, 6, 7, 3, 5)

# Use the modeest package's mfv() function to calculate the mode
mode_v <- modeest::mfv(v)
```

To use `modeest::mfv()`, we first install the `modeest` package:

```{r}
#| eval: false

# Install the modeest package
install.packages("modeest")
```


The notation `modeest::mfv()` indicates that we are using the `mfv()` function from the `modeest` package. The `::` operator specifies which package the function or the object is coming from.

<br><br>

## Question 6
Calculate, `z`, the vector for the standardized values in the vector `x <- c(10, 20, 30, 40, 50)`. The standardized value of the individual value, $z_{i}$, is defined as:

$$
z_{i} \,=\, \frac{x_{i} - \bar{x}}{s},
$$
where\
$\quad$ - $x_{i}$: $i^{th}$ value in the vector $x$\
$\quad$ - $\bar{x}$: the mean of values in $x$\
$\quad$ - $s$: the standard deviation of values in $x$\
$\quad$ - $z_{i}$: $i^{th}$ value in the vector $z$, the vector of the standardized values in $x$.


**Answer**:
```{r}
# Define the data vector
x <- c(10, 20, 30, 40, 50)

# Calculate the standardized values (z-scores) for each element in x
z <- (x - mean(x)) / sd(x)
```



<br><br>

# Short-Answer Questions

1.	Data analytics aims to replace traditional business and economics courses. (False)
	-	Data analytics is seen as a complement to traditional business and economics courses, enhancing decision-making with data-driven insights rather than replacing foundational business knowledge.
<br><br>

2.	Data analysts are the only professionals who benefit from data analytics skills. (False)
	-	Many professionals, including marketers, financial analysts, healthcare professionals, and even educators, benefit from data analytics skills as they apply to various fields.
<br><br>

3.	Python and R are the only programming languages used in data science. (False)
	-	While Python and R are popular in data science, other languages like SQL, MATLAB, and even JavaScript are also used depending on the application.
<br><br>
	
4.	The use of generative AI requires a deep understanding of the subject matter to apply it effectively. (True)
	-	To use generative AI effectively, it is essential to have a solid understanding of the subject area to interpret and apply the results accurately.
<br><br>

5.	Machine learning algorithms need to be explicitly programmed for each task they perform. (False)
	-	Machine learning algorithms learn patterns from data, meaning they don’t need explicit programming for each task. Instead, they adapt and improve their performance based on the data provided.
<br><br>

6.	Which of the following skills is NOT typically covered in traditional business or economics classes?
	-	Finding and cleaning datasets
	  -	**Why?**: Data cleaning and preparation are often covered in data analytics but not typically in traditional business or economics courses.
<br><br>
	
7.	Which of the following is a key reason why R is widely used in data analysis?
	-	It is specifically designed for statistical computing
	  -	**Why?**: R was created for statistical analysis and is widely used because of its extensive libraries for data analysis and statistical modeling.
<br><br>

8.	Which question would you ask to analyze season ticket renewals in sports analytics?
	-	Which type of fan engages most with team merchandise?
	  -	**Why?**: This question helps to understand fan loyalty and engagement, which are important factors in predicting season ticket renewals.
<br><br>

9.	Which of the following is a benefit of using Git in software development?
	-	It tracks changes and helps manage multiple versions of a project
	  -	**Why?**: Git is a version control system that tracks changes and allows developers to manage multiple versions of a project, making collaboration more efficient.
<br><br>

10.	Which of the following is a key application of data analytics in the retail sector?
	-	Enhancing physical store layouts based on customer behavior
	  -	**Why?**: Data analytics helps retailers optimize store layouts by analyzing customer movement patterns, ultimately improving the customer experience and sales.
<br><br>

11.	Why are R and Python considered important tools for data analytics?
	-	R and Python are popular because they have a wide range of libraries and packages for data manipulation, visualization, and machine learning. They also support a large user community and are open-source, making them accessible and adaptable.
<br><br>
	
12.	Explain why understanding the output of Generative AI tools like ChatGPT is important for data analysts or data scientists.
	-	Data analysts need to interpret the output of generative AI tools to ensure that the generated information is relevant, accurate, and aligned with the specific context of their analysis. This understanding helps prevent misinterpretation and misuse of AI-generated insights.
<br><br>
	
13.	How does dynamic ticket pricing work in sports analytics?
	-	Dynamic ticket pricing adjusts the price of tickets in real-time based on demand, opponent strength, weather conditions, and other factors. This approach maximizes revenue by charging more when demand is high and less when demand is low.
<br><br>

14.	How do business intelligence (BI) tools assist in decision-making for businesses?
	-	BI tools help businesses collect, process, and visualize data to support informed decision-making. They provide insights into trends, performance metrics, and areas of improvement, enabling businesses to make data-driven decisions.
<br><br>



