---
title: "Quiz 1 - Addendum"
subtitle: Classwork 7
date: last-modified
from: markdown+emoji

execute: 
  eval: false
  echo: true
---

# Descriptive Statistics

The following provides the percentage of the correct answers for each question in Quiz 1 - Addendum:

```{r}
#| echo: false
#| eval: true

DT::datatable(data.frame(
  Question = c('data.frame-variable',
               'left_join()',
               '== (equality operator)',
               '!= (not-equal operator)',
               '| (or operator)'),
  Percentage = c(69.23, 60.53, 78.95, 89.47, 72.22)
))

```

<br><br>

# Question 1. `data.frame`
In a data frame, what form does a variable take?

**Answer**: Column

- [This slide](https://bcdanl.github.io/danl-lec/danl-101-lec-07-2024-0911.html?q=tidy#/r-basics-9) in Lecture 7 explains what makes a data.frame tidy. In our course, all given data.frames are tidy.
  - It’s been a while since we discussed variable-column, observation-row, and value-cell terminologies in a tidy data.frame, but these concepts are crucial for conducting data analysis.

- In a data.frame:
  - A variable takes a column
  - An observation takes a row
  - A value takes a cell
  
<br><br>

# Question 2. `left_join()`
```{r}
df_joined <- df_1 |> ___(df_2)
```

- Fill in the blank (___) to join the two related data.frames, `df_1` and `df_2`, in a way that keeps all observations and variables in the data.frame `df_1`.

**Answer**: `left_join`

```{r}
df_joined <- df_1 |> left_join(df_2)
```

- [This slide](https://bcdanl.github.io/danl-lec/danl-101-lec-14-2024-0930.html?q=left_join#/relational-data) in Lecture 14 and [this slide](https://bcdanl.github.io/danl-lec/danl-101-lec-12-2024-0925.html?q=relational#/relational-databases) in Lecture 12 explain **join** and `left_join()`.


<br><br>

# Question 3. `==` (Equality operator)
```{r}
df_filtered <- df |> filter(num __ 9) 
```

- Fill in the blank (___) to keep observations, in which the value of the `num` variable is equal to 9.

**Answer**: `==`

```{r}
df_filtered <- df |> filter(num == 9) 
```

- [This slide](https://bcdanl.github.io/danl-lec/danl-101-lec-14-2024-0930.html#/filter-observations-with-filter-4) and [many examples from class R scripts](https://github.com/search?q=repo%3Abcdanl%2F101-05-R%20%3D%3D&type=code) explain this equality operator.

<br><br>

# Question 4. `!=` (Not-equal operator)
```{r}
non_dec <- df |> filter(month ___ 12)
```

- Fill in the blank (___) to keep observations, in which the value of the `month` is not equal to 12 in the data.frame `df`

**Answer**: `!=`

```{r}
non_dec <- df |> filter(month != 12)
```

- [This slide](https://bcdanl.github.io/danl-lec/danl-101-lec-14-2024-0930.html#/filter-observations-with-filter-4) and [Line 63-64 in this class R script](https://github.com/search?q=repo%3Abcdanl%2F101-05-R+%21%3D&type=code) in Lecture 14 explain a not-equal operator, `!=` (not equal to).


<br><br>



# Question 5. `|` (Or operator)
```{r}
df_filtered <- df |> filter(subject == "DANL" ___ subject == "ECON")
```

- Fill in the blank (___) to keep observations, in which the value of the `subject` variable is either "DANL" or "ECON"

**Answer**: `|`

```{r}
df_filtered <- df |> filter(subject == "DANL" | subject == "ECON")
```


- [This slide](https://bcdanl.github.io/danl-lec/danl-101-lec-14-2024-0930.html#/filter-observations-with-filter-2) and [this slide](https://bcdanl.github.io/danl-lec/danl-101-lec-14-2024-0930.html#/filter-observations-with-filter-7) in Lecture 14 explain logical operation, `|` (or).

-	Please note that `|` is not the letter "I" or "l," but a vertical bar character.
  - On a keyboard, `|` is located above the *Enter/Return* key.


<br><br>


# Discussion 


<font size = "5">Welcome to our Quiz 1 Discussion Board! 👋 </font>


This space is designed for you to engage with your classmates about the material covered in Quiz 1.

Whether you are looking to delve deeper into the content, share insights, or have questions about the content, this is the perfect place for you.

If you have any specific questions for Byeong-Hak (**@bcdanl**) or peer classmate (@GitHub-Username) regarding the Quiz 1 materials or need clarification on any points, don't hesitate to ask here. 

Let's collaborate and learn from each other!

