---
title: "`filter()`, `arrange()`, and `distinct()`"
subtitle: Classwork 4
date: last-modified
from: markdown+emoji
---


## Question 1
- Find all flights that had an arrival delay of two or more hours
- Find all flights that flew to Houston (`IAH` or `HOU`)
- Find all flights that departed in summer (July, August, and September)
- Find all flights that arrived more than two hours late, but didn’t leave late
- Find all flights that departed between midnight and 6am (inclusive)
  
*Answer*:
```{r}

```

<br><br>

## Question 2
- How many flights have a missing `dep_time`?

*Answer*:
```{r}

```

<br><br>


## Question 3
- Sort flights to find the most delayed flights. 

*Answer*:
```{r}

```

<br><br>

## Question 4
- Was there a flight on every day of 2013?

*Answer*:
```{r}

```

<br><br>

# Discussion 


<font size = "5">Welcome to our Classwork 4 Discussion Board! 👋 </font>


This space is designed for you to engage with your classmates about the material covered in Classwork 4.

Whether you are looking to delve deeper into the content, share insights, or have questions about the content, this is the perfect place for you.

If you have any specific questions for Byeong-Hak (**@bcdanl**) or peer classmate (@GitHub-Username) regarding the Classwork 1 materials or need clarification on any points, don't hesitate to ask here. 

Let's collaborate and learn from each other!

