---
title: R Basics II
subtitle: Classwork 2
date: 2024-09-18
from: markdown+emoji
---


## Question 1.
- Write an R code to create a numeric vector named `numbers` containing the integers from 10 to 50.

*Answer*:
```{r}

```

<br><br>


## Question 2.
- Create a logical vector `logical_vec` that checks whether the elements of the vector `ages <- c(15, 22, 18, 24, 30)` are greater than or equal to 20.

```{r}
ages <- c(15, 22, 18, 24, 30)
```


*Answer*:
```{r}

```

<br><br>


## Question 3.
- Write an R code to assign the string "Hello, World!" to a variable named `greeting` and display its value on the Console.

*Answer*:
```{r}

```

<br><br>

## Question 4.
- Write an R code to calculate the range (difference between the maximum and minimum values) of the vector `temp <- c(22, 28, 31, 25, 29)`.

```{r}
temp <- c(22, 28, 31, 25, 29)
```


*Answer*:
```{r}

```

<br><br>


## Question 5.
- Write an R code to convert the character vector `char_vec <- c("1", "2", "3", "4")` into a numeric vector named `num_vec`.

```{r}
char_vec <- c("1", "2", "3", "4")
```

*Answer*:
```{r}

```


<br><br>

## Question 6.
- Write an R code to concatenate two character vectors, `first_names <- c("John", "Jane")` and `last_names <- c("Doe", "Smith")`, to create a vector `full_names` containing the full names using the `str_c()` function for vectorized string operations.

```{r}
first_names <- c("John", "Jane")
last_names <- c("Doe", "Smith")
```


*Answer*:
```{r}

```

<br><br>




# Discussion 


<font size = "5">Welcome to our Classwork 2 Discussion Board! 👋 </font>


This space is designed for you to engage with your classmates about the material covered in Classwork 2.

Whether you are looking to delve deeper into the content, share insights, or have questions about the content, this is the perfect place for you.

If you have any specific questions for Byeong-Hak (**@bcdanl**) or peer classmate (@GitHub-Username) regarding the Classwork 1 materials or need clarification on any points, don't hesitate to ask here. 

Let's collaborate and learn from each other!

