---
title: "`select()` and `rename()`"
subtitle: Classwork 6
date: last-modified
from: markdown+emoji

execute: 
  eval: true
  echo: false
---

For this Classwork 6, consider the following data.frame, `nyc_payroll_new`.

```{r}
#| echo: true
library(tidyverse)
nyc_payroll_new <- read_csv("https://bcdanl.github.io/data/nyc_payroll_2024.csv")
```

```{r}
#| echo: false
DT::datatable(nyc_payroll_new)
```

<br>

For detailed descriptions of the variables in this data.frame, please refer to the following link: [Citywide Payroll Data (Fiscal Year)](https://data.cityofnewyork.us/City-Government/Citywide-Payroll-Data-Fiscal-Year-/k397-673e/about_data)

<br>

## Question 1
Create a new data.frame, `df` that keeps only the following five variables---`Fiscal_Year`, `Agency_Name`, `First_Name`, `Last_Name`, and `Base_Salary`---from the data.frame `nyc_payroll_new`.

**Answer**:

```{r}
df <- nyc_payroll_new |> 
  select(Fiscal_Year, Agency_Name,
         First_Name, Last_Name,
         Base_Salary)
```


Here, we use the `select()` function from one of the packages in `tidyverse` to choose only the specified variables (`Fiscal_Year`, `Agency_Name`, `First_Name`, `Last_Name`, and `Base_Salary`) from the `nyc_payroll_new` data.frame. The result is assigned to a new data.frame name `df` containing only these variables, reducing the dataset for further analysis.

```{r}
#| echo: false
DT::datatable(df)
```

<br><br>

## Question 2
Given the data.frame `df` with a variable named `Agency_Name`, how would you rename it to `Agency`?

**Answer**:

```{r}
df_renamed <- df |> rename(Agency = Agency_Name)
```


The `rename()` function allows us to change the name of an existing variable. In this case, we rename the `Agency_Name` variable to `Agency` in the `df` data.frame. The result is a new data.frame, `df_renamed`, where `Agency_Name` is now simply `Agency`.

```{r}
#| echo: false
DT::datatable(df_renamed)
```

<br><br>

## Question 3
How would you remove the `Fiscal_Year` variable using `select()`?

**Answer**:

```{r}
df_dropped <- df |> select(-Fiscal_Year)
```


To remove a column from a data.frame, we use the `select()` function with the `-` sign before the variable name. In this case, `select(-Fiscal_Year)` removes the `Fiscal_Year` variable from the `df` data.frame, resulting in `df_dropped`, which contains all variables except `Fiscal_Year`.

```{r}
#| echo: false
DT::datatable(df_dropped)
```

<br><br>

# Discussion 

<font size = "5">Welcome to our Classwork 6 Discussion Board! 👋 </font>

This space is designed for you to engage with your classmates about the material covered in Classwork 6.

Whether you are looking to delve deeper into the content, share insights, or have questions about the content, this is the perfect place for you.

If you have any specific questions for Byeong-Hak (**@bcdanl**) or peer classmate (@GitHub-Username) regarding the Classwork 6 materials or need clarification on any points, don't hesitate to ask here. 

Let's collaborate and learn from each other!
