---
title: R Basics
subtitle: Classwork 1
date: 2024-09-16
from: markdown+emoji
---

## Question 1.
base-R provides the R object `state.name`. Write an R code to assign `state.name` to a variable, `US_states`.

*Answer*:
```{r}

```

<br><br>

## Question 2.
The `temp_F` vector contains the average high temperatures in January for the following cities: Seoul, Lagos, Paris, Rio de Janeiro, San Juan, and Rochester.

```{r}
temp_F <- c(35, 88, 42, 84, 81, 30)
```

Create a new vector named `temp_C` that stores the converted Celsius temperatures. Below is the conversion formula:

$$
C = \frac{5}{9}\times(F - 32)
$$

*Answer*:
```{r}

```

<br><br>


## Question 3.
- Write an R code to calculate the standard deviation (SD) of the integer vector `x` below manually. That is to calculate the SD without using the `sd()` function.

```{r}
x <- 1:25
```

- Also, write an R code to test whether the standard deviation you calculate manually above is equal to `sd(x)`.



*Answer*:
```{r}


```

<br><br>



## Question 4.
- Consider the vectors:

```{r}
my_vec <- c(-10, -20, 30, 10, 50, 40, -100)
beers <- c("BUD LIGHT", "BUSCH LIGHT", "COORS LIGHT", 
           "GENESEE LIGHT", "MILLER LITE", "NATURAL LIGHT")
```


- Write an R code to filter only the positive values in `my_vec`.
- Write an R code to access the beers that are in positions 2, 4, and 6 using indexing.


<br><br>



## Question 5.
- Write an R code to read the CSV file, `https://bcdanl.github.io/data/mlb_teams.csv` using the `tidyverse`'s `read_csv()` function, and assign it to `MLB_teams`.

*Answer*:
```{r}

```

<br><br>





## Question 6.
Write an R code to provide summary statistics---mean, standard deviation, minimum, first quartile, median, third quartile, and maximum---for variables in the `MLB_teams` data.frame.

*Answer*:
```{r}

```

<br><br>




# Discussion 


<font size = "5">Welcome to our Classwork 1 Discussion Board! 👋 </font>


This space is designed for you to engage with your classmates about the material covered in Classwork 1.

Whether you are looking to delve deeper into the content, share insights, or have questions about the content, this is the perfect place for you.

If you have any specific questions for Byeong-Hak (**@bcdanl**) or peer classmate (@GitHub-Username) regarding the Classwork 1 materials or need clarification on any points, don't hesitate to ask here. 

Let's collaborate and learn from each other!

