---
title: "`select()`, `rename()`, and `mutate()`"
subtitle: Classwork 5
date: last-modified
from: markdown+emoji

execute: 
  eval: true
  echo: false
---

For this Classwork 5, consider the following data.frame, `nyc_payroll_new`.

```{r}
#| echo: true
library(tidyverse)
nyc_payroll_new <- read_csv("https://bcdanl.github.io/data/nyc_payroll_2024.csv")
```

```{r}
#| echo: false
DT::datatable(nyc_payroll_new)
```

<br>

For detailed descriptions of the variables in this data.frame, please refer to the following link: [Citywide Payroll Data (Fiscal Year)](https://data.cityofnewyork.us/City-Government/Citywide-Payroll-Data-Fiscal-Year-/k397-673e/about_data)

<br>

## Question 1
Create a new data.frame, `df` that keeps only the following five variables---`Fiscal_Year`, `Agency_Name`, `First_Name`, `Last_Name`, and `Base_Salary`---from the data.frame `nyc_payroll_new`.

**Answer**:

```{r}
df <- nyc_payroll_new |> 
  select(Fiscal_Year, Agency_Name,
         First_Name, Last_Name,
         Base_Salary)
```

```{r}
#| echo: false

DT::datatable(df)
```


<br><br>

## Question 2
Given the data.frame `df` with a variable named `Agency_Name`, how would you rename it to `Agency`?

**Answer**:

```{r}
df_renamed <- df |> rename(Agency = Agency_Name)
```


```{r}
#| echo: false

DT::datatable(df_renamed)
```


<br><br>

## Question 3
Given the data.frame `df` with variable `Base_Salary`, how would you create a new variable `Adjusted_Salary` that is `Base_Salary` * 1.1 using `mutate()`?

**Answer**:
```{r}
df_new <- df |> mutate(Adjusted_Salary = Base_Salary * 1.1)
```



```{r}
#| echo: false

DT::datatable(df_new)
```

<br><br>

## Question 4
How would you remove the `Fiscal_Year` variable using `select()`?

**Answer**:

```{r}
df_dropped <- df |> select(-Fiscal_Year)
```



```{r}
#| echo: false

DT::datatable(df_dropped)
```

<br><br>

## Question 5
How would you select only `Agency_Name` (or `Agency`) and `Base_Salary` and then create a new variable `Bonus` which is 10% of `Base_Salary`?

**Answer**:
```{r}
df_new <- df |> 
  select(Agency_Name, Base_Salary) |> 
  mutate(Bonus = Base_Salary * 0.1)
```


```{r}
#| echo: false

DT::datatable(df_new)
```


<br><br>

# Discussion 


<font size = "5">Welcome to our Classwork 5 Discussion Board! 👋 </font>


This space is designed for you to engage with your classmates about the material covered in Classwork 5.

Whether you are looking to delve deeper into the content, share insights, or have questions about the content, this is the perfect place for you.

If you have any specific questions for Byeong-Hak (**@bcdanl**) or peer classmate (@GitHub-Username) regarding the Classwork 1 materials or need clarification on any points, don't hesitate to ask here. 

Let's collaborate and learn from each other!

