---
title: Joining `data.frame`s
subtitle: Classwork 3
date: last-modified
from: markdown+emoji
---


## Question 1
- Install the `nycflights13` R package and load it into your R session in your Posit Cloud project.

*Answer*:
```{r}

```

<br><br>

## Question 2
- The `nycflights13` package provides two data.frames: `flights` and `airlines`, which are related by the `carrier` variable.
  - `carrier`: A two-letter abbreviation indicating the full name of the airline.
  
- Use the `left_join()` function to create a new data.frame, `flight_airline`, that includes all observations and variables from the `flights` data.frame, along with the `name` variable from the `airlines` data.frame that corresponds to the `carrier` variable in the `flights` data.frame.

*Answer*:
```{r}

```

<br><br>



# Discussion 


<font size = "5">Welcome to our Classwork 3 Discussion Board! 👋 </font>


This space is designed for you to engage with your classmates about the material covered in Classwork 3.

Whether you are looking to delve deeper into the content, share insights, or have questions about the content, this is the perfect place for you.

If you have any specific questions for Byeong-Hak (**@bcdanl**) or peer classmate (@GitHub-Username) regarding the Classwork 1 materials or need clarification on any points, don't hesitate to ask here. 

Let's collaborate and learn from each other!

